\documentclass[pra,amsfonts,showpacs,showkeys,12pt]{revtex4-1}
%\documentclass[pra,showpacs,showkeys,amsfonts]{revtex4}
\RequirePackage[isolatin]{inputenc}
\usepackage{graphicx}
%\documentstyle[amsfonts]{article}
\RequirePackage{times}
%\RequirePackage{courier}
\RequirePackage{mathptm}
%\renewcommand{\baselinestretch}{1.3}

\begin{document}
%\sloppy



\title{Bob's search for Alice. A Platonistic Exit\\
       Outline of a movie plot}

\author{Karl Svozil}
\email{svozil@gmail.com}
\homepage{http://tph.tuwien.ac.at/~svozil}
\affiliation{c/o Institut f\"ur Theoretische Physik, University of Technology Vienna,
Wiedner Hauptstra\ss e 8-10/136, A-1040 Vienna, Austria}


\begin{abstract}
On a trip to Australia, through contact to {\em dreamtime,}
Bob, a New York stock broker, discovers that his world is a virtual reality.
Just two players are immersed inside this universe, while all the other agents are fake
characters simulated for the sole purpose to enrich the gaming zone with ornament.
Bob and the other gamer codenamed Alice represent their corresponding ``platonistic\footnote{
{http://www.gutenberg.org/dirs/etext99/sympo10.txt},      \\
``The two halves went
about looking for one another, and were ready to die of hunger in one
another's arms.~$\ldots$  The pair are
inseparable and live together in pure and manly affection; yet they cannot
tell what they want of one another.  But if Hephaestus were to come to them
with his instruments and propose that they should be melted into one and
remain one here and hereafter, they would acknowledge that this was the
very expression of their want.  For love is the desire of the whole, and
the pursuit of the whole is called love.  There was a time when the two
sexes were only one, but now God has halved them.~$\ldots$''
%
%
%German translation at URL\\
%{http://www.uni-essen.de/Ev-Theologie/courses/course-stuff/lit-plato-symposion188d.htm}:   \\
%Plato, Symposion (= Das Gastmahl). Reclam: Stuttgart. 188d-193c \S 16,    \\
%``Wenn nun ... jeder andre auf seine eigene Hälfte selbst trifft,
%dann werden sie wunderbar erschüttert von Freundschaft und Vertrautheit und Liebe und wollen voneinander nicht lassen,
%auch nicht einen Augenblick. Diese sind es auch, die gemeinsam das ganze Leben zubringen und nicht einmal
%zu sagen wüssten, was sie voneinander haben wollen. Denn es kann doch wohl nicht die Gemeinschaft des Liebesgenusses sein,
% deretwegen der eine dem andern sich so froh und mit so großem Eifer vereint,
%sondern etwas andres will offenbar die Seele der beiden, was sie nicht sagen kann,
%aber in Zeichen verkündet sie ihr Wollen und in Rätseln.~$\ldots$''
} counterparts.''

[[The movie might involve several such platonistic pairs; with a variety of fates.]]

The goal of the game is a twofold chain of events:
First, Alice  and  Bob have to find themselves.

Second, upon finding themselves, Alice and Bob will be
inseparable and live together in pure and manly affection until
they allow themselves to unite by making love and interlock their eyes in a final orgiastic ecstasy.
When that happens, {\it Satori} is achieved and their minds exit the game.
Otherwise, the players remain locked and stranded in the virtual reality
until they die lonesome and separately at a very old age;
doomed to be reborn to new life to start their search for each other anew.

[[The movie might involve several such generations.]]

After an initial period of denial and perplexity which transforms into agonizing dispair,
Bob sets out to search his world for Alice.
He looks for ``signs,''
screens internet databases and dating sites, performs spiritual practices,
roams public places such as bars and discotheques,
and travels in erratic zigzag around the globe.
First, he attempts many ways of seduction, but later realizes that ideally no seduction would be needed.
Because, he speculates that, once contact with Alice is established,
both of them would be drawn to each other in platonistic affection.
Alas it might be difficult to distinguish mere epistemic passion from the ``true'' ontic one;
and also to disentangle Alice from her present life and convince her to commit
to a destiny she might perceive as weird.
After all, Ann might not be aware of the situation.

After long unsuccessful attempts, disappointments and failures,
Bob stops searching and  relegates to existential philosophy.
He doubts his revelations; waits for miracles.
Then, suddenly, he sees a stranger across a crowded room.
Will this be  Alice, or just another woman?
\end{abstract}

%\pacs{42.50.Ar, 03.80.+r, 05.40.+j, 98.50.-v}
\keywords{Plato, platonistic love, virtual reality, dreamtime, }
\maketitle

%\bibliography{svozil}
%\bibliographystyle{apsrev}


\end{document}
