\documentclass[pra,amsfonts,showpacs,showkeys,12pt]{revtex4-1}
%\documentclass[pra,showpacs,showkeys,amsfonts]{revtex4}
\usepackage{graphicx}
%\documentstyle[amsfonts]{article}
\RequirePackage{times}
%\RequirePackage{courier}
\RequirePackage{mathptm}
%\renewcommand{\baselinestretch}{1.3}

\begin{document}
%\sloppy



\title{The Zoo Hypothesis revised\\
       Outline of a movie plot}

\author{Karl Svozil}
\email{svozil@gmail.com}
\homepage{http://tph.tuwien.ac.at/~svozil}
\affiliation{c/o Institut f\"ur Theoretische Physik, University of Technology Vienna,
Wiedner Hauptstra\ss e 8-10/136, A-1040 Vienna, Austria}


\begin{abstract}
Stuart Kaufman, a New Zealand physicist working at the University of Auckland, suffers from {\em akedia},
culminating in his escape to a lonely sabbatical at Matapouri, one of the seaside beaches of Northland's sunny coasts.
There, he is approached by a stranger: Amy, a woman who walks into his house and into his life,
turns out to be an alien ethnologist\footnote{Cf.
http://en.wikipedia.org/wiki/Zoo\_hypothesis and
John A. Ball,  "The Zoo Hypothesis," Icarus, 19, 347 (1973).} from outer space.

Rather than simply observing Earth from above,
Amy has decided to do some enjoyable field studies with Stuart, who caught her attention whilst publishing
some strange articles relating to  interstellar space travel on preprint archives.
Besides startup perplexities, Amy and Stuard go through a series of funny experiences,
such as surfing both Northland's beaches and Earth's atmosphere, as well as probing a lingerie shop.

As both have the privilege to live almost disentangled from their respective worlds whilst
being aware of the occurrences therein, they soon realize that Amy's world is endangered by a big intragalactic controversy,
whereas Earth is a conveniently unattractive outpost in the middle of nowhere with high chances of remaining unattended.
Because of this and her growing attraction to Stuart, Amy decides to stay with Stuart.
She quits pursuing her anthropological career and starts writing a book of fiction pertaining to the known history of the galaxy.
\end{abstract}

%\pacs{42.50.Ar, 03.80.+r, 05.40.+j, 98.50.-v}
\keywords{Cosmic romance, close encounter with alien culture, UFO}
\maketitle



%\bibliography{svozil}
%\bibliographystyle{apsrev}

\end{document}
