Bob: It is nice to see you. -  How does you thesis do?

Ann: Terrible! I should have taken the topic on Heidegger's metaphysics on why there is something rather than nothing.

Bob: Why? What did you take instead?

Ann: I selected quantum foundations instead.

Bob: sounds cool.

Ann: yeah that gave me shit.

Bob: why?

Ann: well I even drove to Smaland to meet some of those characters involved. Should have stayed at home instead; out of trouble!

Bob: what happened?

Ann: my adviser suggested to go to Sweden. Boy, he framed me! Already upon arrival, someone pushed me into a corner and told me freaky things.

Bob: huh ... like what?

Ann: like the others were all weird, strange quacks, and he was the only one knowing that there are no real observations.

Ann: The entire meeting looked like a communion of a sect of fucking aliens. Everybody was talking freaking nonsense. And assured me that all the others were nuts; completely wacky weirdoes.

Bob: what?

Ann: they constantly referred to guys whose name started with a capital B. Like Born, Bohr, Bohm, Bell, Bennett, and so on. Seems you don't qualify as a quantum hero unless your name starts with a B.

Ann: They kept telling me that nobody understands quantum mechanics anyway; and I should just shut up, stop thinking about how it could be like that,  and apply the established rules to calculate whatever. Else I would get 'down the drain', into a blind alley from which nobody has yet escaped.

Ann: But these rules ... first I thought they quoted Nostradamus   ... let me recall ...  "the impossibility of any sharp separation between the behavior of atomic objects and the interaction with the measuring instruments which serve to define the conditions under which the phenomena appear". When I heard that I knew I was in trouble.

Bob: Troubles? what kind of trouble?

Ann: They spoke about beables and QBism. And that quantum mechanics needs no interpretation. Except of course their own interpretation.

Ann: then they quoted quantum jellification, hidden variables, cats dead and alive at the same time, realism versus idealism.

Ann: someone mentioned quantum mechanics is magic.   Should have taken  Heidegger's fucking Sein und Zeit instead!

Ann: on one night they pretended to get drunk, although there was too little wine there, and the waiters did not wait.

Ann: But on this occasion, someone gave me a paper entitled: a map of madness. I guess at that point I must have completely freaked out.

Bob: what did you do?

Ann: I tried to escape. Took me some time. Swam across a lake... I could not take in this fucking quantum any longer. But finally, here I am.

Bob: Jesus ... and what will you do now?

Ann: I still see a chance to change subject to Foucault and his fucking relationship to Heidegger. Turns out to be much easier. If they will let me do it. Just have to learn German and French; sounds like peanuts compared to the quanta!

Bob: Oh my God
