%%%%%%%%%%%%%%%%%%%%% chapter.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% sample chapter
%
% Use this file as a template for your own input.
%
%%%%%%%%%%%%%%%%%%%%%%%% Springer-Verlag %%%%%%%%%%%%%%%%%%%%%%%%%%
%\motto{Use the template \emph{chapter.tex} to style the various elements of your chapter content.}
\chapter{Early UFO sagas and legends prior to Trinity (July 1945)}
\label{2023-UFO-part-History-chapter-pre-1945} % Always give a unique label
% use \chaptermark{}
% to alter or adjust the chapter heading in the running head


% https://www.afnwc.af.mil/About-Us/History/Trinity-Nuclear-Test/

% https://www.atomicarchive.com/media/photographs/trinity/index.html

\abstract*{The study of unidentified flying objects has a long history, stretching back to ancient times. Throughout history, people have reported sightings of mysterious aerial phenomena, including luminous discs, elongated objects, and glowing shields. These sightings have often been interpreted through religious or cultural lenses and have been used to legitimize political power.
In recent history, the perception of unidentified flying objects has been heavily influenced by technological advancements in aviation and the recognition that meteorites could fall from the sky.
``Foo fighters'' were sporadically recorded during World War II. Trinity, the explosion of the first nuclear device, seemed to be a significant turning point in the history of UFO sightings.}

\abstract{The study of unidentified flying objects has a long history, stretching back to ancient times. Throughout history, people have reported sightings of mysterious aerial phenomena, including luminous discs, elongated objects, and glowing shields. These sightings have often been interpreted through religious or cultural lenses and have been used to legitimize political power.
In recent history, the perception of unidentified flying objects has been heavily influenced by technological advancements in aviation and the recognition that meteorites could fall from the sky.
``Foo fighters'' were sporadically recorded during World War II. Trinity, the explosion of the first nuclear device, seemed to be a significant turning point in the history of UFO sightings.}



Ancient tales of entities in the sky abound and are difficult to evaluate and categorize.
Reports of UFO sightings can be found dating back to the earliest times and have been described as luminous discs, shields, globes, and elongated objects.

According to a RAND corporation report~\cite{Kocher-RAND-1968Jan},
based on studies by Brinsley Le Poer Trench~\cite{lePoerTrench1961Jan} and Jacques Vallee~\cite{Vallee1965Jan},
as well as more recent accounts by Jeffrey J. Kripal~\cite{Kripal2011Nov}, Diana Walsh Pasulka~\cite{Pasulka2019Feb},
and Tom DeLonge and Peter Levenda~\cite{DeLongeLevenda-Gods,DeLongeLevenda-Men}, these events were often interpreted in a religious context.

The repeated appearance of a kind of UFO phenomenology at Fatima, Portugal, in 1917 was witnessed by a crowd of 70\,000 people,
including scientists, reporters, and faithful Catholics.
The phenomenon was described as a luminous and shining disk with shifting tints that spun rapidly,
not like the Moon or the Sun seen through a fog.
In hindsight, when viewed through contemporary eyes, there is no material basis to assume that any of these sightings were of alien origin.

Kings and emperors in ancient times often justified their privileges and origins by claiming that they were descendants of sky people and sky crafts.
Religion and myths were often used in a symbiotic relationship with secular leaders,
as leaders received legitimacy from being blessed by supreme creatures from the heavens---``of God's grace''---in
exchange for supporting religion by worldly power.

In hindsight, could there be more to the ancient mythology of, for example, the Greek god Prometheus, who challenged the gods by stealing fire from them and providing it to humankind in different forms, such as technology, knowledge, and the broader concept of civilization? Perhaps some alien beings either created humankind or showed us mercy by revealing knowledge that we could use to our advantage. Alternatively, did the ancient gods amuse themselves by handing out weaponry and encouraging wars?
Any outrageous claims of this nature are likely to result in a drastically revisionist history.


In the last 200 years or so, the history of the perception of unidentified aerial
phenomena seems to have been heavily stimulated by three significant events.
To begin with, the first ``aerostatic'' flight by the Montgolfier brothers at Versailles in 1783, which
demonstrated that humans could be lifted off the ground using a lighter-than-air vehicle.
Second, the first flight using a steam-powered airship by Henri Giffard in 1852,
which demonstrated that it was possible to use a powered engine to propel an aircraft through the air.
And finally, the recognition that stones (aka meteorites) could fall from the sky.
Between 1794 and 1798, four falls of stones were witnessed and widely publicized~\cite{Marvin_1996}. In 1803, nearly 3\,000 stones fell from the sky at L'Aigle in Normandy.

Those events resulted in people looking very differently into the skies and led to a flurry of strange UFO sightings, many driven by projections of the mind and anxiety, which were exploited in the emergent news media market.

\section{An alien Sumerian civilization}
\label{2023-UFO-part-History-chapter-pre-1945-asc}

Carl Sagan may have been the first to spell out in a scientific article
that human civilization could have been significantly influenced by extraterrestrial contacts.
He speculated that~\cite[p.~496]{Sagan_1963} ``there are  $\ldots$ legends which  $\ldots$
deserve serious study in the present context [[of extraterrestrial contact with Earth during
historical times]]. As one example, we may mention the
Babylonian account of the origin of Sumerian civilization by the Apkdu [[or Apkallu or Abgal]]
of an advanced, nonhuman and possibly extraterrestrial society.''

Sagan quotes~\cite{Sagan_1963,Shklovskii1998Nov,Zygutis2016Nov}
the following fragment~\cite[pp.~56,57]{Cory1876Sep}:
\begin{svgraybox}
``At Babylon there was (in these times) a great resort of people of various nations, who inhabited Chaldea,
and lived without rule and order, like the beasts of the field.

In the first year there made its appearance, from
a part of the Erythr{\ae}an sea which bordered upon
Babylonia, an animal endowed with reason, who was
called Oannes.
[[$\ldots$]]
This Being, in the day-time, used to converse with
men; but took no food at that season; and he gave
them an insight into letters, and sciences, and every
kind of art. He taught them to construct houses,
to found temples, to compile laws, and explained to
them the principles of geometrical knowledge. He
made them distinguish the seeds of the earth, and
showed them how to collect fruits. In short, he
instructed them in everything which could tend to
soften manners and humanise mankind. From that
time, so universal were his instructions, nothing
material has been added by way of improvement.
When the sun set it was the custom of this Being to
plunge again into the sea, and abide all night in the
deep; for he was amphibious.''
\end{svgraybox}

Unlike the restrictive exchange of knowledge and technologies depicted in the ancient Greek legend of Prometheus,
the Babylonian tale portrays a more favorable and fortuitous series of events~\cite{Y_ld_r_m_2022}.
In the Babylonian narrative, the flow of knowledge and technologies
appears to be more open and benevolent.

\section{The Razing of Saubha: Krishna's Battle Against a Flying Fortress}
\label{2023-UFO-part-History-chapter-pre-1945-Saubha}
\index{Saubha, razing of}

``The Razing of Saubha'' is a tale that recounts the epic battle between Krishna,
an incarnation of the Hindu god Vishnu,
and Shalva, a rival king who possessed a powerful flying fortress called Saubha.
The story tells how Shalva, driven by anger over the death of a friend and enemy of Krishna,
laid siege to the city of Dvaraka, causing widespread destruction and loss of life by the deployment of Saubha.
Despite Krishna's absence, he eventually returned to defend his people and successfully defeated Shalva,
destroying the formidable Saubha in the process~\cite{vanBuitenen1975,DebroyBhagavataPurana3}.


The story is told in the Bhagavata Purana. In Bibek Debroy's translation~\cite[Chapter~10(76)]{DebroyBhagavataPurana3}:
\begin{svgraybox}
``Shalva $\ldots$ asked [[the lord and god Pashupati aka Shiva]] $\ldots$  for a vehicle that would be able to go anywhere at will and would be impenetrable by gods,
asuras, humans, gandharvas and rakshasas $\ldots$   terrifying.
$\ldots$
Maya, the conqueror of enemy cities $\ldots$
constructed such a city, made out of iron, and gave it to Shalva. [[``The city could fly through the sky and was known as Saubha''---from now on called Saubha.]]
$\ldots$ that vehicle $\ldots$ was enveloped in darkness. It could travel anywhere at will and was impossible to assail.
$\ldots$ Shalva used it to go to Dvaravati [[aka Dvaraka]].''
Shalva ``showered down weapons from [[Saubha,]] that supreme vimana [[a mythological flying palace or chariot]]
on the turrets, gates, palaces, mansions, terraces and pleasure gardens [[of Dvaraka]]. Rocks, trees, bolts of thunder, snakes and hailstones rained down.
There were fierce whirlwinds and the directions were enveloped in dust.
$\ldots$  Krishna's city [[of Dvaraka]] was thus severely afflicted by Saubha.   $\ldots$
However, [[Saubha]]
had been fashioned by Maya and was full of maya. It would appear as one and as many.
 It would be visible and would not be visible. The enemy found it impossible to determine where it was.
It was sometimes on the ground and sometimes, it was in the sky.
Sometimes it was on the summit of a mountain and sometimes it floated
on water. Saubha whirled around like a circle of fire and was never in a single spot.''
\end{svgraybox}
The Bhagavata Purana describes the annihilation of the flying fortress Saubha by the following terms~\cite[Chapter~10(77)]{DebroyBhagavataPurana3}
``Using a club, he shattered the enemy's Saubha. The club wielded by Krishna's hand shattered it into one thousand fragments and it fell down.''




A complementary narration of the same story~\cite[pp.~253-267]{vanBuitenen1975} can be found in the Vana Parva, The Book of the Forest, the third book of the Mahabharata. This version describes the downing of Saubha by Krishna using his Sudarsana Chakra, a spinning, celestial discus with 108 serrated edges, rather than a club.
Krishna commands the destruction of the flying fortress Saubha with a specific order~\cite[p.~266]{vanBuitenen1975}:
``~`Now smite with your prowess the Saubha and whoever are my enemies inside.' $\ldots$
It approached the now lackluster Saubha city and aloft it cut it in
two as a saw cuts a log. Severed in twain at the impact of the force of
Sudarsana,
[[Saubha]] fell down like the Three Cities razed by the Great God.
When Saubha had fallen, the discus returned to [[Krishna's]] hand. Whirling it
once more, [[Krishna's]] told it, `Now hurry to Shalva!' Just as Shalva was aiming
a heavy mace in that battle royal, the discus cut him in two, and it
blazed with its power.''


In the introduction to his translation of Vana Parva, The Book of the Forest, Johannes Adrianus Bernardus van Buitenen, a renowned Dutch Indologist at the University of Chicago,
described the Saubha story as an example of ``early science fiction''~\cite[p.~182]{vanBuitenen1975}.
He even speculated why Erich von D\"aniken, a Swiss author known for his controversial theories about extraterrestrial beings~\cite{vonDanikenKopp2015Apr}
(see also the earlier Robert Charroux~\cite{Charroux1971Mar}),
did not make more use of it~\cite[p.~202]{vanBuitenen1975}.


\section{The Second Punic War sightings, at 218--201 BC}
\label{2023-UFO-part-History-chapter-pre-1945-spw218-201BC}

%https://listverse.com/2018/03/04/10-ufo-sightings-from-very-early-history/

Throughout the Second Punic War between 218 and 201 BC, reports of strange aerial phenomena emerged, such as descriptions of gleaming phantom ships and shields in the sky. Livius critically noted that many of these reports were too readily believed~\cite{Livius-21-62}. The {\it annales maximi} of Rome also recorded several of these sightings, which frequently occurred during wartime and were observed by numerous individuals.


\section{Flaming Shield in 776}
\label{2023-UFO-part-History-chapter-pre-1945-fs776}

During a military attack on the town of Sigiburg (now Dortmund) in what is now Germany, a strange object resembling two large
flaming shields appeared in the sky. The object was described as being reddish in color and seemed to be floating in the air.
The advancing Saxon army was reportedly so frightened by the sight that they abandoned their siege of Sigiburg and retreated.

This event is recorded in the {\it Annales Laurissenses Maiores}, a series of Latin annals that document the history of the region from 741 to 829 AD. It is believed that the Annales Laurissenses maiores were compiled from accounts written by eyewitnesses at the time of the events described.

\section{Flying plate or hat observed hovering over a church in 1665}
\label{2023-UFO-part-History-chapter-pre-1945-fpohhoc1665}

Erasmus Francisci reported an ``air vision'' that was observed on April 8th, 1665~\cite{VonLudwiger1998Jan} in the city of Stralsund. A flat, round object resembling a plate or a hat was seen hovering over a church. It had the color of the rising moon and remained stationary until evening. The fishermen who witnessed the phenomenon were concerned and later reported experiencing tremors and pains in their limbs and head.

\section{Flying saucer with a cigar-shaped form of 1882}
\label{2023-UFO-part-History-chapter-pre-1945-fswcsf1882}

On the night of November 17, 1882, a UFO was reportedly observed by a number of distinguished scientists in the United Kingdom and over continental Europe~\cite[Chapter~7]{Menzel_1953}. The object was described as a flying saucer with a cigar-shaped form, and it was reportedly the best-observed flying saucer in history. The event is considered one of the most difficult to explain. E. Walter Maunder, a Greenwich astronomer, and J. Rand Capron, a spectroscopist, observed the event from England, while Dutch scientists professors Audemans and Zeeman also witnessed it. Zeeman eventually received a Nobel Prize for his work on magnetism.


\section{Mystery airship wave of 1896--1897}
\label{2023-UFO-part-History-chapter-pre-1945-masw18967}

A series of reports about mysterious airships began in California in 1896 and spread throughout the United States, with many of the sightings moving from west to east. Some of these reports included descriptions of visible occupants on the airships and encounters with the pilots. These airships were described as being larger, faster, and more advanced than any aircraft at the time, and some were even equipped with giant wings~\cite[p.~131,462]{Dash1997}. The reports of these airships from 1896 to 1897 were thoroughly investigated, with researchers examining nearly 1500 newspaper files from across the United States.

The occupants of these airships were often reported as appearing human. In some cases, the occupants claimed to be from Mars. Many of these reports were investigated and found to be misidentifications of planets and stars, hoaxes, or practical jokes, but a small number remain unexplained. Examples of unexplained incidents include an alleged airship crash in Aurora, Texas, on April 17, 1897, whose deceased pilot was reported to be a native of the planet Mars, and an alleged airship encounter involving calf-napping on the ranch of Alexander Hamilton at Leroy, Kansas, on April 19, 1897~\cite{Sanarov-81}.


%  The 1896-97 airship wave is probably the best-investigated of all historical anomalies. The files of almost 1500 newspapers from across the United States have been combed for reports, an astonishing feat of research. The general conclusion of investigators was that a considerable number of the simpler sightings were misidentifications of planets and stars, and a large number of the more complex the result of hoaxes and practical jokes. A small residuum remains perplexing. Clark, The Emergence of a Phenomenon, pp. 17-39; Brookesmith, UFOs: the Complete Sightings Catalogue (London, Blandford, 1995), pp. 22-3.



\section{Mysterious airships with searchlights of 1909}
\label{2023-UFO-part-History-chapter-pre-1945-mawsl1909}
\index{airships}


In December 1909, there were multiple reports of unidentified flying objects (UFOs) in the New York Tribune and the New York Sun. These reports recounted sightings of ``mysterious airships'' or other objects in the sky over different locations in Massachusetts and Connecticut. The objects were characterized as having bright lights or searchlights and were observed hovering or moving slowly through the sky~\cite[New~Lands, Chapter 15]{Fort1975-vd}.





\section{The War of the Worlds scare of 1938}
\label{2023-UFO-part-History-chapter-pre-1945-WW1938}

On October 30, 1938, the radio program ``The Mercury Theatre on the Air'' aired an adaptation of H.G. Wells' novel ``The War of the Worlds.'' Directed and narrated by actor Orson Welles, the program was presented in the form of a simulated news bulletin, causing widespread panic among listeners who believed that a Martian invasion was taking place.

The ``War of the Worlds'' broadcast remains one of the most famous events in radio history and serves as a powerful example of media's influence on public perception and belief. Despite causing widespread panic, the broadcast was a highly creative and well-crafted piece of storytelling that helped popularize the science fiction genre and establish Orson Welles as a prominent figure in entertainment.


\section{Crash retrieval in Cape Girardeau, Missouri, in the spring of 1941}
\label{2023-UFO-part-History-chapter-pre-1945-CGM}


In the spring of 1941, Reverend William Huffman, a Baptist minister, was allegedly summoned to the site of a plane crash outside of Cape Girardeau.
When he arrived, he found a strange craft---a saucer-shaped metallic object that had been broken open.
Inside, he saw a small metal chair, gauges, and dials, along with inscriptions and writings that resembled Egyptian hieroglyphics.
There were three entities or nonhuman people lying on the ground, two outside the saucer and one further away.
Reverend Huffman prayed over them, giving them last rites,
and observed that they appeared to have no bone structure and were soft-looking with crinkled, soft aluminum foil-like skin.
The entity he saw up close was approximately 4 feet tall with very long hands and fingers~\cite[pp.~34--43]{Wood2005Jan}.


\section{The Battle of Los Angeles of 1942}
\label{2023-UFO-part-History-chapter-pre-1945-boLA1942}
\index{Los Angeles, battle of}


The Battle of Los Angeles, also known as the Great Los Angeles Air Raid, took place on February 25, 1942, during World War II. The incident involved the firing of 1400 anti-aircraft shells at an unidentified flying object over Los Angeles, California.

At 2:25 a.m., air raid sirens sounded across the city, and witnesses reported seeing an object flying at high altitude over the city. Initially, thought to be a Japanese aircraft, the object was eventually identified as a UFO. The military fired at the object for several hours, but it is unclear whether any damage was inflicted.

The UFO disappeared from sight at approximately 7:00 a.m., marking the end of the incident.

The event has been the subject of speculation and debate, with some believing it was a genuine UFO encounter, while others suggest it was a false alarm or a misidentified object, such as a balloon, triggered by ``war nerves'' over fears of a Japanese attack. There is no conclusive evidence to support either explanation.




\section{Roman Sabinski's June 1942 encounter with a foo fighter}
\label{2023-UFO-part-History-chapter-pre-1945-RS42}\index{foo fighter}

Here is a dramatic story involving a foo fighter reported in the June 1992 issue of Mufon UFO Journal~\cite{MUFONUFOJ-290},
as well as in Keith Chester's book ``Strange Company: {M}ilitary {E}ncounters with {UFO}s in {W}orld {W}ar {II}''~\cite{Chester2007May}.
Another book on this topic is Graeme Rendall's ``{UFO}s Before {R}oswell: {E}uropean {F}oo-{F}ighters 1940--1945''~\cite{Rendall2021Aug}.
More unconfirmed dramatized foo fighter stories are told
in ``Chapter II: The Coming of the Foo Fighters'' of the book ``Flying Saucers on the Attack''~\cite{Wilkins1954Jan},
by the British journalist Harold Tom Wilkins, who specialized in all kinds of mysteries~\cite{Wilkins1946,Wilkins2017Aug}.


On June 25, 1942, Flight Lt. Ronan Savinsky and his Squadron 301 of the Royal Air Force Polish Division were flying over the
Ruhr Valley in Germany on a bombing mission.
After completing the mission successfully, they were returning over Holland.
Savinsky's rear gunner reported an approaching aircraft, which prompted Savinsky to instruct him to open fire
if the object got too close. Savinsky believed it to be a German fighter equipped with a searchlight
and was not willing to take any chances.

As the object approached from the right (starboard) side of the aircraft, the rear gunner opened fire,
aiming at a range of approximately 200 yards.
From their distance, the object might have been about the size of the Moon and was potentially even much larger than that.
While the rear gunner was shooting, the trajectory of the tracer shells could be followed as they entered the target,
However, the peculiar thing was that they did not fall away but instead simply entered and disappeared.

This lasted for approximately two minutes. After this time, the shiny object changed position and moved at terrific speed over the aircraft to the
other (left port) side,
remaining at approximately 200 yards from the wing.
This time, both the rear and front gunners were ordered to open fire on the object.
Again, the airmen observed that the tracers were entering the target from different angles and disappeared ``inside'' the object.

Eventually, the object took off at a fantastic
speed at a 45-degree angle and disappeared into the stars.
The aircraft behind the observer allegedly had a similar experience a few minutes later~\cite[Chapter~3,note~64]{Chester2007May} and~\cite{MUFONUFOJ-290}.


Here is another (undocumented) foo fighter report quoted by Wilkins~\cite[Chapter~II]{Wilkins1954Jan}:
On December 22, 1944, Lt. David McFalls and Lt. Edward Baker were flying at 10\,000 feet,
south of Hagenau in Alsace-Lorraine, the German Elsa{\ss} Lothringen, 20 miles north of Strasbourg, and 16 miles west of the Rhine.
McFall recalls:
\begin{svgraybox}
``At [[night]] 0600 (six pm) near Hagenau, at 10\,000 feet altitude, two very bright lights climbed toward
us from the ground. They levelled off and stayed on the tail of our plane. They were huge,
bright orange lights. They stayed there for two minutes. On my tail all the time. They were
under perfect control. Then, they turned away from us, and the fire seemed to go out.''
\end{svgraybox}
Two days afterwards, on December 24, 1944, the same crew had another strange encounter:
``A glowing red ball shot straight up to us. It suddenly changed into an airplane which did a wing over! Then, it dived and disappeared.''

