%%%%%%%%%%%%%%%%%%%%% chapter.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% sample chapter
%
% Use this file as a template for your own input.
%
%%%%%%%%%%%%%%%%%%%%%%%% Springer-Verlag %%%%%%%%%%%%%%%%%%%%%%%%%%
%\motto{Use the template \emph{chapter.tex} to style the various elements of your chapter content.}
\chapter{UFO sagas and legends after the Robertson Panel until 2003}
%\chapter{UFO legends after the Robertson Panel (January 1953) and before the USS Nimitz incident}
\label{2023-UFO-part-History-chapter-1953-2003} % Always give a unique label
% use \chaptermark{}
% to alter or adjust the chapter heading in the running head


% https://www.afnwc.af.mil/About-Us/History/Trinity-Nuclear-Test/

% https://www.atomicarchive.com/media/photographs/trinity/index.html

\abstract*{As the number of UFO sightings increased, so did public interest and media attention.
This led to the formation of organizations dedicated to collecting and investigating these events.
This chapter covers some of the most compelling and well-documented cases of UFO sightings, as well as some dramatic anecdotes.}


\abstract{As the number of UFO sightings increased, so did public interest and media attention.
This led to the formation of organizations dedicated to collecting and investigating these events.
This chapter covers some of the most compelling and well-documented cases of UFO sightings, as well as some dramatic anecdotes.}


It is safe to say that most sightings or incidents have prosaic causes rather than extraordinary or ``phenomenal'' origins,
as evidenced by sources such as Geipan~\cite{Geipan}, the Condon report~\cite{Condon-report-Bantam}, or David Clarke's review~\cite{Clarke2013Apr}.
However, neglecting the remaining less than 1{\%}
``hard cases'' could be a mistake~\cite{AA-Condon-1970} because they would carry important evidence and phenomenology.

For an example of a famous yet controversial case, Lonnie Zamora, a police officer in Socorro, New Mexico, claimed to have seen a landed craft of unknown origin and two occupants in 1964.
 Some believe the incident was a genuine UFO sighting, while others suggest that it was a hoax or misidentification of a known object.
Although the incident remains officially unexplained, it is one of the most well-documented UFO cases in history.
However, there is no concrete evidence to support any theory, and experts have not even settled on the symbol written on the craft. In contrast, a similar incident in Valensole one year later may be a genuine ``UFO sighting.''

Therefore, it is necessary to sort out reliable reports to distinguish them from unreliable ones. Being in the ``Russell's teapot'' scenario poses a challenge, as mentioned previously~\cite{Russell1952,Loewe-Christmas}. Making claims about extraordinary events such as alien landings or abductions is equivalent to claiming the occurrence of a Humean miracle, according to Sagan~\cite{Deming_2016}.
Any lack of complete or clear data or any single failure in the data can discredit the entire analysis in the eyes of skeptics.
My personal impression is that the more one looks at some of these sightings and anecdotes, some sightings get ``harder'' and ``more believable,'' gaining ``substance,'' while others tend to decay and fall apart, like sand in one's hand.

\section{Some literature}


For solid investigative journalism accounts on UFOs and related phenomena, I recommend
Ross  Coulthart's ``In Plain Sight''~\cite{Coulthart2021Aug},
Gideon Lewis-Kraus' New Yorker article ``How the Pentagon Started Taking {U.F.O.}s Seriously''~\cite{Lewis-Kraus2021Apr},
Ralph Blumenthal's ``The Believer: Alien Encounters, Hard Science, and the Passion of John Mack''~\cite{Blumenthal2021Mar},
as well as Leslie Kean's ``UFOs: Generals, Pilots, and Government Officials Go on the Record''~\cite{Kean2011Aug}.


For books of some early proponents of the extraterrestrial UFO narrative, I refer to the writings of
Donald Edward  Keyhoe~\cite{Keyhoe1950,Keyhoe1953,Keyhoe1955}
and James Edward McDonald~\cite{McDonald2022Aug,Druffel2003Sep}.

Some of the early critiques include
Donald H. Menzel~\cite{Menzel_1953,BoydMenzel1963Jan},
Philip J. Klass~\cite{Klass1975Jan,Klass1983Jan}.
A review of the literature can be found in Brenda Denzler's evaluation~\cite[p.~78]{Denzler_2001}.
For a contemporary ``debunker,''
containing a ``collection of accounts from people who have escaped from the rabbit hole,'' see Mick West's book ``Escaping the Rabbit Hole''~\cite{West2018Apr}.


For a critical discussion of some of the sightings until approximately 1968, I recommend a report by the Rand corporation,
authored by George Kocher~\cite{Kocher-RAND-1968Jan},
a collection of arguments for and against the extraterrestrial hypothesis~\cite{Sagan1974Jan},
a book by Stanton, L. Jerome~\cite{Stanton1966Jan},
as well as a collection for the National Investigations Committee on Aerial Phenomena (NICAP) by Richard H. Hall~\cite{Hall1964}.
Hall published a follow-up~\cite{Hall2001Jan} in approximately 2001.
An almost complete collection of APRO Bulletins from 1955 to the 1980s,
published by the Aerial Phenomena Research Organization (APRO),
\index{APRO}
\index{Aerial Phenomena Research Organization}
presents a rich set of UFO encounters~\cite{APROBulletins}.
For numerous sightings investigated by the National Investigations Committee on Aerial Phenomena (NICAP)
see their index page~\cite{NICAPIndex}, as well as their selection of ``good cases''~\cite{NICAPGC}.
See also the Sign Oral History Project~\cite{SOHPIndex} with its case studies.
``UFOs: An Insider's View of the Official Quest for Evidence''~\cite{CraigCondon1995}
is a personal account by Roy Craig, who experienced the Condon study as an investigator.


Get a glimpse into the experience of being part of the bigger (maybe second tier Jet-Set) ``UFO (and related) researcher''
community by reading Jacques Vall\'ee's ``Forbidden Science'' chronicles~\cite{ValleeFS1,ValleeFS2,ValleeFS3,ValleeFS4,ValleeFS5},
as well as Diana Walsh Pasulka's ``American Cosmic: UFOs, Religion, Technology''~\cite{Pasulka2019Feb}.
Those books offer some kind of ``window-shopping'' perspective into the ``UFO wonderland/candy shop.''



For a compilation of travels to exotic locations with supposed ``preastronautic'' connections, one can refer to Erich von D\"aniken's
extensive and popular body of work, which is said to have sold 70 million copies~\cite{vonDanikenKopp2015Apr}; in particular, I recommend his
``Reise nach Kiribati''~\cite{vonDanikenKiribati1981Jan} (in German).




\section{British Royal Air Force pilots ``see things'' in 1952-1953}

\label{2023-UFO-chapter-History--1953-2016-brafpst}


\subsection{Exercise Mainbrace UFOs in September 1952}

Exercise Mainbrace was a NATO military exercise that took place in September 1952. It was one of the largest military exercises held in Europe at the time, involving over 200 warships, 1,000 aircraft, and 80,000 personnel from various NATO countries. The aim of the exercise was to test the ability of NATO forces to defend against a potential Soviet invasion of Western Europe. The exercise included simulated amphibious landings, naval maneuvers, and air operations, and was seen as a show of force by NATO in response to Soviet aggression during the Cold War.
The following UFO sightings during Exercise Mainbrace follow a National Investigations Committee on Aerial Phenomena  (NICAP)
collection of Richard Hall~\cite{Hall2009Oct}.
\index{National Investigations Committee on Aerial Phenomena}
\index{NICAP}
\index{Mainbrace}

On September 13, Lieutenant Commander Schmidt Jensen and several members of the crew on the Danish destroyer Willemoes saw a triangular-shaped object emitting a bluish glow that moved at high speed towards the southeast. The object was estimated to be moving at over 900 mph.

%Between September 19 and September 21, there were four important sightings by well-qualified observers.

Just before 11 am on Friday, September 19, 1952
in ``the Yorkshire incident''~\cite{Clarke-Mainbrace-2020Dec}
\index{Yorkshire incident}
a British Meteor jet aircraft was being followed by a silvery object as it was approaching to land at Topcliffe, North Yorkshire, England.
According to a statement made by the Air Ministry's Flt Lt John Kilburn~\cite{AIR16bb1199,Team2022Sep,ClarkeTopcliff2022Sep},
\begin{svgraybox}
``the object was silver in colour and and circular in shape.''
It first seem to pursue the Gloster Meteor jet but then began to descend
``swinging in a pendular motion
$\ldots$
similar to a falling sycamore leaf.''
$\ldots$
``Suddenly it accelerated at an incredible speed towards the west turning onto a south-easterly heading before disappearing.
All this occurred in a matter of fifteen to twenty seconds.
The movements of the object were not identifiable with anything I have seen in the air and the rate of acceleration was unbelievable.''
\end{svgraybox}

In his memorandum for the record on``British Activity in the Field of UFOs''
Marshall H.  Chadwell, the Assistant Director of CIA's Office of Scientific Intelligence (OSI) stated dryly~\cite{Chadwell1952Brit,ClarkeTopcliff2022Sep}
\index{Chadwell, Marshall H.}
\begin{svgraybox}
``In some Royal Air Force (RAF) field [[RAF Topcliffe  in North Yorkshire]], there was some sort of demonstration to which high officials of the
RAF in London had been invited. During the show, a `perfect flying saucer' was seen
by these officials as well as RAF pilots.
So many people saw it that many articles appeared in the public press.
This is distressing to [[Dr (later Professor) Reginald Victor Jones who was Chadwell's opposite number in the British MoD]] because he realises that the creation of the correction of public opinion is a part of his responsibilities.''
\end{svgraybox}
Jones had a significant involvement with unexplained aerial phenomena, which began during World War II and continued in the aftermath of the war. He served as an intelligence officer for the Air Ministry, and he was tasked with investigating the sightings of "ghost rockets" in Scandinavia and parts of occupied Europe.
Following his success in this role, Jones was appointed as the Director of Scientific Intelligence for the UK's Ministry of Defence by Prime Minister Churchill. Jones was well-known for his scientific expertise, and his appointment was seen as a move to strengthen the country's military intelligence capabilities.
Throughout his career, Jones remained fascinated by unexplained aerial phenomena, and he continued to study and investigate these phenomena long after he left his position at the Ministry of Defence. His contributions to the field of scientific intelligence and his efforts to understand these mysterious sightings have had a lasting impact on our understanding of the world around us.


Around September 20, personnel on the USS Franklin D. Roosevelt, an aircraft carrier participating in the Mainbrace maneuvers, observed and photographed a silvery, spherical object moving across the sky behind the fleet. The pictures were examined by Navy Intelligence officers and were said to be excellent.

Also on September 20, three Danish Air Force officers at Karup Field observed a shiny, metallic disk that passed overhead from the direction of the fleet and disappeared in clouds to the east. The following day, six British pilots flying a formation of RAF jets above the North Sea saw a shiny sphere approaching from the direction of the fleet, but the object eluded their pursuit and disappeared. Later that month, on September 27/28, there were widespread UFO reports throughout Western Germany, Denmark, and southern Sweden, including sightings
of a brightly luminous object with a cometlike tail and a cigar-shaped object moving silently eastward.

\subsection{RAF Little Rissington incident on October 21, 1952 in the afternoon}

This incident was first published  by Air Marshal Sir Beresford Peter Torrington Horsley~\cite[Chapter~10]{Horsley1998Jan}, who served as Equerry to the Duke of Edinburgh.
whose narration is summarized in Section~\ref{2023-UFO-part-Perception-types-UK-dede}.
However, Horsley never could research this case.
David Clarke took up this task for the British Broadcasting Corporation (BBC) Radio 4 Series Britain's X-Files back in 2001.
Flt Lt (later Air Commodore) Michael Swiney, who was a student sitting in the front row of the Gloster Meteor VII jet,
along with his flight teacher Lieutenant David Crofts on the back, shared the following details with Clarke in this BBC interview~\cite[time=43m00s]{Team2022Sep}:
\begin{svgraybox}
``We broke through some
Cloud at about 14 000 feet and,  in front of me, in front of the aircraft there appeared to be what I took to be perhaps
three people coming down in parachutes. So I grabbed control of the aircraft from the student instructor [[Crofts]]
behind me to make sure that if they were parachutes we didn't go through the middle of them.
$\ldots$ They were directly in front of the windscreen and that is why I  took the aircraft
to one side to make sure that I didn't go up through them $\ldots$
but it was soon quite apparent that um that they weren't but there were three um circular white or off-white discs [[which appeared like the bottoms of wine glasses]]
straight in front of me,
two on a level heel and one that is slightly canted [[tilted]] to one side.
I can't remember which side that was $\ldots$ I have no idea how far away they were because um I didn't know what size they might have been,
but I know that they were above us and we continued climbing until such time as they lost their disc
appearance and took on a sort of flat plate.
$\ldots$
I's 50 years ago $\ldots$ I don't honestly remember but it was quite a long time it was minutes rather than seconds.

[[Instructor Crofts behind Swiney]]
certainly did [[see the craft]]  because um I managed to put my head to one side and he was able to look past my the open shoulder as it were,
and he could quite clearly see what I could see.
$\ldots$  when we spotted them we were of about uh 14000 feet I think and we continue the climb until we've got to  35000.
When we just were trying to explain  what we were looking at we were trying to tell ourselves that there was some funny light reflections or refractions
in the windscreen or in the  canopy but clearly it wasn't we were saying something.
$\ldots$  We tried very hard to explain away   what we were looking at  but
there's no way we could do that um there was there was something there.
$\ldots$
that absolutely no doubt about it was not a reflection.
$\ldots$ At the time
I didn't feel frightened, I did subsequently feel that we were looking at something that really we shouldn't be seeing
and I remember being told on landing by um either the ground crew or somebody who came out to the aircraft
to meet us that I looked fairly shaken---almost as if I'd seen a ghost''
\end{svgraybox}
Radar had also picked up a signal:
The RAF Southern Sector, headquartered in Wiltshire, detected the entry of three objects into their airspace at a velocity of 3,000 miles per hour.
ATCC Gloucester, located at what is now Gloucestershire Airport, verified the sighting of three objects reported by the pilots.
RAF Fighter Command dispatched two Meteor aircraft to intercept the objects, which were traveling eastward at a speed of 600 knots.
However, the fighter planes were unable to establish any further contact with the objects.



\section{Hearsay about Wernher von Braun 1959}

\label{2023-UFO-chapter-History--1953-2016-hswvb59}
\index{von Braun, Wernher}


Wernher von Braun was a member of the prewar NSDAP in Nazi Germany since 1937 and was promoted to SS-Sturmbannf\"uhrer in 1943. He was later transferred to the USA as part of the Operation Paperclip, as detailed in~\cite{Jacobsen2014}. He continued his work on rocket propulsion, which eventually led to the use of the Saturn family of rockets in NASA's Apollo program.

In various UFO talk shows and publications, von Braun is (mis)quoted in the UFO context.
It is based on some ``Dr. Stalter from New York'' who
allegedly relayed a statement attributed to von Braun.
However, Dr. Stalter is typically not mentioned as the source of the statement, which goes as follows:
\begin{svgraybox}
``Wernher von Braun:
``There are extraterrestrial forces that are stronger than we previously believed.''

``We seek connection with them''

The German-American rocket scientist Wernher von Braun, the mastermind behind the launch of ``Juno~2'', declared after its deflection from its orbit to the moon:
There are extraterrestrial forces, the seat of which is still unknown to us, but which are much stronger than we have assumed so far.
I am not allowed to say more about them.
In a few months, however, we will be able to be clearer. We are at present in the process of coming into closer contact with these extraterrestrial forces.''

Dr . Stalter, New York
\end{svgraybox}

The original German text, found in an esoteric publication called ``Neues Europa,'' extensively also discusses supposed visits from ``Venusians,'' that is, beings from Venus.
Here is the exact quotation of the article in the magazine reporting von Braun's alleged statement~\cite{vonBraun1959}:
\begin{svgraybox}
Wernher von Braun:
``Es gibt au{\ss}erirdische Kr\"afte, die st\"arker sind als wir bisher glaubten''

``Wir suchen Verbindung mit ihnen''

Der deutsch-amerikanische Raketenforscher Wernher v . Braun, der leitende Kopf bei dem Abschu{\ss} der ``Juno~2'', erkl\"arte nach dessen Ablenkung aus seiner Bahn zum Monde:
,,Es gibt au{\ss}erirdische Kr\"afte, deren Sitz uns noch unbekannt ist, die aber weitaus st\"arker sind, als wir bisher annahmen.
Mehr darf ich dar\"uber nicht sagen.
In wenigen Monaten aber k\"onnen wir deutlicher werden. Wir sind zur Zeit dabei, mit diesen au{\ss}erirdischen Kr\"aften in n\"ahere Verbindung zu kommen.''

Dr . Stalter, New York
\end{svgraybox}


Such messages may be dismissed as rumors, but it is difficult to overlook the recollections~\cite{SmithWilbertNotes} of the late Robert Irving Sarbacher,\index{Sarbacher, Robert Irving}
as mentioned in Section~\ref{2023-UFO-part-Perception-crash-retreivals-SM}.
According to Sarbacher, it was highly likely that Wernher von Braun may have been consulted
on matters relating to retrieved spacecraft if indeed such craft had been retreived.


\section{Socorro policeman Zamora's encounter in 1964---a student prank?}

\label{2023-UFO-chapter-History--1953-2016-SZ64}
\index{Socorro incident 1965}

According to Project Blue Book~\cite[1964-04-8694587-Socorro-NewMexico]{bluebook-directory-listing}, on April 24, 1964, a Socorro, New Mexico policeman named Lonnie Zamora reported sighting an object about a mile south of the town at approximately 5:45 pm in an unpopulated area full of hills, gullies, and sagebrush. While chasing a speeding car north on US 85, Mr. Zamora heard a roar and saw flames in an area where a dynamite shack was known to be located. He abandoned the chase of the car and proceeded to where he believed an explosion had occurred. After traveling a little-used road and experiencing considerable difficulty trying to drive his car up a gravel-covered hill, he observed what he thought was an overturned car standing on end. He was approximately 800 feet away from the object, and his car was at the crest of a hill with the object ahead of him in a gully. During this first glance, he saw one or two figures in coveralls whom he assumed to be occupants of the object. He did not see them again.

After radioing to Police Headquarters at Socorro that he was proceeding to investigate what he believed to be an auto accident, he drove to a point approximately 150 feet from the gully where the object rested and stopped the car to proceed on foot. He said the object was white, egg or oval-shaped [[like a rugby ball]] and apparently supported on girder-like legs. He heard a roar and saw smoke and flame coming from the bottom of the object. Mr. Zamora believed that the object was about to explode, and he became frightened, turned, and ran to shield himself behind the police car, bruising his leg and losing his glasses on the way. He crouched down, shielding his eyes with his arm while the noise stopped and he glanced up. The object had risen to a point approximately 15-20 ft above the ground, and the smoke had ceased. At this point, he noted a design on the object, which he described as markings in red approximately 1 to 1.5 ft in height, shaped like a crescent with a vertical arrow and horizontal line underneath. The object remained stationary for several seconds and then flew off in a southerly direction following the contour of the gully.

Sgt. Chavez of the New Mexico State Police arrived on the scene in response to Mr. Zamora's earlier radio call. He observed no object, but he reported that there were some slight depressions in the ground and apparently burned brushes in the area where Mr. Zamora had reported seeing the object. The brush was cold to the touch. Sgt Chavez reported the incident to local military authorities who conducted the initial investigation. The Air Force sent investigators from their Air Technical Intelligence Centre (ATIC, handling Project Blue Book requests) at Wright-Patterson AFB, Ohio, who found that no other witnesses to the object reported by Mr. Zamora could be located. There were no unidentified helicopters or aircraft in the area. Observers at radar installations had observed no unusual or unidentified blips. There was no unusual meteorological activity and no thunderstorms, and the weather was windy but clear.


The story takes a twist with claims that the Socorro incident was a hoax, potentially directed at Socorro Policeman Lonnie Zamora and committed by students at the New Mexico Institute of Mining and Technology. In a letter to Linus Pauling dated June 19, 1968, the institute's president, Stirling A. Colgate, made this claim (although it has been contested by some~\cite{Moss2021Jan})~\cite{Bragalia-hoax}.

There is also ambiguity regarding the symbolism on the craft. The body of the craft allegedly had a couple of lines drawn as a symbol.
It is unclear whether Hynek\index{Hynek, J. Allen} or Captain Richard Holder
asked Zamora and others not to reveal the true symbol so that potential hoaxers would reveal themselves.
As a result, there are various alleged symbols, including an inverted ``V'' under an ``umbrella''
type half-circle with a horizontal baseline under the inverted ``V.''
An alternative symbol is a line within the inverted ``V'' stretching from half the baseline to the top of the inverted ``V.''
This information is contained in my version of the Project Blue Book files~\cite[1964-04-8694587-Socorro-NewMexico]{bluebook-directory-listing}.
However, another popular symbol is formed by three horizontal bars of equal length, stacked upon each other and overlaid by an inverted ``V.''

For a recent investigation of the case, see also Chapter~4 of Daniel Coumbe's ``Anomaly'' book~\cite{Coumbe2022Oct}.

\section{A lavender farmer's encounter in Valensole, France, in 1965}

\label{2023-UFO-chapter-History--1953-2016-val65}
\index{Valensole incident 1965}

The following event was later recounted by GEIPAN, the French government study and information group on unidentified aerospace phenomena~\cite{Valensole-GEIPAN}.

On July 1, 1965, at 5:45 a.m., Maurice Masse, a farmer, was working in his lavender field near Valensole in the Provence-Alpes-C\^ote d'Azur department of Alpes de Haute Provence when he heard a whistle that caught his attention. He went to investigate and found a strange, dark machine on the ground that was shaped like a rugby ball. At first, he thought it was a Renault Dauphine. He watched as two individuals entered the machine, which quickly took off and disappeared toward Manosque. Approximately ten minutes later, he noticed traces on the ground where the machine had been, in the shape of a star with a cylindrical hole in the middle, on soaked ground. He returned to the location with his daughter in the evening at approximately 8:30 p.m. and noticed that the ground had hardened like cement. Ren\'e Gouin, a neighboring farmer visiting the site, noticed that in a field close by, the machine had burned the blossoming lavender as it flew away: first, many plants were heavily damaged, and then fewer plants were burned as the machine flew higher up~\cite[time=23min]{ExoMagazinTV2020Aug}.

Shortly afterwards, on July 4, 1965, a theory that it was a helicopter was put forth in an article in Le Monde. Military circles considered the machine to be a light army helicopter, most likely of the Alouette II or Alouette III type. There were known maneuvers called ``Provence 65'' taking place in the region since June 29, which included the participation of this light aviation. Since Valensole is only 19 km away from Manosque, the eastern border of the maneuver zone, it is almost certain that an aircraft flew over Valensole, and a helicopter could very well have landed in Maurice Masse's field.

On August 18, 1965, one and a half months after the event,
maybe because he previously was afraid that the police would consider him insane if he had told them the entire happening,
the farmer approached the police and reported very strange additional information.
He reported that, as he approached the craft and the two figures
standing outside of it, he got about seven meters away from them.
At that moment, one individual facing the farmer noticed him. This one must have given a signal to the other, and both stood up.
The one with his back to the farmer turned around and made a gesture with his right hand.
In it, the farmer saw an object. At that moment, he became paralyzed, completely immobile, and could not make a single movement~\cite{Niemtzow-paralysis-1975}. He felt nothing. He was neither numb nor paralyzed. He simply could not move his head, arms, legs, or body. He described the strangers as being approximately 1.2 meters tall, not larger than children of 10 years of age, with a very big head in comparison to the rest of the body. After approximately 20 minutes, he allegedly could move again.

Some believe that the farmer experienced a ``missing time'' after this encounter. Others claim that he saw a UFO a second time, this time hovering above him as he prepared to water young lavender plants. He only told his wife of this event, and she remained silent until after his death, seeing his ontological shock~\cite{ExoMagazinTV2020Aug}.

\section{``Foo fighters'' sightings by the chief pilots of Lufthansa and Austrian Airlines}

\label{2023-UFO-chapter-History--1953-2016-scpla}
%\index{Brazil's night of UFOs 1986}

\index{foo fighter}

The following reports, translated from German, are accounts given by Werner Utter,
chief pilot of Lufthansa\index{Lufthansa}, and
Alexander Raab, chief pilot of Austrian Airlines\index{Austrian Airlines} (AUA).
These reports were collected by Andreas M\"uller~\cite{Muller2021Nov} and by Mufon~\cite{ExoMagazinTV2013Feb}. Utter was one of the most experienced German pilots of his time. During WWII, he flew the first jet aircraft for the Luftwaffe, and after the war, he flew many heads of (West) Germany on official state visits. Later, Utter became a member of the Board of Directors for Lufthansa and was responsible for flight operations.

%\begin{svgraybox}
%Utter:
%``[[In the 1950's]] on a flight from Beirut to Baghdad, as we were flying at approximately 3000 meters altitude and had crossed over Lebanon,
%I suddenly saw a light in the sky ahead on the left.
%From my experiences in the last war I first thought it is a light bomb.
%This light got closer and closer very fast until I turned on the cockpit lights,
%thinking that another plane was approaching. In addition, suddenly, a huge fireball was next to the cockpit and the inner engine,
%about 5 meters in diameter. It was a warm, reddish light. The whole fireball was moving.
% I wasn't afraid of it at all. I was completely absorbed by this light.
%I looked at my co-pilot, he nodded, he saw it too, it was daylight inside the cockpit.
%When I looked again, this light quickly and abruptly disappeared into the sky -- and we were back under the stars.''
%\end{svgraybox}
%
%% Utter: Auf einem Flug von Beirut nach Bagdad, als wir in etwa 3.000 Metern Höhe waren und den Libanon überflogen hatten, sah ich plötzlich am Sternenhimmel eine Lichterscheinung, wie etwa eine Leuchtbombe. Diese Lichterscheinung kam nähcr und nähen so dass ich ganz aufgeregt die Bordscheinweil'er einschaltete, ich dachte, es käme uns ein anderes Flugzeug entgegen. Und plötzlich stand ein riesiger Feuerball neben dem Cockpit und dem Innenmotor, etwa im Durchmesser von 5 Metern. Es war ein warmes, ein rötliches Licht. Der ganze Fezierball bewegte sich. Ich hatte überhaet keine Angst davor. Ich war ganz eingenommen von dieser Lichterscheinungo Ich sah zu meinem Copiloten, er nickte, er sah es auch, es war la taghell im Cockpit. Als ich wieder hinsah, ging diese Lichterscheinung mit einer ungeheuren Geschwindigkeit abrupt in den Himmel -- und wir waren wieder unter den Sternen.''

\begin{svgraybox}
Utter:
``[[In the 1950's]] My first conscious experience was on a flight from Beirut to Baghdad,
on 3000 meters altitude, and it was at that time a Super Constellation,
one of the best three-blades per engine aircraft with four engines.
And as we initially took off and had to make a right turn over the Mediterranean and fly over Lebanon, and as we were almost at the crest of Lebanon,
I suddenly saw a very bright light on the left ahead of me.
And I immediately remembered my experiences during the last war as a pilot and thought to myself, someone is throwing a flare.
The light fascinated me.
I looked at the light and it was coming toward me at such a speed that I initially turned on my headlights because I thought it could not see me.
I told the copilot: watch out, something is coming toward us.
And it all happened in seconds and suddenly it was there,
a fireball, I would say, a fireball that was not dead but alive, pulsating, and it was suddenly next to me between
the left inner engine and the cockpit, which means you could not see exactly how far it extended.
It flew around me and it was daylight inside the cockpit and I looked at the copilot,
he saw it too and we did not say a word,
we were so amazed by something like that and while we were staring into this fireball,
it flew or shot, I would say, with an eerie speed at approximately 45 degrees in front of me,
upward, getting smaller and disappearing and it was dark and that was it.

And the three of us in the front of the plane, we were so impressed that we could not talk to each other properly for a while.
In Baghdad, we took some time to recover from the whole thing. We could not explain anything at all.

We did not know that at that time there were already known UFO sightings -- I certainly was not aware of any of it.~\cite{ExoMagazinTV2013Feb}''
\end{svgraybox}

%Original German:

% Utter: mein erstes bewusstes Erlebnis hatte ich auf einem Flug von Beirut nach Bagdad und das war damals eine Super Constellation,
%eines der besten drei-rotorigen Flugzeuge mit vier Motoren.
%Und als wir  zun\"achst starten heraus und musste eine rechtskurve über dem mittelmeer  gewinnt und den libanon zu überfliegen und als wir dann fast auf dem
%scheitelpunkt des libanon waren sah ich plötzlich vor mir  voraus ein sehr helles licht.
%
%Und ich erinnerte mich sofort an meine erlebnisse im letzten krieg als flieger und dachte mir: irgendwo wirft da einer eine Leuchtbombe.
%Es hat mich fasziniert.
%ich hab das Licht angesehen und es kommt auf mich zu; und zwar mit einer solchen geschwindigkeit,
%dass ich zunächst mal meine scheinwerfern anmachte weil ich dachte es sieht mich nicht.
%Den kopiloten  sagte ich: pass mal auf, es kommt was auf uns zu.
% und es geht alles sekunden und plötzlich war es da und es stand ein feuerball möchte ich sagen ein feuerball der nicht tot war,
%sondern er lebte , der pulsierte stand dann plötzlich neben mir zwischen dem linken inen motor und der kanzel, das heißt man kann sich genau sehen
%wie weit er nach außen ging dieser feuerball.
%Er flog mit mir rum und es war tag hell im cockpit und stufte de Copiloten an, der sah es auch, und wir sprachen kein wort,
%waren w bass erstaunt über so etwas und während wir so in diesem feuer ball hinein starten, flog oder ging das ging der feuerball möchte ich sagen mit einer
%unheimlichen geschwindigkeit etwa mit 45 Grad vor mir nach oben wurde immer kleiner und verschwand und dunkel und das war's.
%
%und wir drei vorne fahren wir waren dermaßen beeindruckt das wir schwer miteinander reden konnten in der schlucht ein paar mal tief.
%In Bagdad haben wir selbst mal erholt von dem ganzen .
%wir konnten uns überhaupt nichts erklären.
%
%Dass damals schon das  irgendwie ufo-sichtungen bekannt waren wußten wir nicht--- mir war auf jeden fall nichts davon bekannt.


Here is Utter's account of his second sighting that took place on November 21st, 1978 over Labrador:
\begin{svgraybox}
Utter: ``[[On November 21st, 1978]] this was over Labrador, at an altitude of approximately 12,000 meters on a flight from Frankfurt to New York.
It seems we had a flying saucer in sight. Very bright, pulsating and emitting rays, sometimes red, sometimes white, sometimes violet, such as jellyfish or spider.
I was just as fascinated by this light appearance, but there was not the warmth there as in the first appearance.''
\end{svgraybox}
The sighting was preceded by a radio message from a TWA plane reporting a light, a possible UFO,
that was supposed to cross the course of Utter's plane.

% Utter: Das war über Labrador, in einer Höhe von rund 12,00 Metern aufs einem Flug von Frankfurt nach New York. Es scheint wir haben eine fliegende Untertasse in Sicht. Ganz hell und sendet Strahlen aus, mal rot, mal weiß, mal violett, wie so eine Spinne. Ich war von dieser Lichterscheinung genauso fasziniert, aber es war nicht die Wärme da wie bei der ersten Erscheinung. Der Sichtung ging ein Funkspruch von einer TWA-Maschine voraus, der von einem Licht, einem möglichen UFO, berichtete., daseigentlich den Kurs von Utters Maschine kreizen sollte.



Utter's third encounter was on a cargo flight approximately in 1978:
\begin{svgraybox}
Utter: ``You may laugh about it, but I have it on tape.
I was returning from New York in the evening on a cargo flight,
it was still daylight. Suddenly, a huge thing comes toward us.
The flight engineer, who was standing behind me and talking to the copilot,
quickly shouted---you can hear it on the tape---`watch out', and then I said, 'what is that?'
It was a cucumber or cigar, coming toward us at 35,000 feet (12,000 meters),
and I immediately said, `that is a UFO.'
The air traffic controller below saw nothing on radar~$\ldots$
 I thought it was going into my one engine (left outboard engine).
If I had seen it alone,~$\ldots$ but the engineer also saw it and shouted the alarm.''
\end{svgraybox}

% Utter: Sie mögen darüber lachen, aber ich hab's auf Tonband. Ich kam abends von New York zurück mit dem Frachter; es war noch taghell. Da kommt mir plötzlich so ein Riesending entgegen. Der Flugingenieur, der hinter mir stand und sich mit dem Copiloten unterhielt, der ruft sehr schnell das hört man auf Tonband, ,Achtung', und dann ich ,Was ist das?: Das war eine Riesenzigarre, die uns entgegenkam in 35.000 Fuß (12.000 Metern) und ich gehe gleich runter und sag: ,Das ist ein UFO'. Da sagt der Fluglotse unten, er hoffe, dass es keine fliegende Unterasse' gewesen sei... Ich dachte, es geht mir in den Einser-Motor rein (linkes Außentriebwerk). Wenn ich das alleine gesehen hätte, ... aber der Ingenieur hat es auch gesehen und diesen Notschrei ausgestoßen.''



Alexander Raab, who was chief pilot of Austrian Airlines (AUA) in the 1960s, also reported a UFO sighting.
The incident occurred in the evening hours of March 18, 1972, during a flight with an AUA DC-9 from Vienna to Frankfurt,
at approximately 6000 meters above the provincial capital of Linz, Upper Austria.
Eight minutes before the sightings, the on-board compasses went crazy and deviated by several degrees~\cite{Raab-Kletter-Linz}.

Raab's aircraft was overtaken by a cone-shaped flying object~\cite[p.~165]{Muller2021Nov}.
According to his account, the ``fiery body'' was in the shape of a funnel, with a downward-pointing tip
that emitted a dazzlingly bright light that turned reddish later~\cite{Raab-Kletter-Linz}.
Raab saw the UFO together with his copilot Otto Herold for approximately half an hour~\cite[p.~165]{Muller2021Nov}.

In the first moment, he had made ``no comparisons,'' in particular, not with a UFO.
Instead, he first thought of a meteorite. However, the angle of entry was much too steep.
The funnel itself had clearly visible contours~\cite{Raab-Kletter-Linz}.

``Putting up any theory, I cannot because I think too realistically.''
He explained his reservations with respect to the observed phenomenon.
Additionally, he had ``not thought of Martians'' during the sighting.

Two minutes later, he said, the apparition suddenly changed direction and now overtook the AUA aircraft on the left,
before it disappeared from the field of vision of the two pilots on the horizon.

When asked by a journalist if the sighting could have been compared to satellites,
Raab reported that he had already seen numerous satellites from the cockpit but that the UFO appearance was ``completely different.''
\begin{svgraybox}
``The thing was much bigger from my point of view.
It was to the left of the aircraft.
In my opinion, it had nothing to do with satellites or the remnant of a rocket stage.
($\ldots$) By the way, my copilot and I had not been the only observers.
This object had also been sighted at the same time by crew members of a Cessna as well as a Lufthansa aircraft [[Brouwer above Radstadt?]].''
\end{svgraybox}

% Raab: Auch Chefpilot der Austrian Airlines war UFO-Zeuge Neben Utter als erstem Chefpiloten der Deutschen Lufthansa, wusste auch Alexander Raab, der in den 1960-er Jahren Chefpilot der Austrian Airlines (AUA) war, in späteren Jahren von UFO-Sichtungen zu berichten. Wie aus damaligen Medienberichten und aus einem Interview Raabs mit dem Wiener Buchautor Peter Krassa (Wien 1 980) liervorgeht, ereignete sich der Vorfall in den Abendstunden des 18. März 1972 während eines Fluges mit einer DC-9 der AUA von Wien nach Frankfurt. Rund 6.000 Meter über der Landeshauptstadt I,inz wurde Raabs M aschi iicli iiks von einem kegelförmigen Flugobjekt überholt. Der feurige Körper'' hatte demnach die Form eines trrichters, mit einer nach unten gerichteten Spitze, die ein gleißend helles Licht ausstrahlte. Raab sah Glast JF0 gemeinsam mit seinem Kopiloten Otto Icrold etwa eine knappe halbe Stunde(!) lang. Im ersten Moment habe er "keine Nrgleiche'' angestellt, auch nicht direkt an ein UFO, stattdessen zunähst an einen Meteoriten gedacht. Allerdings sei der Eintrittswinkel viel zu steil gewesen, bei. selbst habe keine wirklichen Konturen gehabt. "Irgendeiner2ltel aufstellen, das kann ich nicht, da ich zu realistisch denke", erklärte re seine Distanz zu dem beobachteten Phänomen. Auch habe adael, während der Sichtung "nicht an Marsmenschen'' gedacht. Zwei Minuten später habe die Erscheinung plötzlich die Richtung verändert und überholte die AUA-Maschine nunmehr zur Linken. Bevor sie am Horizont aus dem Blickfeld der beiden Piloten verschwand, spielten die Bordkompasse verrückt und wichen um sieben Grad ab.

% Raab: Auf die Nachfrage des Journalisten, ob die Sichtung mit Satelliten zu vergleichen gewesen seien, berichtete Raab, dass er bereits zahlreiche Satelliten aus dem Cockpit heraus gesehen habe, die UFO-Erscheinung aber ``völlig anders'' gewesen sei. "Das Ding war aus meiner Sicht heraus viel größer. Es war links neben dem Flugzeug. Meiner Meinung nach hatte es mit Satelliten oder dem Rest einer Raketenstufe nichts zu tun. (1 Übrigens waren mein Kopilot und ich nicht die einzigen Beobachter gewesen. Dieses Objekt war zur selben Zeit auch von Besatzungsmitgliedern einer Cessna sowie einer Lufthansa-Maschine gesichtet worden.''


Afterwards, Leopold Kletter, the head of the Austrian weather service, declared that the luminous phenomenon undoubtedly had natural and prosaic causes in the form of a large meteor that broke up in the upper atmosphere and caused many light phenomena,
including those observed above Radstadt and in multiple locations in southwest Europe~\cite{Raab-Kletter-Linz}.

Recently the experienced pilot and photographer Christiaan van Heijst~\cite[time=4225s=1~h:10~min:27~s]{Heijst2023Feb}
that approximately ``30 to 40  percent of my colleagues have potentially seen something that they cannot explain.''
Van Heijst recalls four such sightings. In two instances, a very bright light descended vertically from above at great speed.
One sighting occurred near N\"urnberg, Germany,
while during the other sighting the object vanished into the Mediterranean near the Greek-Albanian border.
He also visually witnessed a bright light in the sky ``stuttering towards the east,''
appearing and disappearing intermittently four times before moving out of sight at an instant speed.

Van Heijst also observed a cylindrical-shaped object of presumably significant size, which remained stationary at a very high altitude over an unknown location presumably south of Spain for about an hour.
He witnessed this event with another experienced pilot, and both were struck by the object's steady shape and position.
After contacting flight control on the ground, Van Heijst was advised to contact military radar control as soon as possible.
The military controller showed a great deal of interest in his sighting and informed him that nothing of any known type, not even a balloon, had been over Spanish airspace.

\section{Malmstrom incident in 1967}
\label{2023-UFO-chapter-History--1953-2016-dicbm1967}
\index{Malmstrom incident 1967}

UFOs may have an affinity for nuclear sites, and Robert Hastings~\cite{Hastings2008Jan}
has collected a large number of UFO sightings in connection with such facilities,
especially nuclear weapons.
One of the most prominent incidents reportedly occurred in 1967 in central Montana at Minuteman Intercontinental Ballistic Missile fields.
of the 341st Space/Missile Wing (341 MW) of the Air Force Global Strike Command at Malmstrom Air Force Base.

According to the reports-partly official and private recollections-twenty
such Minuteman Intercontinental Ballistic Missiles had allegedly been deactivated.
The story has been retold by James Klotz, a UFO researcher and
(at the time) 1st Lt. Robert Salas~\cite{Klotz-Salas-1996,Klotz2005Feb}.


\subsection{Oscar-flight Missile Combat Crew}

In the early hours of the morning of March 16th, 1967, Salas was Deputy Missile Combat Crew Commander (DMCCC) of the Missile Combat Crew
at the Launch Control Center (LCC or capsule) of ``Oscar-Flight,'' which was below ground.
Initially, one of the airmen serving above ground of the Oscar Flight facility,
spotted what appeared to be a star in the sky, zig-zagging across its expanse.
Although he was hesitant to believe it, another light soon appeared, mirroring the previous maneuvers but closer and larger in size.
Wanting another opinion, the Airman called for the Flight Security Controller to join him.
Together, they watched in awe as the lights streaked above them,
suddenly halting or changing direction at lightning accelerations.

At that point, the flight security controller quickly phoned Salas in the underground capsule to relay the unbelievable sighting~\cite{Klotz2005Feb}.
At first, Salas gave little attention to this message.
However, a few minutes after the first call, he received another call from the flight security controller.
This time, there was urgency in the guard's voice. He was clearly frightened and was shouting his words.
``Sir, there's one hovering outside the front gate!~$\ldots$
A UFO! It's just sitting there. ~$\ldots$
I can't really describe it. It's glowing red.''

Salas immediately went over to his commander, 1st Lt. Fred Meiwald, who was on a scheduled sleep period on a small cot behind a curtain.
He woke him and began to brief him about the phone calls and what was going on at the topside of their installation.
(i.e., above their capsule at the surface).

In the middle of this conversation, they both heard the first alarm klaxon sound through the confined space of the capsule,
 and they immediately looked over at the panel of annunciator lights at the Commander's console.
A ``No-Go'' (inoperable) light and two red security lights were lit, indicating problems at one of their missile sites.
Meiwald jumped up to query the system to determine the cause of the problem.
Before he could do so, another alarm went off at another site,
then two more simultaneously.
Within the next few seconds, they had lost eight to ten missiles to a ``No-Go'' condition.

Salas and Meiwald tried to find the cause of the malfunctions and bring the Minutemen missiles back to operational status.
When they had completed their checklist procedure, they discovered that each of the missiles had gone off alert due to a Guidance and Control System fault. Power had not been lost to the sites, the missiles simply were not operational due to some inexplicable reason, each of their Guidance and Control systems had malfunctioned. After reporting to the Command Post, Fred turned to his colleague
and stated that the same kind of missile shutdowns had happened at Echo-Flight!

\subsection{Echo-flight Missile Combat Crew}

Maybe at the same time (there exist some uncertainties about the concurrence), in the morning of March 16th, 1967, another crew, Captain Eric Carlson and 1st Lt. Walt Figel, were the
Missile Combat Crew Commander (MCCC) and the Deputy Missile Combat Crew Commander (DMCCC) of the
Missile Combat Crew
at the Launch Control Center (LCC or capsule) of ``Echo-Flight'' Missile Combat Crew that was also (just as LLC of ``Oscar-Flight'') located below ground
in central Montana.

Maintenance crews and security teams were stationed at two of the Echo-Flight Launch Facilities (LFs)
after performing maintenance work during the previous day.
They had camped out overnight to ensure that the missile sites were secure.
In the early hours of the morning, multiple reports came in from the security patrols
 and maintenance crews stating that they had witnessed unidentified flying objects).
One of the UFOs was spotted directly above one of the Echo-Flight minuteman silos.
The incident was so unsettling that at least one security officer was deeply impacted,
and never returned to missile security duty again.

At approximately 8:30 a.m., Figel, the DMCCC, was providing Carlson, the MCCC, with an update on the flight status when an alarm horn abruptly blared.
It was discovered that one of the Minuteman missiles they oversaw had become inoperable and gone off alert.
 This occurred at one of the two sites where maintenance crews had set up camps.
Feeling disconcerted, Figel believed that the maintenance staff had neglected to inform him, as per protocol,
that maintenance work had been conducted on the missile and that it was now transitioning to an `off-alert' status.

Consequently, Figel promptly reached out to the missile site.
However, he was informed by the on-site security guard that no maintenance had been carried out that morning.
Moreover, the guard claimed to have witnessed a UFO hovering over the site.
Figel initially dismissed the guard's claim as a result of being drunk.

Suddenly, multiple missiles began sounding alerts in quick succession.
In a matter of seconds, the entire flight of ten minuteman ICBMs---one by one across the board---became nonoperational, reporting a ``No-Go'' condition.
Every missile was rendered inoperable, one after the other.

Upon completion of the checklist procedure at each missile site,
it became clear that a fault in the Guidance and Control  System had caused all the missiles to deactivate from alert status.
The sites had not experienced any power outages, but inexplicably,
the guidance and control systems in each missile had malfunctioned, rendering them nonoperational.

Subsequently, the ``Echo-Flight'' Missile Combat Crew---Figel and Carlson---dispatched two Security Alert Teams (SATs, ``strike teams'')
to the sites where maintenance crews were present without informing them that one of the on-site guards had reported a UFO.
After arriving at the Launch Facilities, the SATs reported back to Figel
that all maintenance and security personnel present at each site had witnessed UFOs hovering over the two locations~\cite[p.~12]{Klotz2005Feb}.
(It is unclear whether the two SATs saw these UFOs personally or whether they just relayed hearsay.)

Later that morning, the Echo-Flight Combat Crew was replaced by Captain Don Crawford's crew. Crawford noticed that both Carlson and Figel were still visibly affected by what had happened earlier. During his shift, Crawford witnessed the maintenance crews tirelessly working on the missiles all day and at night to restore them to their alert status. It was a daunting task, as almost twenty missiles had been lost, rendering our deterrent forces vulnerable.
Those missiles had remained out of service for an entire day.



After their debriefing at Malstrom Air Force Base, no explanation was given to the soldiers who had experienced this alleged UFO encounter.
Instead, as Klotz and Salas~\cite[p.~18]{Klotz2005Feb} report, the
official wing history covering the time period January through March 1967 states to the contrary:
\begin{svgraybox}
``Rumors of Unidentified Flying Objects (UFO) around the area of
Echo Flight during the time of the fault were disproven. A Mobile
Strike Team (MST), which had checked all November [[Oscar or Echo]] Flight's LFs on
the morning of 16 March 1967, were questioned and stated that no
unusual activity or sightings were observed.''
\end{svgraybox}

\subsection{Postmortem analysis by Boing Corporation}

Boeing played a crucial role as a support contractor for the Minuteman Missile system,
and at the request of the Air Force, was tasked with conducting an investigation into the incident.
A message from the Strategic Air Command to Boeing Corporation, which built and maintained the Minuteman Intercontinental Ballistic Missiles,
obtained by a FOIA request reads (in capital letters)~\cite{Klotz-Salas-1996}:
\begin{svgraybox}
``All ten missiles in Echo Flight at Malmstrom lost strat alert within ten seconds of each other. This incident occurred at 0845[[?]] on March 1967.''

$\ldots$

``The fact that no apparent reason for the loss of ten missiles can be readily identified is cause for grave concern to this headquarters.''
\end{svgraybox}


Robert Kaminski, an engineer in the Logistics Engineering Group at Boeing responsible for resolving field issues with the Minuteman system,
was designated as the team leader to coordinate the contractor's response when the Echo Flight incident took place.
According to a letter from Kaminski to Klotz and Salas~\cite[p.~21]{Klotz2005Feb}, this investigation
which was of ``grave concern'' to the Strategic Air Command was terminated by them
in a most remarkable way:
\begin{svgraybox}
``Meanwhile, I was contacted by our representative at OOAMA [[Ogden
Air Material Command support facility]] (Don Peterson) and told by
him that the incident was reported as being a UFO event - that a UFO
was seen by some airmen over the LCF at the time E-Flight went down.
Subsequently, we were notified a few days later that a stop work order
was on the way from OOAMA to stop any further effort on this project.
We stopped. We were also told that we were not to submit the final
engineering report. This was most unusual since all of our work
required review by the customer and the submittal of a final
Engineering Report to OOAMA.''
\end{svgraybox}

\section{Minot Air Force Base incident on October 24, 1968}

\label{2023-UFO-chapter-History--1953-2016-mi68}
\index{Minot incident}

An encounter of an alleged UFO with a B-52 strategic bomber of the US Air Force took place on a flight to, and over,
Minot Air Force Base, North Dakota on October 24, 1968. Later, a Minuteman silo became involved as well.

The following summary is based on a very detailed documentation~\cite{minotb52ufo} of this case by the Sign Oral History Project~\cite{SOHPIndex}.
On May 5, 2000 Captain, ret. USAF, Bradford Runyon, Jr. was interviewed by Thomas Tulien and Jan Aldrich~\cite{MinotAFB1968}.

A corresponding investigation was reported in the Project Blue Book entry ``24 Oct 68 24/0030, 24/0530Z Minot AFB, North Dakota''~\cite{MinotAFBPBB1968,bluebook-directory-listing}.

\subsection{Stimulated encounter with an UFO}

Runyon recounts a time when he was a copilot on a flight in a B-52 strategic bomber
that had been out on a mission for 10 hours. As requested permission to land, the controller asked him to check on something.
They directed him in a certain direction, but he was not sure what he was supposed to look for.
Upon requesting more details the controller just said, ``Well, you'll know it if you find it.''

His navigator and radar navigator noticed an object coming toward them at a high speed of approximately
3,000 miles per hour. They were ready for impact, but the object stopped dead off the right wing or off the tail,
being in the right seat.
The object was visible only as a haze. At that point, their radios went out.
Then, it moved to the other side of the airplane and stayed a mile and a half just off the left tail.
It was not identifiable because it did not have an Identification, Friend or Foe (IFF) code.
The skin paint of the object-the radar indication caused by the reflected radar signal from the object-was huge,
possibly six times as large as a tanker plane.

With their radios out, they decided to land at Minot Air Force Base without permission. They started to descend. Finally, only 10 miles from the runway, their radios returned.
At that point, the object must have already separated from their plane, dropped away, and set down.

\subsection{Observation of the landed UFO}

As they proceeded to the base, a general officer came on the radio and instructed them to return and fly over the object.
and observe it and take pictures if possible.
The object had landed, and the radars on the ground and in the place showed its location.
They turned their plane and went back at approximately 150 and 200 mph, keeping a constant distance of approximately 1,000 feet above ground,
and went around the landed object quickly.
Therefore, they did not have a chance to observe it in full detail.
Moreover, their airplane surrounded them, so they had limited views on the side and below.

As they flew around it, Runyon was able to make out that the object was egg-shaped.
The object was aligned with the runway and consisted of three connected parts:
Its first, orange glowing portion resembled molten metal or lava,
but did not appear shiny or radiant. Rather, it had a dull appearance.
Second, there was a shiny tubular section---it was smooth metallic
looking, round tunnel. Attached to it was a third and final part,
a sort of crescent moon-shaped
object that is wrapped around one end of the larger mass, and it
was smooth, shiny, metallic-looking---like a big bumper.
Viewed from above, the entire object was shiny, and it was slightly lit.

As they approached the object, their radios had turned off again. In addition, after they made some distance, the radio went back again.
They finally landed at Minot Air Force Base. All airmen went in for the mission debriefing.


\subsection{Silo slit opened}

During a briefing the day after, a general officer told Runyon that,
while they approach the base and had been still in the air the other day,
an alarm had gone off in one of the missile silos.
Air Police were dispatched to investigate, but the first group of Air Policemen did not check in as expected.

Another group was sent to check on them, and they found the first group unconscious near their vehicle,
with the paint on top of the vehicle burned off.
When they regained consciousness, the first group reported that they had seen an object that appeared to be descending on top of them,
causing them to flee.


The alarms were triggered because the 20-ton concrete lid, which covered the top of the missile silo, had been removed.
Additionally, the chain link fence around the area had been crushed,
 and there were signs of radioactivity present.
The inner alarm within the silo had also been activated.

One could imagine two conventional ways to remove the concrete lid:
by using a large crane to lift it off and set it aside or by activating explosive charges that would blow it off in the event of a missile launch
during war.
However, the explosive charges had not been activated.
The 20-ton concrete lid was discovered just off to the side, in close proximity to the now-open missile silo.

\subsection{Project Blue Book Entry}

The entry of Project Blue Book investigation is entitled ``1968\_-Minot\_AFB\_ND\_Runyon-Brad-2000''~\cite{bluebook-directory-listing}
and carries the information ``24 Oct 68 24/0030, 24/0530Z Minot AFB, North Dakota''
(see also~\cite{MinotAFBPBB1968}).
Project Blue Book entry claims that ``the following conclusions have been reached after a thorough study of the
data submitted to the Foreign Technology Division'':
\begin{svgraybox}
The ground visual sightings appear
to be of the star Sirius and the B-52 which was flying in the area. The B-52 radar contact
and the temporary loss of UHF transmission could be attributed to plasma similar to
ball lightning. The air-visual from the B-52 could be the star Vega which was on the
horizon at the time, or it could be a light on the ground, or possibly a plasma.
\end{svgraybox}
No further investigation by the Foreign Technology Division was contemplated.
Case closed.


\section{John Lennon' New York UFO in August 1974}

\label{2023-UFO-chapter-History--1953-2016-jl1974}
\index{Lennon, John}

John Lennon, the famous musician and member of the Beatles, had a life-long interest in UFOs, even before this sighting.
On the 23rd of Aug. 1974 at 9 o'clock, according to their own unconfirmed story,
Lennon and his muse May Pang saw a UFO from his New York apartment balcony.
He described the object as round, black or gray in the middle with white blinking lights encircling the rim
(bottom circumference)---like light bulbs, on-off-on-off-on-off alternatively---and a red light on top.

The unknown object flew close to his balcony on a summer night without making any noise.
He assured interviewers that it was not a helicopter or a balloon due to its proximity to the rooftop and its maneuvers.
According to his narration, the episode lasted five to ten minutes.

Lennon called the police, who allegedly received similar reports that evening.
The incident took place during Lennon's separation from Yoko Ono and a period of drug and alcohol abuse,
although he stated he was sober at the time of the sighting.

Lennon recalled that the UFO flew down the East River, turned right toward the United Nations, then turned left and down the river.
% https://www.beatlesbible.com/1974/08/23/john-lennon-sees-ufo-new-york-city/

% https://thenewdaily.com.au/entertainment/music/2021/08/23/on-this-day-john-lennon-ufo/

% https://www.facebook.com/watch/?v=436404954002918

% https://www.finebooksmagazine.com/fine-books-news/out-world-collection-john-lennons-drawings-and-sci-fi-magazines

May Pang, who was with him at the time, confirmed most of Lennon's story.
She noted that, after hovering over a building, the UFO turned sideways (turned on its side) and flew down the river.
Then, it shot up into the sky.

\section{Tehran incident in 1976}

\label{2023-UFO-chapter-History--1953-2016-tui1976}
\index{Teheran incident}

On September 18th, 1976, a strange object was spotted in the sky over Tehran, Iran.
The object was seen by multiple witnesses and appeared to change colors and shape.

\subsection{Sighting by airport personnel}

Hussain Perouzi, a night shift supervisor at the air traffic control tower at Mehrabad Airport,
observed the object through binoculars and confirmed its unusual behavior~\cite{Sprague2022Sep}:
The object moved northward, oscillating up and down, at an estimated distance of 5 miles and altitude of 6,000 feet.
It vanished abruptly, only to reappear a mile away within a few seconds.
Seen through binoculars, the object's right and left ends appeared blue in color,
with a flashing red light in the middle.
To confirm that his observation was accurate, he passed the binoculars to one of his trainees who saw the same thing.
Despite the object's stunning brightness, Perouzi was able to discern its shape, which resembled that of a starfish.

The tower radar was under repair, so Perouzi could not confirm the sighting by radar.
However, concerned about the object's potential impact on incoming flights, Perouzi reported it to the Iranian Air Force.


\subsection{First attempt to approach the object: outrun and confusion}

Two Imperial Iranian Air Force F-4 Phantom II jet interceptors were dispatched for investigation.
First, Captain Mohammad Reza Azizkhani flew an F-4 Phantom Fighter jet from Shahrokhi Air Force Base.
The object was so large and brilliant that it could be seen from almost 70 miles away.
Despite increasing his velocity and surpassing Mach 2 speeds, Azizkhani was unable to catch up to the incredibly fast object as it
swiftly made its way to the city limits of Tehran.
Azizkhani was then ordered to return.

However, the object suddenly reappeared directly above the city of Tehran, having traveled a great distance at implausible speeds.
It appeared to have stopped mid-flight above Teheran.

Azizkhani headed toward the object again. When he approached within approximately 25 nautical miles,
his navigational equipment on board began to malfunction, causing confusion and concern.
The equipment returned to normal when he veered away from the object, but every time he got close,
it would shut down again.

Azizkhani returned to base without any clues or explanations for what he had witnessed.
He reported instrumentation malfunctions whenever he was within a certain proximity to the object.


\subsection{Attempt to shoot down an approaching uncategorized object}

Squadron Commander, Parvis Jafari, with First Lieutenant Jalal Damirian serving as the weapons officer,
was then scrambled to intercept the object and took off  in another F-4 Phantom jet.
%They took off to where the object was last tracked on radar.
Here is a short excerpt from a (maybe inaccurate in some details) Defense Intelligence Agency (DIA) report~\cite{DIA76}:
\begin{svgraybox}
C. The size of the radar return was comparable to that of a 707 tanker [[Boeing 707 Stratotanker]].
The visual size of the object was difficult to discern because of its intense brilliance.
The light that it gave off was that of flashing strobe lights arranged in a rectangular pattern and alternating blue, green, red, and orange in color.
The sequence of the lights was so fast that all the colors could be seen at once.
\end{svgraybox}

As Jafari's F-4 Phantom jet came closer, he could faintly discern a diamond-shaped object with lights on all points of the diamond.
He then noticed that four other objects with different shapes separated from the main object at different times.
Whenever the objects were close, his radar jammed, and radio communications were garbled.
One of the objects approached his jet, and he tried to launch a heat seeker AIM-9 missile at it, but the missile panel went out.
The experience was both interesting and terrifying.

Jafari was uncertain about his next move, as he had no way to counteroffend if the object fired a missile at him. The object continued to approach him, and he became increasingly scared as his missile panel was down and communication with the tower was lost. Jafari believed that if the object got any closer, he would have to eject from the plane, but even his ejection equipment was malfunctioning. To avoid a head-on collision, Jafari veered to the left, only to find the object had disappeared. Another detached object began circling his jet, and even when he tried to get away from it, it continued to follow him.
Jafari decided to perform a quick negative G nose dive-an aviation maneuver where the pilot pulls the nose of the aircraft down and accelerates toward the ground. During the dive, the pilot experiences negative g-forces, which push them out of their seat toward the aircraft's ceiling---and
The object shot passed through him and disappeared.
With his instrumentation failing, Jafari finally received a signal to communicate with the tower and was ordered to return to base.

On his descent, Jafari saw that the main diamond-shaped object was still stationary.
As he continued in his turn away from it, the second object that had approached him went to the
returned to the primary object for a perfect rejoin.
This or another object broke off from the diamond-shaped one and made its way toward Jafari.
However, to his relief, the object did not appear to be targeting him as it descended toward the ground.
The object emitted
a luminous light that was so bright that it made the sands visible from as far as fifteen miles away.

As he approached the landing site, Jafari was instructed to conduct a final flyover to gather any valuable observations
of an impact site and to identify the object more closely.
However, as he got closer, his radio malfunctioned, and his instruments went haywire.
Jafari was only able to resume communication with the tower after he had distanced himself from the object.
Despite his attempts to gather more information, he was ordered to return to base.
The main diamond-shaped object in the sky had vanished into thin air while he was landing.

%An initial report of the incident was sent to the US Joint Chiefs of Staff on the same day of the incident.
On the next day, Jafari and Damirian flew by helicopter to the area where they thought the light hit the ground but found nothing.
Occupants of nearby houses only reported hearing a loud noise and a bright flash of light during the night.

\section{Terry Lovelace encounter at Devil's Den in Arkansas in 1977}

\label{2023-UFO-chapter-History--1953-2016-tl1977}
\index{Lovelace, Terry}

The following is a condensed version of Terry Lovelace's memories as recorded in a book~\cite{Lovelace18}
as well as in a podcast~\cite{Iandoli2023Jan}.

%\begin{svgraybox}
%Terry Lovelace and his friend Toby, in 1977 both in their twenties,  went on a ``spontaneous'' camping while they were on active duty as a medic and emergency medical technician. Both worked at Whiteman Air Force Base in Missouri.
%Toby had approached Lovelace in the middle of the night,  suggesting they go camping at a plateau near Devil's Den in Arkansas.
%In hindsight, it is noteworthy that neither of them had prior experience or inclination towards camping.
%
%After driving for about six hours they ended up accidentally finding a plateau (they thought was still within Devil's Den State Park)
%owned by the Bureau of Land Management and setting up camp there.
%They started enjoying the experience, sitting at a campfire after dusk, until they noticed a strange phenomenon: the forest became completely still and silent.
%At some later point they observed lights in the sky that appear to be moving and rotating.
%Lovelace describes the lights as three bright stars that rotate on an axis and move straight up into the air,
%stopping at a certain point and then beginning to move towards them. As the object gets closer,
%they can see that it is a triangle shape with lights on the apex, but they don't hear any noise.
%
%He also mentions feeling a sense of calm and detachment while observing the object, as if their emotions were muted.
%As they watched a light came on from underneath the center of the craft beaming down from about 3000 feet to the ground.
%It was white and about the diameter of a softball on the ground.
%This light hit their campfire for about a minute or less and clicked off.
%Then, instead of this broader light, thin laser-type lights popped on from the same center place of the triangle.
%On the ground these thin lights were about the diameter of a lead pencil.
%One of these lights danced around them and struck them and their belongings multiple times, including Lovelace's chest.
%He did not feel any pain. After the light disappeared, they felt a sense of calm and drowsiness.
%
%They experienced the need to go to sleep immediately, and did so. Lovelace woke up later to see flashing lights outside the tent and assumed they were from a park ranger.
%He also noticed that his combat boots were unlaced and unbuttoned, which he knew he hadn't done before going to sleep.
%Later, after taking off his boots, he  noticed that his socks were out sideways.
%
%When he looked outside the tent, he saw that the triangle craft had descended to 30 feet above the floor of the meadow.
%12 maybe 15 kids seemed to be roaming around, but on closer inspection he realized that
%they were not human: they were small creatures, with spindly arms, large heads and long limbs.
%They were walking around the object in groups of twos and threes, in uneven or irregular ways.
%
%Eventually, a bright milky white column of light came from the center of the object.
%This light was about 30 feet in diameter, about as broad as the craft was high off the floor of the meadow.
%As this happens the gray beings turned their attention toward the light column, walked towards it, and entered it in groups of two or three.
%(Lovelace suggests that the scene is similar to the ``beamer'' from the original Star Trek where characters would ``pixelate'' and vanish.)
%The beings disappeared as well and after the last group enters the light column wereturned off and the [[vertex?]] lights on the triangle shape craft change to solid white before taking off.
%
%In panic Lovelace and Toby rushed to their car, leaving their belongings behind.
%On their return journey home, they became acutely aware of a severe thirst. They stoped to buy drinks which they consumed immediately.
%They decided not to tell anybody because of fear of being considered insane and being dismissed from the Air Force.
%They both experienced a strong and strange sense of alienation and disconnection towards each other.
%
%Back home they suffered from various health problems, such as overexposure to light.
%They were ordered not to communicate with each other, and to avoid the other. Tony became soon transferred to Japan.
%Lovelace's health condition was carefully scrutinized. He was admitted to the hospital.
%During recovery Lovelace was visited by two police officers
%from the  investigative branch of the Air Force's Security Police (OSI) who nudged him into signing some forms of (to him) unknown content.
%
%After some time Lovelace is contacted [[by OSI?]] and brought to an interrogation facility.
%They attempted without success to make him undergo hypnosis in order to discover what occurred during the missing time inside the triangle craft.

% \begin{svgraybox}

In 1977, Terry Lovelace, a medic and emergency medical technician, and his friend Toby, both in their twenties, were on active duty at Whiteman Air Force Base in Missouri. In the middle of the night, Toby suggested they go camping on an impromptu basis. They spontaneously decided to go to a plateau near Devil's Den in Arkansas.
At the time, neither of them had prior experience or inclination toward camping.

After driving for approximately six hours, they accidentally found a plateau owned by the Bureau of Land Management and set up a camp there. As they sat around a campfire, they noticed that the forest had become completely still and silent. They then observed lights in the sky that appeared to be moving and rotating. Lovelace described the lights as three bright stars that rotated on an axis and moved straight up into the air, stopping at a certain point and then moving toward them. As the object got closer, they could see that it was a triangle shape with lights on the apex, but they did not hear any noise.

While observing the object, they felt a sense of calm and detachment, as if their emotions were muted. A light then came on from underneath the center of the craft, beaming down from approximately 3000 feet to the ground and hit their campfire for about a minute or less before clicking off. Instead of this broader light, thin laser-type lights popped on from the same center of the triangle. One of these lights danced around them and struck them and their belongings multiple times, including Lovelace's chest, but he did not feel any pain. After the light disappeared, they felt a sense of calm and drowsiness, and they went into their tent and immediately fell asleep.

When Lovelace woke up later, he saw flashing lights outside the tent and assumed they were from a park ranger. He also noticed that his combat boots were unlaced and unbuttoned, which he knew he hadn't done before going to sleep. Later, after taking off his boots, he noticed that his socks were turned out sideways. When he looked outside the tent, he saw that the triangle craft had descended to 30 feet above the floor of the meadow, and 12-15 small, spindly-armed creatures with large heads and long limbs, who he first assumed to be children but shortly afterwards identified as aliens, were roaming around the object in groups of twos and threes.

Eventually, a bright, milky-white column of light emerged from the center of the object. This light was approximately 30 feet in diameter---roughly the same width as the craft was high---off the ground. As the light appeared, the gray beings turned their attention to it, walked toward it, and entered it in groups of two or three. Lovelace suggests that the scene was similar to the ``beamer'' from the original Star Trek, where characters would ``pixelate'' and vanish. The beings disappeared along with the light, and after the last group entered the light, it turned off, and the lights on the triangle-shaped craft changed to solid white before taking off.

After the craft had disappeared, Lovelace and Toby, still in panic, rushed to their car, leaving their belongings behind. On their journey home, they experienced severe thirst. They decided not to tell anybody for fear of being considered insane and dismissed from the Air Force. They both felt an inexplicable but strong and strange sense of alienation and disconnection toward each other.

After this experience, both suffered from various health problems. They were ordered to avoid each other and not communicate. Eventually, Lovelace was hospitalized and visited by two police officers from the investigative branch of the Air Force's Security Police (OSI), who made him sign unknown forms.

Later, Lovelace was contacted by the Office of Scientific Intelligence (of the CIA).
They took him to an interrogation facility and attempted to use hypnosis to uncover what might have happened during the time he supposedly spent inside the triangular craft.

% \end{svgraybox}

The following is a summary of Terry Lovelace's recollections of the missing time he allegedly spent in the triangle-shaped craft,
as he recalled it through flashbacks~\cite{Lovelace18,Iandoli2023Jan} and later hypnoses.

%\begin{svgraybox}
Lovelace believed to have been brought inside of the triangle-shaped craft. He was paralyzed and unable to move. He stood next to Toby and described the place as being enormous and confusing, as it appeared to be much larger on the inside than the craft he had seen hovering over the campsite. He suggests that the craft could have been in a different location or that the laws of physics were different for the beings controlling the craft. In his angle of sight and while being paralyzed, Lovelace encountered little gray beings. He speculates that these may have been robots running around the craft. He also saw what appeared to be big garage doors and a row of silver disc-shaped flying saucers.

Lovelace also described one other type of alien who was six feet tall, with pinkish flesh-colored skin and sparse hair, large black eyes with no sclera, pupil, or cornea, and wearing a knit garment with shoes that seemed to be integrated into their pant legs. Lovelace asserted that this alien had caused the most frightening experience for him, as he felt like ``the guy was in his head'' and knew everything about him, his life, his plans, and his secrets. The exchange with the alien took place as if the other had ``just raw intellect in the eyes.''

He felt inferior: this alien might have been as superior to humans, as he was to his dog.

Lovelace also recalled seeing a group of humans in unknown uniforms but identical boots as Lovelace, but they took no notice of him and vanished from his view while he was paralyzed.

%\end{svgraybox}

Lovelace also mentioned the accidental discovery of an implant after his legs had been X-rayed following an accident.
He later became contacted by DeLonge\index{DeLonge, Tom} and Elizondo\index{Elizondo, Luis}, who suggested an examination.
Lovelace mentioned that he may have had previous encounters with aliens and that there may be a covert military operation dealing with
``alien abductees'' like himself.
He speculated that the military may be aware of and involved in these types of encounters.

It should be noted that all of these claims and narrations are not confirmed and are based on Lovelace's own speculation and interpretation of his experiences. Lovelace did not come forward with his account of the 1977 incident at Devil's Den until 2018, which is 40 years after the event occurred.


\section{Abduction at a Minuteman Missile launch site in November 1977}

\label{2023-UFO-chapter-History--1953-2016-mw77}
\index{Woods, Mario}
\index{Ellsworth Air Force Base}

The following is an account of events that took place at a Minuteman missile launch facility controlled by the
Ellsworth Air Force Base (AFB) in South Dakota. In November of 1977
USAF Sergeant  Mario Woods served there as a member of the 44th Security Police Squadron.
The narration is based on a podcast with Robert Fleischer~\cite{FleischerWoods2023Mar}.
Just like Minot AFB encountered in Section~\ref{2023-UFO-chapter-History--1953-2016-mi68}
and Malmstrom AFB encountered in Section~\ref{2023-UFO-chapter-History--1953-2016-dicbm1967},
Ellsworth AFB has been part of of a system of cold war bases in the northern part of the US
equipped with a
strategic B-52 bomber wing and  Minuteman Missiles tipped with nuclear warheads
as both aircraft and missile deployment.

Already as a child, Mario Woods had a UFO sighting with his family.
In 1961, his mother was driving him to De Queen Elementary School in Port Arthur, Texas.
They lived close by and the school was a single building that housed kindergarten through 12th grade.
Across the street, a road was blocked off and used as a playground. As they approached the drop-off point,
they were in a line of cars, maybe fifth or eighth. Suddenly, his mother pointed out three silver white, disc-shaped flying saucers
with lights on the bottom, pulsating around them,
hovering in a triangular formation about 50 feet in the air, directly above a nearby church. The saucers were 50 to 70 feet in diameter and had pulsating lights on the bottom. They stayed in the sky for a couple of minutes, during which time everyone on the playground and even people in the school were looking up at them. His sister was scared, but their mother didn't say anything. The saucers suddenly shot straight up and out of the atmosphere, getting smaller until they disappeared from sight.
Woods would never forget that day, and his sister Carolyn and he talk about it often.

Around 2017, Woods and his sister had another encounter with  two Tic-Tac shaped UFO's in Florida,
over the Dade City Florida area, just north of Tampa,  when he stopped going to buy lunch with her. A lot of people saw these UFOs.
The UFOs were there and all of a sudden gone in a flash.

Woods' encounter at Ellsworth AFB in November 1977 started wehn he went outside his Launch Control Facility to smoke a cigarette.
He saw two distant large lights that may have been planets or star, but at first he thought they were helicopters or a B-52 bomber
flying in in tandem. They seem to zero in on him but did not change position.
Woods had the intuition to communicate with whatever was out there.
So he flipped the facility lights using just one switch, which controlled 12 to 14 light cans or perimeter lights on the building and fence line.
The lights were about two feet in diameter and were only moderately bright for the time, as light emitting diodes (LEDs) were not as advanced as they are today.
Woods used them as a signaling device, without following any particular sequence like SOS.
He flipped them on and off three or four times before running back outside.
When he walked outside again about 30 seconds later, the object flipped its lights off and on repeatedly.
It almost looked as if the objects in the sky  reacted to him.

At that point Woods got excited and wanted to repeat this again.
So he went back into his Launch Control Facility and told a colleague Michael  Johnson who was watching TV to come ond and see this reaction.
When they went outside, he first showed Michael this light in the sky.
Then Woods ran inside and switched the facility lights on and off. It did not come back on.
And sure enough the objects would flash back at him: it went off went on went off went on and it went off.
Michael didn't seem very interested and went back inside. The narrator went back outside alone and saw the light had moved closer and further to the north.
He was excited, thinking it might be an airplane flying over. He had told Bill Holloman at the flight security control center,
who was at the Flight Security Control Center, but they didn't find it unusual.
So he flipped the lights at it again and ran back outside, now by himself.
The object came back on, but it was not in the same direction; it was further north and had moved back at a 45-degree angle, indicating that it had gotten closer, but it was still far away. It had moved more to the north of him and stayed lit up for just a few moments before going out again.
He waited, but it did not come back on, despite flipping the light several times.

A while afterwards there was an alarm ``set four'' at one of the Minuteman missile sights called ``November Five.''
Woods and Johnson were ordered to investigate.
Already from afar, as they approached
Woods happened to turn over his right shoulder and looked over toward November Five, which was about seven to eight miles away.
He saw a long-range glow kind of pulsating up off the ground. He couldn't see the ground except he just saw a glow in the atmosphere in that direction.
He didn't know if it was really at November Five but he just knew it was in that direction.

When Woods and Johnson came around the little dog leg on Ormond Road,
they saw a huge object--- of the size of a super Walmart building---sitting there above November Five---exactly on top of that blast door.
The object was not touching the ground but about ten feet in the air.
It was the biggest thing he had ever seen in the air.
The object was a round sphere and appeared to have electricity or flames emitting from the outer area,
like an electrical coil fire of some type.
It had no leading edges, wings, or any sign of thrust, and there was no noise at all.
The object was gargantuan and they pulled up right in front of the site.
Their vehicle was parked at a 45-degree angle about 20 yards from the cattle gate that keeps cattle on the ranchers' properties.

They panicked. Woods kept saying, ``What is it? What is that man? What is that, Michael?'' but Johnson did not answer.
 All of a sudden, before he could think of the next thought, it was as if the atmosphere,
the brightness of this object increased intensely in the cab of the truck.
It was like the sun was directly in their vehicle, and then the air or the atmosphere,
the best way to describe it, was as if it was all pulled out at one time,
and they couldn't breathe. There was no air to breathe.
It was as if he was being pulled up into the back of that seat.
Woods only knew one thing to do, and he didn't even know where the thought came from.
He reached up and rolled that window down, a hand crank window.
He had mittens on the left hand, and he had his Maglite in his right hand.
The only thing that he knew to do was to roll the window down.
He pulled himself  up on that windowsill and sat on it, afraid to touch the ground, afraid to get out of that vehicle.
He held onto the blue  lights on top of the roof, and  flashed his Maglite at this object, five, six, seven, eight times.
As soon as he did that, he felt that as if in his mind he was told that is enough, and he slid back down on the seat. Things eased up,
so they could breathe again, and he told  Johnson it will be okay.
But Johnson never said anything. He just kept looking ahead, sitting there holding on the steering wheel.

As Woods vision began to fade, an object appeared in front of their windshield. It was black, about twice the size of a beach ball, and had lines running from top to bottom with a slight curvature. The object was jumping around erratically, and he saw it three or four times before it disappeared over the top of the vehicle. As he started to pass out, he saw three small beings wearing some type of uniform, standing next to a taller being behind them. They had large eyes and were around four feet tall. He couldn't be sure, but there may have been different types of beings. One of the beings held a straight stick with a glowing tip, which briefly caught his attention. The taller being had something protruding from its chest, which he feared.
Just before he blacked out, he heard and felt a vibration repeating the words ``do not fear.''

When he woke up, he didn't know exactly what that meant but he just felt as if he was really alive. He was in complete darkness, unable to see anything, but he was sitting in the seat of the vehicle. He asked Michael where they were, but there was no answer. As his vision started to return, he looked around and saw some really strange things.
He wondered where they were and how they got there. Without hesitation, he reached down and popped the door.
Woods later notified the base which had already searched for them for five hour.
They triangulated them and found their car tucked on a place close to a dam a couple of miles away.
Both men got debriefed, supposedly by an unknown ``man in black suite,''
Captain Jack Reed and then a young officer by name of Richard Doty~\cite{Pilkington2010Sep}\index{Doty, Richard} who was allegedly
trained in the Strategic Air Command Communications Protocol.

\section{Frederick Valentich disappearance in 1978}

\label{2023-UFO-chapter-History--1953-2016-vd1978}
\index{Valentich, Frederick}

In 1978, Frederick Valentich, an inexperienced Australian pilot and UFO enthusiast, disappeared while flying a small plane, a Cessna 182L with call sign Delta Sierra Juliet, across the Bass Strait from Melbourne to Hobart, Tasmania at dusk on October 21st, 1978. He reported to the Melbourne Flight Service that he had encountered an object with a metallic shine and bright lights and that it was hovering above him. His last words were, ``That strange aircraft is hovering on top of me again$\ldots$ and it's not an aircraft.''

A search for Valentich and his aircraft was conducted but was unsuccessful. During the investigation, it was discovered that some people had seen strange luminous objects in the area on the night of Valentich's disappearance. Additionally, an amateur photographer allegedly took color photos of a disk surrounded by cloud-like vapor near the Strait around the time of Valentich's disappearance~\cite{Isaac-1978.1021_Valentich_disappearance, Sanarov-81, Haines2000}.


%https://www.isaackoi.com/ufo/19781021-valentich-disappearance.html
%https://skepticalinquirer.org/2013/11/the-valentich-disappearance-another-ufo-cold-case-solved/



\section{Rendlesham Forest incident in 1980}

\label{2023-UFO-chapter-History--1953-2016-rf1980}
\index{Rendlesham Forest incident}

This incident involves the alleged visitation of alien craft near or at the twin bases of
Royal Air Force (RAF) [[station]] Bentwaters/Woodbridge and Royal Air Force (RAF) [[station]] Bentwaters.
RAF Bentwaters and RAF Woodbridge were two air bases leased by Britain's Royal Air Force to the US Air Force.
Rendlesham Forest was situated between them, and the bases were approximately 72km apart.
Rumors circulated that one of the biggest stockpiles of US-controlled nuclear weapons in Europe was stored at RAF Bentwaters.
Allegedly, ``nuclear weapons stored in the Weapons Storage Area at RAF Woodbridge were struck by light beams
fired from an unidentified craft seen over the base in the period 25--30 December 1980''~\cite{Hill-Norton1997}.


Unconfirmed accounts leaning toward flying saucers of some type has been presented in several books~\cite{Bruni2000Nov,Pope2014Apr,Heseltine2023Feb}
by Georgina Bruni, Nick Pope, John Burroughs, and James (Jim) W. Penniston and Gary Heseltine.
Pope had worked on the British Ministry of Defence's ``leaky''~\cite{BibEntry2018May}
UFO desk, and the other two authors were USAF security personnel at Bentwaters/Woodbridge involved in the events in Rendlesham Forest.
A Telegraph article~\cite{Hawksley-18}
by Rupert Hawksley is just one in a long series of press reports on the incident.

Ian Ridpath~\cite{Ridpath-Rend} and David Clarke~\cite{Clarke2011Mar}.
drafted a vastly different story:
While it was suggested that at least one RAF radar had tracked the UFOs, the Ministry of Defence
file revealed that no unusual targets were seen on the UK radar picture during the Christmas period in 1980.
Squadron Leader Derek Coumbe confirmed that there was nothing unusual on the radar picture despite claims made to the contrary.
The marks found on the trees in the area were made by foresters' axes, and the decision to remove the trees was made before the sightings.
The ground traces found by USAF airmen were deemed to be of no significance by British police officers.
The levels of radiation detected were confirmed by at least three
independent scientific experts and the makers of the equipment used to be nothing unusual or significant.
The readings were simply background levels of radiation that would be expected in a pine forest~\cite{Clarke2011Mar}.
Finally, the base commander and his deputy of RAF Bentwaters/Woodbridge disagree on what they had seen~\cite{ClarkeBlog2023Jan}.


Let me state up front that debunking stories abound on this topic.
These include a burn-up in the atmosphere of an exceptionally bright meteor~\cite{Ridpath-Rend,Clarke2011Mar}
and the since-dismantled Orfordness Lighthouse~\cite{Rigby2020Dec}.
The UFO sightings may have been triggered by a Russian object that entered the Earth's atmosphere after being in orbit for only a few hours:
Cosmos 749, a satellite launched by the Soviets, re-entered the Earth's atmosphere on Christmas Day and broke into several pieces,
creating a stunning light display in the sky.
Astronomers tracked the object's course across southeast England and noted that the final fragment disappeared in the East Anglia area,
where the Rendlesham sightings were reported.
In addition, 12 satellites decayed during the week of the UFO scare, with six being large objects such as rocket bodies that would have been visible to ground observers. Furthermore, astronomers logged three fireball meteors during the Christmas holiday period,
including one at 2:50 am on Boxing Day morning~\cite{Clarke2011Mar}.

According to another narrative, the ``Historic {B}ritish {UFO} mystery was `prank played on {US} Air Force by
[[the British Special Air Service]] {SAS}'~''~\cite{telegraph-2018-12-30}
in retaliation for the imprisonment and beating of British SAS soldiers by their US captors.
The British SAS had regularly tested US security by probing the perimeters of RAF Woodbridge in Suffolk, where nuclear warheads were said to be stored.
However, these US captors refused to believe who these
SAS soldiers were and repeatedly referred to them as ``unidentified aliens.''


Therefore, with these cautionary provisos, let me proceed with the narrative involving strangeness of the UFO kind.


\subsection{The Halt memo, Halt's signed a notarized statement and Conrad's statement}

Two weeks after the incident,
Lieutenant Colonel Charles I. Halt (of the USAF), the deputy base commander of RAF Woodbridge,
wrote a memorandum~\cite{HaltMemo1981,Ridpath} to the Ministry of Defence  detailing alleged encounters experienced in Rendlesham Forest.
This memorandum, known as the ``Halt memo,'' was made publicly available in the United States in 1983 under the US Freedom of Information Act.
The memo was titled ``Unexplained Lights'' and was dated ``13 Jan 1981.''
However, the two-week delay between the incident and the report may have caused errors in the dates and times given.
The memo was not classified in any way.
\begin{svgraybox}
1.  Early in the morning of 27 Dec 80 (approximately 0300L) two USAF security police patrolmen
saw unusual lights outside the back gate at RAF Woodbridge. Thinking an aircraft might have crashed or been forced down,
they called for permission to go outside the gate to investigate. The on-duty flight chief responded and
allowed three patrolmen to proceed on foot. The individuals reported seeing a strange glowing object in the forest.
The object was described as being metallic in appearance and triangular in shape,
approximately two to three meters across the base and approximately two meters high.
It illuminated the entire forest with a white light.
 The object itself had a pulsing red light on top and a bank(s) of blue lights underneath.
The object was hovering or on legs. As the patrolmen approached the object, it maneuvered through the trees and disappeared.
At this time the animals on a nearby farm went into a frenzy. The object was briefly sighted approximately an hour later near the back gate.



2.  The next day, three depressions 1.5 inches deep and 7 inches in diameter were found where the object had been sighted on the ground. The following night (29 Dec 80) the area was checked for radiation. Beta/gamma readings of 0.1 milliroentgens were recorded with peak readings in the three depressions and near the center of the triangle formed by the depressions. A nearby tree had moderate (0.05-0.07) readings on the side of the tree toward the depressions.



3.  Later in the night a red sun-like light was seen through the trees. It moved about and pulsed. At one point it appeared to throw off glowing particles and then broke into five separate white objects and then disappeared. Immediately thereafter, three star-like objects were noticed in the sky, two objects to the north and one to the south, all of which were about 10 degrees off the horizon. The objects moved rapidly in sharp, angular movements and displayed red, green and blue lights. The objects to the north appeared to be elliptical through an 8-12 power lens. They then turned to full circles. The objects to the north remained in the sky for an hour or more. The object to the south was visible for two or three hours and beamed down a stream of light from time to time. Numerous individuals, including the undersigned, witnessed the activities in paragraphs 2 and 3.



(Signed)

Charles I. Halt, Lt Col, USAF

Deputy Base Commander
\end{svgraybox}

David Clarke investigated the background of the memo, specifically the reaction of Col Theodore J. Conrad,
who served as the Bentwaters/Woodbridge base commander, while Halt was his deputy.
Conrad denied many of the claims made by Halt in both the 1981 memo and a notarized statement from Halt in 2010. In a statement excerpted on a blog by David Clarke, Conrad judged Halt's memo as having generated
``the frenzy of speculative reporting and the inevitable allegations of cover-up.''



\subsection{December 26, 1980}

At 03:00, in the very early morning hours of December 26,
1980---confusingly reported as December 27, 1980 by Lieutenant Colonel Charles Halt in his memo to the UK Ministry of Defence---a
security patrol stationed near the east gate of RAF
Woodbridge, including
Airman First Class John Burroughs, spotted strange lights in the forest, flashing red and blue,
that appeared to be descending into the nearby Rendlesham Forest.

Burroughs
and his supervisor, who also saw the lights, made a small excursion off base in their car.
At first, they believed it to be a downed aircraft.
Down a small track, white light became visible, in addition to red and blue lights.
This white light was particularly odd and at one point appeared to be coming closer to them, down the small track.
The color, configuration and movement
of the lights were like no aircraft or vehicle they were familiar with.
They decided to go back to the gate and inform the
on-duty flight chief, Staff Sergeant James W. Penniston.
Soon after, Penniston and a few others arrived on the scene.

The control tower checked the radar and contacted other locations for information. They learned that a ``boogie'' (for unknown air contact)
had been tracked 15 minutes earlier but had disappeared over the Woodbridge base.

Burroughs and Penniston proceeded on food to the source of light that they finally arrived at. Their radio equipment failed and
they were exerting static electricity in the air around them.
Suddenly, as they approached, there was a silent explosion of light.
At one point, Penniston and Burroughs threw themselves to the ground.
Penniston saw no immediate harm from the intense light. Therefore, he stood up, looked to his right and saw Burroughs
engulfed in a huge beam of light that seemed to be coming from above, completely encompassing him.

The initial sphere of light that was in front of him had disappeared and transformed into the shape of some sort of craft.
It had a roughly triangular shape and appeared to be hovering just above the ground or resting on tripod-like legs at each edge.
The craft had a bank of blue lights on its side and a bright white light on top. No sound could be heard.
Penniston noticed strange symbols on the side of the craft that resembled ancient Egyptian hieroglyphs. He took photographs and sketched the craft and symbols in his notebook. When Penniston touched the craft, he described the hull as hard and smooth, like smooth opaque black glass. However, the symbols felt rough, like running his fingers over sandpaper. As he touched the symbols, the white light on top of the craft flared up and became so intense that he was temporarily blinded. When he removed his hand, the light dimmed, and the sense of panic receded. After some time, the craft lifted off the ground and slowly rose above the trees surrounding the clearing. There was no noise during this time.
The craft seemed to maneuver through the trees as it cleared them and then accelerated away in an instant~\cite[Chapter~1]{Pope2014Apr}.


Shortly afterwards, at 04:00, local police were called to the scene but could only see lights from the Orfordness lighthouse
several miles away on the coast.

When the servicemen returned to the area at daybreak,
they identified three small impressions in a triangular pattern on the ground,
along with burn marks and broken branches on nearby trees.
The police were again called out at 10:30 to view the impressions, which they thought could have been made by an animal.


As stated by Penniston~\cite[Chapter~19]{Pope2014Apr},
\begin{svgraybox}
``after what I had seen in the early morning hours on that night, I left that forest a different man than when I walked in. I was in awe of the technology and, yes, a knowing that it was not an aircraft that could have been manufactured in 1980 or even now.
The replication of this craft in my assessment is far more advanced than any we could develop in the next few centuries at the very least.''
\end{svgraybox}


\subsection{UFO scanning a weapons storage area in late December 1980}

Sergeant Steve Longero of the USAF Security Police
allegedly saw an object hovering over the treetops outside the perimeter of the Bentwaters (nuclear) Weapon Storage Area~\cite{Heseltine2023Feb,FleischerHeseltine2023April},
with  different colored lights on it,
disc-shaped, and with a greenish glow. He estimated that it was about 200 to 300 yards away, at an approximate altitude of perhaps 40 to 50 feet,
 and was located on the forest side of the nuclear Weapon Storage Area.
He described the object as looking like an eye with something dripping off it, and he observed a searchlight-like beam coming from the object
and shining onto the ground within the Weapon Storage Area.
The beam was continually moving and shining across the entire area in a grid search pattern,
scanning the entire row of bunkers with nuclear weapons, casting left to right.
The object then moved away over the forest out of view, and Longero began to hear a lot of radio chatter.


\subsection{A reddish/orange object on late December 27, 1980}

According to a recent reinvestigation of the Rendlesham Forest Incident by a formal criminal investigator Gary Heseltine~\cite{Heseltine2023Feb},
Lt. Bonnie Tamplin's traumatic incident during late December 1980 has long been established, but only fragments of the incident were previously known.
According to the generally accepted story before the Heseltine's reinvestigation,
Two USAF Security police officers, Airwoman Lori Buoen and Airman John Trementozzi, saw a red ball of fire descend into the forest while at the East Gate.
Lt Tamplin, who happened to be on mobile patrol in the area,
agreed to investigate and soon after, Buoen and Trementozzi heard Tamplin's distressed screaming voice over the radio requesting help.
Master Sergeant ``Bobby'' Ball was believed to have responded to her cries over the radio first and reached her first.
Tamplin was so traumatized that she was relieved from duty and seemingly did not return to shift duty.

However, during the filming of Capel Green, former Airman First Class Michael Stacy Smith came forward with direct witness testimony to the events involving Lt Tamplin.
Although he cannot tell the whole story as a direct witness,
a trusted supervisor on shift had told him the rest of the story.
According to Smith, Lt Tamplin had visited him on the night of the incident and told him she was going to Woodbridge.
A few minutes later, he heard her voice over the radio, but it wasn't clear.
He saw several vehicles leaving Central Security Control in a hurry and was told to go to Butley Gate.

A short time later, MSgt Ball returned in his vehicle with Lt Tamplin, who was crying, had dirty clothes, and was not wearing her beret.
Smith learned from a supervisor after the shift that an orange/reddish ball had come in front of Tamplin's jeep as she was driving.
Apparently, in trying to get away from the object, the jeep ran off the road and rolled onto its side. Lt.
Tamplin was able to crawl out of it but the reddish/orange object was still close to her location, so she had opened fire on it.
After this incident, when her weapon was checked into the armory after the incident, a number of rounds were found to be missing.
The missing rounds were replaced after the weapon was cleaned and returned to the armory.
Heseltine mentioned that so far had Tamplin refused to come forward and tell her story~\cite{FleischerHeseltine2023April}.



\subsection{December 28, 1980}

In the early hours of 28 December 1980, Lieutenant Colonel Charles Halt,
accompanied by several servicemen visited the site of the alleged Rendlesham Forest UFO sightings.
They took radiation readings using a military survey meter and allegedly detected a small increase in radiation levels.
Halt recorded the events on a microcassette recorder and reported seeing a flashing light across the field to the east,
as well as three star-like lights in the sky that seemed
to hover for two to three hours and beam down a stream of light from time to time.
The transcript of the Halt tape includes the following:
\begin{svgraybox}
``It looks like an eye winking at you. Still moving from side to side and when you put the starscope on it, it sort of like has a hollow center,
a dark centre, it's~$\ldots$ like a pupil of an eye looking at you, winking.
And the flash is so bright to the starscope, that it almost burns your eye~$\ldots$''

``04.00 hours. One object still hovering over the Woodbridge base at about five to ten degrees off the horizon,
still moving erratic and similar lights and beaming down as earlier.''
\end{svgraybox}

On the following night, on December 28, 1980, a group of military personnel, including Larry Warren,
an 18-year-old soldier, Sgt Adrian Bustinza, and other soldiers, were sent out on patrol in Rendlesham Forest,
investigating the reports of strange lights seen in the area on the previous two nights.
Sometime after 11 pm, the men left their trucks and walked toward the field where the lights had been spotted.
According to Warren and others, they had a close encounter with an unidentified flying object and other
strange experiences. Warren claims to have witnessed ``disaster preparedness officers''
with Geiger counters moving in a semicircular pattern around a small object on the ground.


In Larry Warren's very early account of the Rendlesham Forest incident, he describes being taken to a field known as the ``farmer's field'' with a group of others.
In the field, they saw approximately 15 to 20 US Airforce security officers standing in a circle around a yellowish-green fog that
was hugging the ground and glowing. Suddenly, a small light, the size of a beach ball, self-illuminated, and not quite red in color,
appeared and stopped over the mist. There was then a bright explosion of light that left a shimmering,
translucent craft on the ground where the mist had been.
Warren claims that the incident was being filmed by two people from a distance using early video and cine film cameras.
Warren says that he witnessed something akin to a bubble emerging from the craft and dividing into three bubbles, each containing the upper torso of a humanoid figure.
The three entities floated beside the craft, and Warren claims that he saw Colonel Williams have a silent face-off with them.
 Eventually, Warren was told to leave and return to the staging area,
where vehicles were parked and could go no further into the forest~\cite{Heseltine2023Feb,FleischerHeseltine2023April}.




The Condign Report~\cite{Clarke2022Nov} mentions the Rendlesham Forest incident
in Volume 2, Working Paper 1, Annex F under the heading
``Non-ionizing EM effects on humans'' and ``Potential mental Effects on humans'' the report states~\cite{CondignReportVol2}:
\begin{svgraybox}
``Neurological, rather than biological
effects may be the clue to some human
behaviour after exposure  in the near field of
UAP radiation.  Particularly sensitive are the
temporal lobe areas of the brain. Unlike other
parts of the brain, the temporal cortex can be
rendered electrically unstable.''

$\ldots$

``The well-reported Rendlsham
Forest/Bentwaters event is an example where it
might be postulated that several observers were
probably exposed to UAP radiation for longer
than normal UAP sighting periods. There may
be other cases which remain unreported. It is
clear that the recipients of these effects are not
aware that their behaviour/perception of what
they are observing is being modified.''
\end{svgraybox}


Both Penniston and Burroughs have reported suffering from Post-Traumatic Stress Disorder (PTSD) following the incident.
On the basis of the aforementioned passage in the Condign Report, Burroughs later won compensation for illnesses he blames on exposure to radiation from the event~\cite{Potter2022Sep}.










\section{Soviet UFO missile crises in 1982}

\label{2023-UFO-chapter-History--1953-2016-sumc}
\index{Soviet UFO missile crises}

The Soviet UFO missile crisis was the opposite of what US Missile Combat Crews experienced in the Malmstrom incident in 1967
(see Section~\ref{2023-UFO-chapter-History--1953-2016-dicbm1967}): instead of deactivating US intercontinental ballistic missiles,
UFOs activated USSR intercontinental ballistic missiles without authorization from Moscow.


The incident allegedly took place on October 4, 1982, in Byelokoroviche, Ukraine.
It affected the 50th missile division of the Strategic Missile Forces of the Carpathian Military District.

On October 4, 1982, at 18:30 Moscow time, several unknown aircraft appeared in the sky
moving along trajectories that were impossible for any known earth-sourced technology.
According to military reports, strange lights were seen in the night sky.
One or more UFOs were spotted hovering above the missile base in Usovo, Ukraine.

Alarming events occurred inside the underground bunker of military unit 52035,
where the missile launch control panels were located.
The signal lights on both control panels suddenly lit up,
which meant that the missiles were ready to be launched.
This could have only occurred after receiving appropriate authorization from Moscow,
but Moscow did not send such a signal.

The situation was confirmed by a report signed by Major M. D. Kataman,
assistant commander of military unit 52035 for communications.
He did not see the UFOs flying above because he was, at the time, on shift in the underground bunker (capsule).
However, ``on October 4, 1982 at 21:37,'' he writes, ``I observed the spontaneous switching on of all signal lights on the panels.''
Someone or something was apparently manipulating the series of precise control codes, four spaces,
and control code combination, which regulate the computerized missile control launch panel~\cite{Huneeus2011Jan}.
Within fifteen terrifying seconds, the base lost control of its deadly weapons.

One of the eyewitnesses to the appearance of a UFO over the base was Lieutenant Colonel Vladimir Platunov.
He described the object as ``just as a flying saucer'' without hatches or portholes,
300 meters in diameter, and with a ``flawless geometric shape.''
The saucer made a sharp right-angle turn and made no sound.
The object was also seen by servicemen from a neighboring unit and many civilians from nearby villages.
They all reported that the ``flying saucer'' was huge and had a ``flawless geometric shape.''

Subsequently, the Soviet Academy of Sciences determined that the UFO hovering near the missile site was,
in reality, only a military flare.

The incident was later reported by ABC TV News Moscow correspondent David Ensor~\cite{ABC-transcript-1994,Huneeus2011Jan}
in its documentary ``Soviet KGB Files on UFOs,'' which was broadcast on October 6, 1994.
The story was retold in the TV show entitled
``Military Secret: UFOs and the Ministry of Defense of the Russian Federation''~\cite{Hvar2012,TT-SU}
and reviewed by Hastings~\cite[Chapter 23]{Hastings2008Jan}.

\section{Brazil's night of UFOs on May 19, 1986}
\index{Brazil's night of UFOs}
\label{2023-UFO-chapter-History--1953-2016-bui1986}

By listening to the reports of Brazil's ``night of UFOs,''
I am reminded of the Washington Flyovers in 1952, described in Section~\ref{2023-UFO-part-History-chapter-post-1945-pre-1953-WF52}
(page~\pageref{2023-UFO-part-History-chapter-post-1945-pre-1953-WF52}).
As in the former flap, these phenomena could have been ``debunked'' as natural atmospheric events
such as inversion layers~\cite{Menzel_1953},
bright stars or meteors or optical illusions,
but just as then, elements of uneasiness remain.

On the night of May 19, 1986, a series of unidentified flying objects were tracked by ground radar and reported by pilots over several states in southeastern Brazil.

An air force air traffic controller named Sergeant Sergio Mota da Silva reported seeing unusual lights in the sky while working in the control tower of an airport near S\~ao Paulo~\cite{Tahir2022Jun}. He stated that the lights did not move up or down, left or right. He asked other controllers if any planes were headed to his airfield, to which they replied there were not. At one point, he said he dimmed the lights on the airport's runway, causing the objects to come closer to the control tower and then move away when he turned up the brightness.

At approximately 8 PM, approximately 2000 cadets and officers from an air force training school near the S\~ao Paulo airport reported seeing
lights in the sky, similar to those that were previously reported by da Silva.
An hour later, a plane piloted by the president of aircraft maker Embraer reported several objects in the sky as it was landing at the airport.
These lights were also seen by two passenger planes.

Radar returns were picked up by airports in S\~ao Paulo~\cite{Brazil86} and were later tracked by the Integrated Air Defense {\&} Air Traffic Control Center I (CINDACTA I) in Brasilia. The objects were described as spherical in shape and self-luminous white, green, or red in color. The radar controller and some pilots described the objects as performing unusual and erratic maneuvers.

The incident was considered significant because it was the first time that government authorities of any country published the existence of unidentified aerial phenomena in their airspace. CINDACTA I radars that covered the whole Southeast Area of the country, as well as the Federal Capital, captured as many as twenty-one unidentified radar plots, many of which demonstrated very uncommon flight profiles. The unidentified objects were tracked by radar as they moved across the airspace, but no physical contact was made.

Brazil's night of UFOs received significant media attention and remains one of the most well-documented cases of a military encounter with unidentified flying objects. Despite the efforts made by the Brazilian government and the Brazilian Air Force, the objects remained unidentified. To date, there is no official explanation for these events.

Here are some personal anecdotal accounts in more detail. As the radar returns were considered unidentified, the Brazilian Air Force scrambled five fighter jets to investigate. Two Northrop F-5E, flown by the pilot of the alert jet Lieutenant Aviator Kleber Marinho and Captain Aviator M\'arcio Brisola Jord\~ao from Santa Cruz Air Base in Rio de Janeiro state, and two Dassault F-103 Mirage and one Dassault Mirage 500 jet fighters, flown by Captain Aviator Rodolfo da Silva Souza, Captain Aviator Armindo Souza Viriato de Freitas and Captain Aviator J\'ulio C\'ezar Rozenberg,
respectively, from An\'apolis Air Base in Goias state were launched.
The pilots were instructed to intercept the objects without taking an aggressive stance.

Pilots of the scrambled fighters reported visual and radar sightings of the objects. Kleber, who had activated the afterburner of his F-5E, recalls~\cite{Brazil86}:
\begin{svgraybox}
``I began to descend, going directly toward the
object, but being on the whole careful and considering the possibility of an optical illusion,
due to the darkness. I could be seeing a light under water, a large ship with a spotlight or any
other thing. For this reason I didn't want to rely only on my visual impressions so (I) tried my
radar, even without instruction to do so. And, really, at about 8 to 12 miles away, an object
appeared on the screen, confirming the presence of something solid in front of me. This
coincided with the direction of the light that I had sighted visually. In the radar systems that
equipped the fighters of that time, the size of the plot is in general agreement with the size of
the contact. The radar indicated an object of about one centimeter (screen image size), that
means something in the span of a Jumbo (Boeing 747).

$\ldots$

I arrived close to the object, positioning me about six miles away from him; this is still
far enough away so that I can accomplish a necessary verification at night. The object
stopped moving in my direction and it began to climb. I didn't lose my initial radar contact
and I started to climb with [[it]]. I continued following the contact up to about 30 thousand
feet, when I lost the radar contact; I could still see it. But then the strong light became merged
with the lights of the stars.''
\end{svgraybox}

All other aviators failed to intercept and returned to their bases.
In his attempts to locate the ``object'' Rozenberg recalled that he did not have anything on his radar and did not see anything,
although ground control
announced that the alleged object was a mile ahead.
This night, according to the consensual memories of the pilots, was exceptionally clear and bright.

Geraldo Souza Pinto, the pilot of a Boeing-707 cargo plane, was subsequently contacted by CINDACTA~I [[ground control?]].
They [[ground control?]] told him, ``we have a UFO at that location [[at their 11 o'clock position]],'' appearing on their radar.
Pinto then recalls
``It was at this position that I saw it. A very strong light shone, a white light.''
Next, the crew lowered their cockpit lights to be able to search for the contact visually.
Nivaldo Barbosa, the copilot, saw something outside the cockpit that left a luminous trace.
Pinto recalls,
\begin{svgraybox}
``$\ldots$
the controller informed us that the object had suddenly moved to our right at a very
high rate of speed, reaching in a fraction of a second, an incredible speed, something above
Mach~5. A human being could not tolerate an acceleration of this magnitude. He would die
with such displacement.

$\ldots$

We were some 30 miles from the object. The impression that we had was that
the contact was (now) moving at a low speed, and that we were approaching [[it]].
The approach continued. The radar continued to inform us of the (separation) distances:
fifteen miles, ten, five~$\ldots$
But I looked, [[Barbosa]] looked and didn't see anything else. Then
the controller said: `Three miles, two, ah~$\ldots$ the traffic is merging with the plot of your
(own) airplane.' We looked upward, down and we didn't see anything! We were told by
ground control that the object was passing behind our aircraft, but his transmission began to
have a lot of interference and the radar lost him of view.''
\end{svgraybox}


A short unconfirmed message or comment from the USDAO (US Defense Attach\'e Office) in Rio de Janeiro to the Defense Intelligence Agency (DIA) was entitled~\cite{Brazil86} ``BAF [Brazilian Air Force] has a Close Encounter of the First Kind.'' It states: ``Three visual sightings and positive radar contact from three different types of radar systems lead one to believe that something arrived over Brazil the night of 19 May. Pilots or ground personnel reports no EMI [[electromagnetic interference?]] effects due to the presence of UFO/UAP.


\section{UFOs over Gulf Breeze from November 11, 1987, to 1988}
\index{Gulf Breeze UFO incident}
\label{2023-UFO-chapter-History--1953-2016-gbui}

The Gulf Breeze UFO incidents might have been good examples of how the UFO community sometimes gets steered, nudged, and spun by occasional hoaxers and persons and institutions who subsequently rock the boat of attention~\cite{franck-attention-1999,franck-attention-1993,Davenport2002Sep}.

On November 11, 1987, the first alleged UFO sighting in this persistent ``UFO wave'' was made by Ed Walters~\cite{Walters1991Jan}, a local contractor who, according to an article in the Pensacola News Journal, had, at the time, been ``convicted in 1967 of forgery charges from Duval County, Fla, and two auto theft charges from Alachua County, Fla. He had served $1\frac{1}{2}$ years in prison~$\ldots$''~\cite{PensacolaNewsJournal1990}. Walters supplied the paper The Gulf Breeze Sentinel with Polaroid photos that he claimed to show a UFO.
He also claimed that he had been briefly immobilized by a blue beam and had taken photos of a UFO hovering above his Gulf Breeze home. He alleged that he had videotaped and taken 32 photographs of the UFO and had seen aliens that communicated with him telepathically in both English and Spanish. He claimed to have been lifted off the ground by a bright blue beam of light and that his wife had attempted to outrun the same beam.

Over time, Walters and his family reported 19 sightings or encounters and claimed that humming in his head signaled the arrival of UFOs. On one occasion, Walters said he felt a presence, saw the UFO, took a photo of it, and then lost consciousness for an hour.

In 1990, after the family had moved out of the house at 612 Silverthorne Road in Gulf Breeze in December 1988, a new resident found a UFO model made up of ``four plastic foam plates and some drafting paper'' in the attic~\cite{PensacolaNewsJournal1990MF}. The model found in Walters' house closely resembled the UFOs in his photos. Walters denied knowing about the model but refused to take a lie detector test. Instead, he signed a statement denying any knowledge of the UFO model. The person who found the model also signed a statement claiming to not know who made it. Photographers were able to recreate photos of UFOs from Walters' book using the model.

Already in June 1988, Robert Boyd, a critic, had been forced out of the Mutual UFO Network (MUFON) because of his conclusions that Walter's photos were probably fakes. MUFON, as well as Bruce Maccabee~\cite{Walters1991Jan} and other photo analysts, still supported Walters' photos.

In early 1990, MUFON hosted its annual UFO symposium in Pensacola, attracting hundreds of attendees. The event garnered free publicity and brought in tourists, with UFO-focused tours being organized and merchants selling UFO-related merchandise. A local travel agent arranged 600 flights for the conference. The success of the symposium led to the first Gulf Breeze UFO Conference in October 1993, featuring alien abduction researcher John E. Mack, which was expected to attract 600 attendees. The second conference in October 1994 drew over 700 attendees, many of whom suspected they had been abducted by aliens. Additionally, for Walters, the Gulf Breeze UFO incident was profitable: according to an article in the Pensacola News Journal~\cite{PensacolaNewsJournal1990}, he had been given an advance of {\$}200,000 for his book~\cite{Walters1991Jan} and was offered {\$}450,000 by a TV station for the mini-series rights on his UFO encounters.

Later, the Pensacola Bay Bridge was a popular spot for UFO watchers in 1991, with many reporting seeing a ``pearl necklace''
of lights in the sky and a red glow over 74 times. People frequently gathered near the bridge at night with equipment to capture footage of the sightings. In 1996, the mayor of Gulf Breeze suggested that the lights in the sky were likely from nearby military airfields and airports. The MUFON UFO hotline was disconnected in 2000, and the president of the organization stated that there was not much going on in the area. The Center for Inquiry executive director agreed with colleague Phillip Klass's statement that the sightings were a hoax and suggested that people let their imaginations get carried away. Just because something cannot be identified does not mean it is from another planet.

