%%%%%%%%%%%%%%%%%%%%% chapter.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% sample chapter
%
% Use this file as a template for your own input.
%
%%%%%%%%%%%%%%%%%%%%%%%% Springer-Verlag %%%%%%%%%%%%%%%%%%%%%%%%%%
%\motto{Use the template \emph{chapter.tex} to style the various elements of your chapter content.}
\chapter{Executive summaries}
\label{2023-UFO-part-Speculation-executive-summary} % Always give a unique label
% use \chaptermark{}
% to alter or adjust the chapter heading in the running head



\abstract*{
As I do not know what actually happened, I will present three scenarios:
(i) One in which no alien intrusion occurred but waves of mass hysteria created mistaken UFO sightings, with governments everywhere trying to calm their populations, dealing with an avalanche of reports, and a press that sells sensational cases.
(ii) Another scenario involves alien visitations and even crashes of alien craft, but those aliens prefer not to be disclosed. Governments try to maintain the illusion of autonomy while attempting to take advantage of any salvaged or gifted alien craft.
(iii) The last scenario involves the Others dictating strict nonfraternization while pursuing their interests, particularly breeding and reproduction, by abducting a sizable number of humans.}


\abstract{
As I do not know what actually happened, I will present three scenarios:
(i) One in which no alien intrusion occurred but waves of mass hysteria created mistaken UFO sightings, with governments everywhere trying to calm their populations, dealing with an avalanche of reports, and a press that sells sensational cases.
(ii) Another scenario involves alien visitations and even crashes of alien craft, but those aliens prefer not to be disclosed. Governments try to maintain the illusion of autonomy while attempting to take advantage of any salvaged or gifted alien craft.
(iii) The last scenario involves the Others dictating strict nonfraternization while pursuing their interests, particularly breeding and reproduction, by abducting a sizable number of humans.}





\section{The ``Great Nothingburger'' scenario}
\label{2023-UFO-part-Speculation-executive-summary-nb} % Always give a unique label


In this scenario, UFO sightings lack any ontological basis. They are simply delusions and projections of the mind.

As Steven J. Dick pointed out~\cite[Chapter~3]{Dick2015}, throughout history, there have been instances where people thought they had found evidence of life beyond Earth, including the 1835 Moon Hoax/Satire, the Mars canal controversy (1894--1909), the Orson Wells War of the Worlds broadcast in 1938, the discovery of pulsars in 1967, the Viking landing on Mars in 1976, and the 1996 claim of Martian nanofossils.

The following episode from 1835 is an excerpt of a recollection by Dick, as quoted earlier. That year, the New York Sun published a series of six illustrated articles claiming that Sir John Herschel had made great astronomical discoveries while observing the moon at the Cape of Good Hope. The articles were later revealed to be a hoax written by Richard Adams Locke, a 34-year-old reporter who had just joined the Sun.

According to the articles, Herschel had observed large, winged creatures on the moon that were approximately four feet tall, with faces ``of a yellowish flesh color'' and appearing to be in conversation: ``these creatures were evidently engaged in conversation: their gesticulation, more particularly the varied action of their hands and arms, appeared impassioned and emphatic. We hence inferred that they were rational beings.''

Despite the fact that the journal the articles claimed to be from had ceased to exist three years prior, the hoax was widely read and generated significant interest at the time.

The author, Richard Adams Locke, was soon revealed to be a reporter for the Sun, and the articles were exposed as a hoax. However, it was not until 150 years later that historian Michael Crowe argued that Locke was actually writing satire aimed at advocates of inhabited worlds such as the Scottish astronomer Thomas Dick, who had calculated the number of inhabitants on each planet in the Solar System to be over 21 trillion. The Moon Hoax, according to Crowe, was a way for Locke to ridicule such absurd conclusions.

The public and scientific reaction to the Moon Hoax articles was significant. The story captured the attention of a large audience and had a major impact on the newspaper's success. At the time, the New York Sun had a circulation of approximately 8000, but during the Moon episode, it reached 19\,000 and continued to remain high. The story was so popular that the Sun sold 60\,000 copies of the series in brochure form and released lithographs of the supposed lunar creatures. The story gained international attention, with French, Italian, Spanish, and German editions of the brochure appearing and numerous other newspapers reporting on it. The Moon Hoax may not have been true, but it had great popular appeal and was highly profitable for the Sun, much as sensational tabloid journalism continues to be today.

From a debunking point of view, Arnold's sighting in 1947 can be perceived as a series of misperceptions, as pointed out by Hynek's Project Blue Book report \cite{1947-KA-PBB}. The length of the objects was one of the misperceptions. It was probably ``pelicans flying in V-formation'' \cite{Maccabee2017Jun}. The incidents in Roswell were initially misinterpreted as alien spacecraft due to the sighting of a downed Project Mogul balloon \cite{Weaver1995Jan}. People everywhere gazed at the sky and imagined extraterrestrial spaceships in the form of flying saucers. Even the Air Force's Project Sign was being increasingly consumed by the flying saucer craze, and the Washington Flyovers, caused by atmospheric inversions \cite{Menzel_1953}, generated massive attention at the heart of the US empire.

At that time, the UFO craze in the USA became so intense that the CIA~\cite{Haines-CIA-UFO} feared it could undermine established authority
and disrupt the ``orderly functioning'' of the government.
They were concerned that UFO reports might clog communication channels with irrelevant information and cause ``hysterical mass behavior.'' Potential adversaries---the Soviets---might take advantage of the UFO phenomenon
and use it to confuse US air defenses during an attack.

Several high-ranking experts---individuals like Menzel and Hynek, researchers at Battelle Memorial Institute,
renowned scientists in selected committees such as the Robertson Panel and the Condon Committee---were asked to dampen the saucer frenzy, and
convince the public that there was no national security threat posed by flying saucers.
They were also asked to suggest, if not recommend, that any official investigations into the increasing number of
alleged sightings reported to the Air Force should be terminated.
With very few exceptions, due to a lack of data, all previously investigated cases had turned out to have prosaic and mundane explanations.
These saucer investigations would never yield anything valuable.

In particular, two groups of experts, the Robertson Panel and the Condon Committee,
were tasked with calming the hysteria and convincing the public that there was no threat to national security posed by UFOs.
The Condon report from the latter group also recommended that although a tiny fraction of ``hard cases'' remained unexplained and ``strange,'' the opportunity costs of ongoing reporting were unjustified. Therefore, as intended and planned, the recommendation was made for the US Air Force to cease any official investigations into the constantly growing number of reports.

Despite occasional efforts from external sources, such as the Rockefeller Initiative~\cite{Berliner2000Jun} or the Disclosure Project by Greer,
the US Air Force~\cite{Greer-dp} has been able to maintain a peaceful state of non-involvement until the present day.

A congressional initiative, initiated by a UFO enthusiast and real estate tycoon with strong connections to a powerful senator,
led to the creation of a small private contract and an associated section in the Pentagon.
Despite not gaining Special Access Program status, this project aggressively attempted
to uncover what its protagonists believed to be secret government programs related to UFOs,
like a screwdriver trying to open a corked bottle.
To stay with this allegory, the bottle they were trying to open, which they believed to contain UFOs, was empty.

Congress became suspicious that there might be more to the UFO (now called UAP) issue and created legislation for people to come forward with historic and current carve-out unacknowledged secret information. They also established a permanent ``All-domain Anomaly Resolution Office (AARO),'' a kind of follow-up to Project Blue Book, to thoroughly investigate the contents of the (empty) UFO bottle and report directly to the Under Secretary of Defense for Intelligence and Security~\cite{AARO22}.

In addition, the national security apparatus, including the CIA, this
Mighty Wurlitzer~\cite{Loory1974,Bernstein1977,Wilford2009,Covert2016}, has been
using the chimeric UFO phenomenon as a camouflage~\cite{SnowdenTweet2023Feb}
to their own advantage:
to disguise all kinds of secret operations, nuts-and-bolts craft, and distractions from inconvenient exposures~\cite{HershsubstackFeb3,Hersh2023Feb}.


\section{The hidden breakaway scenario}
\label{2023-UFO-part-Speculation-executive-summary-hb} % Always give a unique label

This scenario basically follows Oliver Harry Turner's\index{Turner, Oliver Harry} recollections~\cite{TurnerAustralia1971}.

Arnold may have been the first popular witness to the UFO flap following the creation and deployment of the atom bomb.
Roswell was one occasion of a downed craft, among others~\cite{Vallee2021May}.
The USA attempted to ``process'' alien artifacts similarly to Nazi and other Earth-bound (e.g., Soviet)
technology~\cite{SmithWilbertNotes, Sarbacher-83}. The matter became highly secretive, and only a small group of scientists,
such as von Neumann,\index{von Neumann, John} Eric Walker,\index{Walker, Eric} and Vannevar Bush,\index{Bush, Vannevar} were involved, along with military and government officials.

However, any ``Paperclip''-style~\cite{Jacobsen2014} processing of these downed crafts was inapplicable because of alien nonfraternization. Nobody could explain the technology like former Nazi party member SS Sturmbannf\"uhrer von Braun did.


Eisenhower's Military-Industrial Complex (MIC) oversaw these efforts, which were coordinated by the Air Material Command at Wright-Patterson Field, Ohio, and executed and processed by various DOE/DOD/industry partners (including Battelle and Lockheed).

The MIC achieved success with some ``proximity technology,'' which may have been ``on the verge of discovery'' regardless~\cite{Corso,Corso1998Jun}. However, the MIC failed to develop advanced technology, particularly in propulsion. It is thought that the science and technology of propulsion was based on ideas that were at least two scientific breakthroughs ahead of the resources available on Earth. The US made efforts to develop ``anti-gravity,'' but these proved fruitless.

This situation has persisted until today. As a result, the US may have obtained advanced crafts (although this is purely speculative), but is unable to replicate them on a larger scale. Perhaps the propulsion systems salvaged from some advanced crashed craft were used to construct new, self-made craft.

However, as sensors accessible to the general public and various interest groups have improved, the timeframe
for maintaining confidentiality and for ``kicking the saucer can further down the nondisclosure alley'' seems to be getting shorter.

A congressional initiative, started by a UFO enthusiast and real estate tycoon with strong connections to a powerful senator, led to the creation of a small private contract and an associated section in the Pentagon. Despite not achieving Special Access Program status, this project aggressively attempted to uncover what its proponents believed to be secret government programs relating to UFOs, like a screwdriver trying to open a corked bottle. They sent out one of their own, who wrote a secret memorandum of a conversation with a high-ranking military officer who suffered a bureaucratic defeat while attempting to determine what was inside that UFO bottle. Those in-the-know let the officer taste one bottle but refused access to more, in particular, the collection in their wine cellar.

In this scenario, the legacy media in 2017 got it all wrong: they mistook an anarchic grassroots movement---AAWSAP/AATIP---for the real thing, staring at the top of the iceberg, but not at its huge extent deep down in the water.

\section{Total control of Earth by ``Others'' leading to loss of human autonomy}
\label{2023-UFO-part-Speculation-executive-summary-tl} % Always give a unique label

This scenario involves the sizable abduction of a large number of humans, as described by David M. Jacobs~\cite{Jacobs1993Apr,Jacobs1999Mar,Jacobs2015Sep}, Don Crosbie Donderi~\cite{HJW1992,Donderi2013Mar,Donderi2013Jun},
John E. Mack~\cite{Mack-NYT1994Mar,Mack1994Apr,Mack1999PC}, and Budd Hopkins~\cite{Hopkins1981}.

The Others have prevented world governments and the United Nations from revealing their presence. One particular effort, near the end of the Cold War era, was thwarted by the Others when they abducted the Secretary General of the United Nations after a coordination session for disclosure. This was meant to send a clear message that such a revelation was not possible~\cite{hopkins-bbi}.

Extraterrestrial beings, including Reptiloids, Insectoids, and Humanoids, continue to carry out their activities on Earth, considering us their possessions~\cite{FortBotD}. Unfortunately, nothing can be done about this, as the situation is hermetic.



\section{The theological wild card: UFOs are daemonic}

According to an alleged secret group that is often referred to as the Collins Elite~\cite{Redfern2010Aug,Heiser2010Nov},
aliens are actually demons and fallen angels working for Satan.
These followers reinterpret UFO-related phenomena in a theological context~\cite{Kripal2011Nov,Pasulka2023Feb}
and believe that there is a satanic agenda behind alien abductions.
It presents a darker and more ominous perspective on the UFO phenomenon and consequently rejects related experiences~\cite[p.~219]{Pasulka2019Feb}.

Let me end this brief spectrum of UFO perception with a quote from Aurelius Augustinus, 354--430, who, in his ``Confessiones'' (Book XI, chapter 25), wrote:
{\em ``Ei mihi, qui nescio saltem quid nesciam!''}
(English translation: ``Alas for me, that I do not at least know the extent of my own ignorance!'')


\section{Uniform recommendation of government UFO panels: continue qualified research}

All government panels I am aware of recommended the initiation or continuation of some
(though modest size) academic efforts of research into unidentified phenomena of the UFO sort.
Let me quote from three of them.

The Durant Report~\cite{RobertsonPanelDurantReport}, subsuming the recommendations of the 1952--53 Robertson Panel,
emphasized that, regarding UFO's (aka ``Flying Saucers''),  there was ``no indication that these phenomena constitute a direct physical threat to national security,''
but stated:
\begin{svgraybox}
``Dr. Robertson at least was of the opinion that [[$\ldots$]] there would still be a role for a [[$\ldots$]] section
to cope with the residuum of items of possible scientific intelligence value.
This section should concentrate on energetically following up [[$\ldots$]]
those cases which seemed to indicate the evidence of unconventional enemy artifacts.''
\end{svgraybox}




In  ``Section I: Conclusions and Recommendations'' of the Condon Report~\cite{Condon-report,Condon-report-Bantam,Condon-report-Dutton,BibEntry2023Jan} of 1968,
Edward U. Condon writes:
\begin{svgraybox}
``Scientists are no respecters of authority. Our conclusion that study of UFO reports is not likely to advance science will not be
uncritically accepted by them. Nor should it be, nor do we wish it to be. For scientists, it is our hope that the detailed analytical
presentation of what we were able to do, and of what we were unable to do, will assist them in deciding whether or not they agree with
our conclusions. Our hope is that the details of this report will help other scientists in seeing what the problems are and the difficulties
of coping with them.

If they agree with our conclusions, they will turn their valuable attention and talents elsewhere. If they disagree it will be because
our report has helped them reach a clear picture of wherein existing studies are faulty or incomplete and thereby will have stimulated
ideas for more accurate studies. If they do get such ideas and can formulate them clearly, we have no doubt that support will be
forthcoming to carry on with such clearly defined, specific studies. We think that such ideas for work should be supported.''
\end{svgraybox}


In the ``Subsidiary Recommendations'' section of the ``Executive Summary'' of the UK Condign Report~\cite{CondignReport} of 1996,
Ron Haddow, its alleged author~\cite{Clarke2022Nov}, writes:
\begin{svgraybox}
``The relevance of plasma and magnetic fields to UAP were an unexpected feature of
the study. It is recommended that further investigation should be into the
applicability of various characteristics of plasmas in novel military applications.''
\end{svgraybox}

This summary also notes that:
\begin{svgraybox}
``Attempts by other nations to intercept the unexplained objects, which can
clearly change position faster than an aircraft, have reportedly already caused
fatalities  [[$\ldots$]]''

 ``The conditions for the initial formation and sustaining of what are apparently
buoyant charged masses, which can form, separate, merge, hover, climb, dive and
accelerate are not completely understood. Nevertheless, the underlying physics may
have some military application in the future in the form of active visual, radar and IR
decoys and passive electromagnetic spectrum energy absorbers.''
\end{svgraybox}

With regard to civil aviation, the following recommendations hold:
\begin{svgraybox}
\begin{itemize}
\item[--] ``It should be stressed that, despite the recent increase in UAP events, the
probability of encountering a UAP remains very low.
\item[--]  No attempt should be made to out-manoeuvre a UAP during interception.
\item[--]  At higher altitudes, although UAP appear to be benign to civil air-traffic,
pilots should be advised not to manoeuvre, other than to place the object
astern, if possible.''
\end{itemize}
\end{svgraybox}
