%%%%%%%%%%%%%%%%%%%%% chapter.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% sample chapter
%
% Use this file as a template for your own input.
%
%%%%%%%%%%%%%%%%%%%%%%%% Springer-Verlag %%%%%%%%%%%%%%%%%%%%%%%%%%
%\motto{Use the template \emph{chapter.tex} to style the various elements of your chapter content.}
\chapter{Abductions and experiences}
\label{2023-UFO-part-Perception-abductions} % Always give a unique label
% use \chaptermark{}
% to alter or adjust the chapter heading in the running head




\abstract*{There have been numerous reported cases of alien encounters and abductions throughout history. Many individuals claim to have been taken aboard an extraterrestrial spacecraft and subjected to various experiments, procedures, or observations. These experiences have been described by many as life-changing events, causing a dramatic shift in their attitude toward the existence of UFOs and existence in general. However, there is a lack of concrete evidence to support these claims. Some scientists have offered alternative explanations for these experiences, such as sleep paralysis and the possibility of false memories generated through hypnosis.}



\abstract{There have been numerous reported cases of alien encounters and abductions throughout history. Many individuals claim to have been taken aboard an extraterrestrial spacecraft and subjected to various experiments, procedures, or observations. These experiences have been described by many as life-changing events, causing a dramatic shift in their attitude toward the existence of UFOs and existence in general. However, there is a lack of concrete evidence to support these claims. Some scientists have offered alternative explanations for these experiences, such as sleep paralysis and the possibility of false memories generated through hypnosis.}


\section{Caveat}

This chapter presents accounts of alleged abductions that have the potential to be authentic. However, it should be noted that these accounts are speculative and may be purely fictional. They are highly controversial and have not undergone independent verification.

It is crucial to approach such claims with skepticism and to bear in mind that the credibility of the sources and the evidence presented should always be critically evaluated~\cite{Appelle1995,Appelle2000}. Additionally, it is worth noting that false or exaggerated claims are not uncommon, particularly in the context of controversial or sensational topics such as UFO sightings and alleged alien contacts.



\section{Deception operations against experiencers and UFO reporters}
\label{2023-UFO-part-Perception-abductions-do}
\index{deception operations}

Before discussing experiencers, I would like to mention that there have been speculations regarding some government agencies running deception operations against them and groups of UFO reporters and ``researchers'' trying to manage the situation and categorize what they have seen or been told. For instance, the CIA has feared that UFO reports might congest communication channels, thus diminishing alertness and response to attacks. However, increased public attention to flying saucers from outer space may also have positive side-effects. This category of flying saucers could be an esoteric classification of highly advanced, top-secret technologies and weapons developed by the respective states.


\subsection{UFOs serving as camouflage for secret military technology and spy planes}

As already stated, the UFO experiencer
and the wider community of UFO reporters and researchers who are trying to understand and categorize the phenomenon
is confronted with possible attempts at deception or ``education'' operations targeting them.
For instance, the CIA has been known to be concerned about ``hysterical mass behavior''~\cite[Chapter~4]{Jacobsen2011} and the potential for UFO reports to congest communication channels,
potentially lowering the ability to respond to attacks~\cite{Haines-CIA-UFO}.

However, heightened public interest in flying saucers from outer space may also have advantages,
such as the possibility that the category of flying saucers serves as an esoteric camouflage or disguise for
highly advanced and classified technologies and weapons developed by the respective countries.

For instance, reports of UFO sightings by commercial airline pilots and air traffic controllers began to increase after
U-2 planes started flying from Area 51.
The U-2s, which were silver at the time, created a reflection that led citizens to believe they were UFOs due to their high altitude
and unique wing shape. The planes flew at an altitude of 70\,000 feet, much higher than the commercial airplanes of the mid-1950s,
and their long, shiny wings made them look like a flying cross. Therefore, even if the CIA knew that U-2 sightings accounted for
a large portion of UFO sightings at the time, it did not disclose this fact~\cite[Chapter~4]{Jacobsen2011}.

\subsection{The controversial story of Paul Bennewitz}

Paul Frederic Bennewitz, Jr.\index{Bennewitz, Paul} was an American physicist and electronics engineer
known for his work in analyzing extraterrestrial signals and his strange story of possible government cover-up and deception.
Born on October 17, 1925, Bennewitz grew up in a family of scientists and engineers. He became interested in electronics at a young age and went on to study physics and electrical engineering at the University of Notre Dame.

During the 1970s, Bennewitz was a member of Arizona's branch of the Aerial Phenomena Research Organization (APRO).\index{APRO}
In that late decade, he became
intrigued by reports of strange lights and alleged unidentified flying objects (UFOs) near the Kirtland Air Force Base in Albuquerque, New Mexico. He began monitoring the signals he believed were coming from these objects and claimed to have decoded messages that he believed were from extraterrestrial beings.

Bennewitz shared his findings with others, including Air Force officials and researchers in the UFO community. At this point, the military became concerned about the security implications of Bennewitz's work regarding secret yet prosaic activities at Kirtland Air Force Base.

Some authors~\cite{Pilkington2010Sep,Pilkington2013} have speculated that Bennewitz was the victim of a deception operation attempted by the Air Force.
He was allegedly contacted by an agent who invited him to the base and shared ``secret UFO information of the government'' with him. The original goal of this operation might have been to recategorize secret military research, technology, and equipment as UFOs galore.
(One name dropped in this connection is Richard Doty\index{Doty, Richard}, but as I can neither confirm nor disprove this claim, I leave it at that.)

In the 1980s, Bennewitz became increasingly isolated and paranoid, and his health began to deteriorate. He was eventually diagnosed with schizophrenia and was admitted to a mental hospital, where he died in 2003. The story of Paul Bennewitz remains controversial to this day.


\section{Abduction saga I: Barney and Betty Hill abduction}
\label{2023-UFO-part-Perception-abductions-bbha}
\index{Hill abduction}

This is a summary of the Barney and Betty Hill abduction, as reviewed by Don Crosbie Donderi~\cite{Donderi2013Jun}.
Barney and Betty, an interracial couple, were driving home from Vermont to Portsmouth through northern New Hampshire on US Route 3 on the night of September 19. They stopped at a restaurant in Colebrook, New Hampshire and left just after 10:00 p.m. A storm was moving up the east coast, giving them an incentive to get home soon. While they were driving, they noticed a light in the sky above Jupiter that seemed to be moving and bigger than the planet. They slowed the car to get a better look and thought it might be a satellite. As they drove south, the light was sometimes visible as the road turned and trees or mountains obstructed their view, so they found it hard to tell whether it was actually moving. They stopped the car, saw that it was moving, and got out to look through binoculars. They saw a disc with a row of brightly lit rectangular windows, sixty to eighty feet in diameter and approximately twenty feet tall. The disc hovered over the road in front of their car, then moved silently to the left
and came to rest over the tree line just past the edge of a field.
Barney got out of the car and walked closer to the disc with his binoculars, but saw black-clad figures inside.
One figure stared at him, and he felt he was about to be captured.
Betty screamed for him to come back to the car, and they drove away. As they fled, the disc followed them, and they heard buzzing tones coming from the car trunk. They had unclear memories of seeing a manned roadblock and a fiery globe on the ground.
They continued on their journey and reached home just after 5:00 am, later than their predicted arrival time.

Approximately ten nights after the UFO sighting, Betty began experiencing nightmares that lasted for five consecutive nights. Barney did not have similar nightmares and did not think they were anything serious. Betty wrote down her nightmares to alleviate her anxiety. Two years after the sighting, Barney had a traumatic flashback of the event, and several people suggested hypnosis to remove the amnesia that Barney and Betty had experienced. They were referred to a psychiatrist, Dr. Benjamin Simon, who used hypnosis to help soldiers with post-traumatic stress disorder. Under hypnosis, Barney and Betty recalled being taken into a UFO and examined by the occupants. Barney remembered a rectal, groin, and
genital/urological probe, and the occupants were surprised to find that Barney had removable teeth and Betty did not.

Betty recalled an alien being whom she referred to as the ``leader.''
Despite not communicating in English, Betty was able to understand the leader as if he were speaking the language.
The leader performed a medical examination on Betty, which included the insertion of a large needle into her navel, causing her pain.
However, the leader could wave his hand over her face and make the pain go away.

After the examination, Betty was left alone with the leader, and they had a conversation.
She expressed her desire to take a souvenir with her.
 The leader offered to let her take something.
She saw a book-like object lying on a cabinet and asked if she could take it, even though she could not read it.
The leader agreed to give her the book, and she then asked him where they were from. The leader inquired if she knew much about the universe, and after she shared her limited knowledge, he opened a star map. Described as being three feet wide, oblong-shaped, and almost self-luminous, like a holographic projection, it displayed bright circles of different sizes connected by solid lines between larger circles and some dotted lines. The leader explained that the circles represented stars or planets, the lines represented trade routes, and the dotted lines represented expeditions. He asked her if she knew where she was on the map, and she said no. The leader then stated, ``If you do not know where you are, there would not be any point in me telling you where I come from,'' and closed the map.

As Betty was getting ready to leave after her encounter with the leader, a commotion broke out among the crew. A shorter, round-faced individual in a position of authority objected to Betty being allowed to keep the book, and she was forced to give it back. The leader also informed her that they had decided to prevent Barney and Betty from remembering their abduction. They explained that even if Betty remembered anything, Barney would not, and if he did, his memories would contradict hers and they would be left confused. Betty returned to the car where Barney was already sitting behind the wheel, and Delsey the dachshund had been left during the abduction.


Barney and Betty's mechanical wristwatches stopped during this incident.


\section{Abduction saga II: Travis Walton incident}
\label{2023-UFO-part-Perception-abductions-twui}
\index{Travis Walton incident}
\index{Walton, Travis}

The following story might have been invented for the sake of money.

Travis Walton was a 22-year-old man working on a logging contract in Arizona.
On November 5, 1975, he was working with a timber stand improvement crew in the Apache-Sitgreaves National Forest near Snowflake, Arizona.
In the evening, while leaving the forest in a truck, they noticed a craft descending from the trees.
All seven witnesses, including Walton, saw a saucer-shaped object hovering over the ground approximately 110 feet away,
 emitting a high-pitched buzz.
Walton, seated in the front passenger seat, left the truck and approached the craft.
Despite his friends' warnings to come back, he stood in front of the object.
Suddenly, a bright blue light beam shot out, hit him in the chest, and knocked him back 10 to 15 feet onto the ground.
His friends, fearing for their lives, drove away in panic.

After several minutes of distancing themselves from the dangerous craft that had shaken Walton to the ground, Walton's friends
regained their composure and decided to return to the site. However, when they arrived, Walton was nowhere to be found.
His disappearance triggered a massive search effort that involved many people, police, tracking dogs, helicopters, and horseback riders.
Despite the extensive search, no trace of Walton was found for five days.

For five days, the authorities could not exclude that Walton had been murdered by his coworkers.
He eventually turned up unshaven and dehydrated and made a phone call from a payphone.
This sparked media attention and a circus-like atmosphere, with UFO researchers descending upon the area.

Walton claimed that during his abduction, he woke up in a hospital-like room with three short, bald creatures.
Later he changed location by leaving a small craft into some sort of mothership or hangar. A ``man'' in blue uniform with a helmet
brought him into a room with other human-like beings without helmets
who tried to push him over to a table where they put a mask over his face.
No verbal or telepathic communication took place. The next thing Walton knew,
he was on the ground on Earth with the craft taking off above
him~\cite{WaltonNelson1,WaltonNelson2,Walton1978Jan,DolanDisclosure2018Sep}.

Walton wrote a book about his experiences~\cite{Walton1978Jan} and made several appearances on talk shows~\cite{WaltonPowerfuljre2021Jan}.

Skeptics view the Travis Walton abduction case as a hoax perpetuated by a credulous media circus.
Walton allegedly exploited his publicity to make money.
Klass, a UFO skeptic, pointed to alleged discrepancies in Walton's and his coworkers' accounts.
Klass also reported that the polygraph tests were poorly administered and that Walton
used countermeasures during the test. Clancy speculated that Walton may have been influenced by the NBC television movie called
``The UFO Incident.'' The movie aired two weeks before Walton's claimed abduction and
dramatized the claims of Barney and Betty Hill. After the airing of the movie, there was a rise in alien abduction claims.
Clancy concluded that Walton's narration made him an instant celebrity.



\section{Different types of experiences and abductions}
\label{2023-UFO-part-Perception-abductions-dt}

\subsection{Individuals having sightings versus experiencers versus abductees}

%The percentage of individuals who have experienced a ``hard case'' UFO/UAP sighting---one that cannot be immediately ruled out by obvious mundane phenomena or causes---is
%challenging to determine.
%On the one hand, many sightings remain unreported for various reasons such as neglect or disinterest, or fear of ridicule.
%
%On the other hand, many such reports turn out to be simple fabrications, for instance
%by people seeking to exploit the public interest for their own gain,
%as well as the subconcious wish to be extraordinary, to obtain meaning and gratification in an otherwise boring and monotonous life.
%Thus distinguishing between credible sightings and hoaxes or misidentifications can be difficult.
%
%Moreover it remains unclear what proportion of individuals, referred to as experiencers,
%have had an ``essential experience.''
%Here I aim to differentiate between either a occational simple sighting without deeply affecting that person,
%or an experience that is associated with an accompanying ``ontological shock,''
%as explained in Section~\ref{2023-UFO-part-Perception-abductions-Ontologicalshock}.
%
%A third category of deep experiencers is those of so-called abductees, people who allegedly have been abducted by aliens.
%The number of people classified as abductees is uncertain. Although it may seem challenging to differentiate between abductees and experiencers, according to Jacobs~\cite{Jacobs2015Oct}, an individual is either one or the other, with few cases in-between. An estimate~\cite{HJW1992,Donderi2013Mar} from 1992 suggests that approximately 2{\%}
%of Americans may have experienced the ``UFO Abduction Syndrome.'' Don Crosbie Donderi\index{Donderi, Don Crosbie} claims to be able to use a questionnaire to determine with a high level of certainty whether a person has been abducted~\cite{Donderi2013Jun}.
%
%Let me outrightly state that I myself have had none such experiences, nor do I wish to enter into debates into the factuality of such experiences or abductions.
%My own position is that those who claim to have an experience and cannot outrightly be dismissed as profeteers or hoaxers
%have some genuine experience which should be taken seriously in an epistemic but not necessarily ontic kind of way.
%Stated differently, I here discuss experiences rather than facts.
%Factuality is important but remains uncertain.


Estimating the percentage of individuals who have encountered a ``hard case'' UFO/UAP sighting---one that cannot be easily dismissed as being due to ordinary phenomena or causes---is a complex task. On the one hand, numerous sightings go unreported due to various factors such as:
\begin{itemize} 
\item neglect or disinterest, 
\item fear of ridicule. 
\end{itemize}
On the other hand, a significant portion of these reports may be mere fabrications, created by individuals who:
\begin{itemize} 
\item seek to exploit public interest for personal gain, 
\item have a strong desire to feel extraordinary and find meaning in their otherwise mundane lives. 
\end{itemize}
Consequently, distinguishing between credible sightings and hoaxes or misidentifications proves to be challenging.

Furthermore, it is still unclear what percentage of individuals, also known as experiencers, have had a profound encounter, which is typically accompanied by an ``ontological shock,'' as discussed in Section~\ref{2023-UFO-part-Perception-abductions-Ontologicalshock}. Such an encounter goes beyond a simple sighting that may have little impact on the individual.

A third category of deep experiencers consists of the so-called abductees, individuals who claim to have been abducted by extraterrestrials. The number of people identifying as abductees remains uncertain. Although differentiating between alleged abductees and alleged experiencers may appear challenging, Jacobs~\cite{Jacobs2015Oct} posits that an individual is typically either one or the other, with few cases falling in-between. A 1992 estimate~\cite{HJW1992,Donderi2013Mar} suggests that approximately 2{\%} of Americans may have experienced the UFO Abduction Syndrome. Don Crosbie Donderi\index{Donderi, Don Crosbie} asserts that he can use a questionnaire to determine with a high degree of accuracy whether a person has actually been abducted~\cite{Donderi2013Jun}.

Allow me to clarify that I have not had any such experiences myself, and I do not intend to engage in debates regarding the factuality of these experiences or abductions~\cite{Sagan1997Feb}. My stance is that many individuals who claim to have had an experience and cannot be easily dismissed as profiteers or hoaxers likely have a genuine experience that should be treated seriously in an epistemic manner, though not necessarily in an ontic sense. In other words, this discussion focuses on experiences rather than facts. While factuality is important, it remains uncertain and highly questionable.

\subsection{Time and situation of abductions}

I believe that, just as in cases of UFO sightings, it is of great importance to separate the wheat from the chaff,
to sort out actual abductions or close encounters with aliens (if they exist)
from more prosaic experiences such as nocturnal sleep paralysis or false memories.

However, while it is rather likely that a high percentage of claimed nocturnal encounters with aliens might result from
symptoms of sleep paralysis interpreted as alien abductions, some of them might actually be caused by aliens.
Daytime experiences---with people experiencing ``missing time'' and being transferred to different locations than their pickup positions---might
be an entirely different sort.

In any case, my impression is that there is a divide between people who are ``experiencers'' of some sort---some of whom claim to have occasionally
or regularly seen some ``glowing orbs,'' others who claim to have stumbled upon some craft, sometimes even with inhabitants, and others who claim to have
been forced into an alien craft---and ``other folk'' (like myself) who have had no such experiences.

My observation is that experiencers feel very strongly about the existence of an alien presence, even if the experience took place a long time ago or is sporadic.
It is as if they need no further proof, and skepticism by non-experiencers appears totally inappropriate to them.
It is almost as if they were trying to explain color to the color blind.

Whatever these ``experiences'' or ``abductions'' may be, my feeling is that they have not been taken seriously in the medical community.
Schizophrenic or psychotic mental conditions may cause some of these.
Likewise, post-traumatic stress disorder can result, similar to what occurs in situations of war or terrorism.
I am not suggesting that all cases can be explained by mental conditions, or that they are alien-induced; although one should not a priori rule out such causes.

Many individuals affected by such incidents require assistance but are not adequately supported by psychologists, psychiatrists, or society. They deserve empathy and a greater understanding from us,
and this process may eventually also help ``us''---the majority---to understand ourselves better.

\section{Ontological shock}
\label{2023-UFO-part-Perception-abductions-Ontologicalshock}
\index{ontological shock}

Ontological shock\index{ontological shock} refers to an experience that challenges one's understanding of reality, causing disruption. This can be a profound and unsettling experience that requires individuals to re-evaluate their long-held beliefs and assumptions.

John E. Mack, a Harvard psychiatrist, worked with patients who claimed to have experienced ontological shock related to UFO encounters. Specifically, these individuals believed that their experiences of being abducted by aliens were genuine and that their perception of reality had been forever altered. Initially, many of these patients approached Mack with what they called ``vivid dreams,'' which Mack referred to as denial. Mack used hypnosis to help his patients overcome this denial and accept their experiences as genuine~\cite{Mack1994Apr,Mack-NYT1994Mar}.


\section{Sleep paralysis may account for a large number of ``abductions''}

Nocturnal UFO abduction experiences are frequently described as people being taken against their will by extraterrestrial beings and subjected to various physical and psychological examinations. These experiences are typically reported by individuals who claim to have been abducted by aliens and often involve vivid hallucinations and a sensation of being paralyzed or unable to move.

Therefore, it has been suggested that ``alien abductions'' have a mundane explanation, such as sleep paralysis~\cite{Holden_2002,Clancy2007Apr}. During sleep, the body becomes paralyzed as a protective measure to prevent injury. However, it is possible to fall asleep or wake up while still in a paralyzed state, leaving the person incapable of movement. During an episode of sleeping paralysis, an individual may also experience vivid hallucinations and an intense sense of fear or anxiety.

When this happens, people may experience hallucinations that seem real but are actually a product of their imagination. Sleep paralysis is a common experience, affecting approximately 25{\%}
of people worldwide, and approximately 5{\%}
of people experience a very dramatic form thereof---that is, they also have hallucinations involving their sense of sight, sound, touch, and even abduction experiences. Some of them become obsessed with seeking an explanation for it, which can lead them to adopt unscientific explanations~\cite{Spanos_1993}.

Furthermore, a study found that adults who reported memories of childhood sexual abuse, whether repressed, recovered, or continuous,
were more likely to experience sleep paralysis than a control group without such experiences~\cite{McNally-Clancy}.
Sleep paralysis may be linked to irregular sleep patterns, stress,
and even certain medications~\cite{Sharpless_2016}.
It has also been suggested that ``many of the frequently reported particulars of the abduction experience
bear more than a passing resemblance to medical-surgical procedures
and propose[[d]] that experience with these may also be contributory''~\cite{Forrest_2008}.

However, it may be more than that. This group of ``pseudo-abductees'' might find depth and meaning in their lives through ``alien communion.'' In her book, ``Abducted"\cite{Clancy2007Apr}, Susan A. Clancy suggests that those who claim to have been abducted by extraterrestrials are not mentally ill but rather intelligent individuals who have formed false memories due to a combination of nightmares, exposure to alien-related media, and a desire for something greater than science can provide. It gives their lives meaning\cite{Frankl2006Jun}, a glowing light of orientation and proof of the experiencer's exceptionalism in an otherwise dreary life. For these individuals, the experience of abduction can be transformative and meaningful. Clancy's book delves into the intricacies of memory and the nature of belief.


\section{The sexual connotation: arousal and rape}
\label{2023-UFO-part-Perception-abductions-sexual}
\index{sexual arousal}

There is a sexual component to abduction that Mack acknowledged in ``Abduction''~\cite{Mack1994Apr}.
Mack writes, ``the reported intrusive sexual and reproductive procedures that are a central aspect
of the abduction phenomenon can profoundly affect the intimate life and general well-being of abductees.''

In his book ``{C}ommunion: {A} True Story''~\cite{Strieber1987Jan}---whose title page shows an archetypical ``gray alien''---Whitley Strieber
observes almost romantic feelings toward aliens as follows: ``To me this is a woman, perhaps because her movements are so graceful,
perhaps because she has created states of sexual arousal in me, or maybe it is simply the memory of her hand touching the side of my chest one time,
so lightly and yet with such firmness.''
He also mentions that ``in recent years, many of the taken have reported having sexual experiences with the visitors.''

Strieber also mentions
a notable case of a sexual encounter with an extraterrestrial that occurred in Brazil in 1957,
where a farmer drove his tractor at night in one of his fields. The tractor suddenly died, and he allegedly
saw an object that had landed in front of him. He was stripped, washed, taken inside the object and left on a table.
Later, he saw a woman, resembling a human---her hair being blond and parted in the center, with an extremely wide face,
blue and slanted eyes, thin, nearly invisible lips---entered the room and had sexual intercourse with him.
She pointed to her belly and then to the sky and left him.

In one of Strieber's podcasts with Mack, the latter discusses the connection between sex
and alien abductions at some length~\cite[time=26m30s]{Strieber2023Jan} and Mack's quotation of Strieber's testimony
in Mack's book ``Passport to the {C}osmos''~\cite{Mack1999PC}:
\begin{svgraybox}
``The sexual part of [[Strieber's]] relationship [with the beings] has been very complex and very rich and very difficult at times
because I'm a married man.
I've taken marriage vows, which I believe in.
To an extent this aspect of it has been thrust on me, and it's not something I've been able to control.
If it had been under my control, I would have felt very guilty.''
There is a specific alien female with whom [[Strieber]] feels mated.
It's like having a second wife with whom I have a secret relationship~[[$\ldots$]]
There was never a seduction,~[[$\ldots$]]  I would wake up in a state of sexual excitement, in mid-intercourse~[[$\ldots$]]
The physical dynamic is different in the sense that the sensation of intercourse moves through your whole body,
and you become totally devoted to it for longer than I do in normal intercourse~[[$\ldots$]]
as if a level of sexuality I'm normally not involved with is engaged. It's very, very powerful.''
\end{svgraybox}


\index{rape}
Nocturnal abductions may not always be enchanting but can be traumatizing. According to David M. Jacobs, a UFO researcher, aliens may be abducting humans and performing invasive medical procedures to learn about the human body, particularly the reproductive system. Jacobs believes that aliens are collecting human sperm and eggs for their own breeding program, which could result in the creation of human-alien hybrid beings.

In a 1994 article in the Los Angeles Times~\cite{Hulse2019Mar}, a story is recounted that is similar to approximately 60 stories in Jacob's book ``Secret Life: Firsthand Documented Accounts of UFO Abductions"~\cite{Jacobs1993Apr}. The story involves Bill, a 42-year-old married medical professional with two children and a devout Christian belief. Bill underwent regressive hypnosis with a hypnotherapist, which led him to believe that he had been abducted by aliens and taken to a spaceship for examination. During the examination, an instrument painfully probed his rectum, and another collected semen from him. Bill described the experience as feeling like he was being raped. Despite not profiting from these claims, the experience has caused strain on his marriage, and he is unable to discuss the topic with his wife.

\section{Possible psychotic or schizophrenic origins}
\label{2023-UFO-part-Perception-abductions-pposo}

Psychosis and schizophrenia are medical conditions that can lead to changes in a person's thoughts, emotions, and behavior. Psychosis refers to a state in which a person loses touch with reality and may exhibit symptoms such as hallucinations (seeing or hearing things that are not there) or delusions (fixed, false beliefs). Schizophrenia is a mental disorder that frequently includes psychosis as a symptom and may also involve disorganized thinking, communication difficulties, and alterations in mood and motivation.

The exact causes of psychosis and schizophrenia are not yet entirely understood, but research suggests that a combination of genetic, environmental, and brain chemistry factors may be involved. While some individuals experiencing psychosis may report hallucinations or delusions involving extraterrestrial abductions or other unusual events, it is important to note that not all such experiences are necessarily caused by psychosis. Other factors, such as sleep deprivation, stress, use of certain drugs, false memories~\cite{Clancy_2002m}, or physical conditions such as sleep paralysis, can also contribute to similar experiences. It is also important to bear in mind that many individuals who report such experiences do not have any underlying mental health conditions~\cite{Holden_2002}.

Therefore, it is possible that some of these recollections may indeed be caused by actual extraterrestrial abductions or other strange events, making it challenging to distinguish and cope with~\cite{Appelle1995,Appelle2000}. As for myself, I would be uncertain how to categorize such experiences if I were to encounter them.


\section{Alien subversion?}
\label{2023-UFO-part-Perception-abductions-alsub}

David M. Jacobs~\cite{Jacobs1993Apr,Jacobs1999Mar,Jacobs2015Sep,Jacobs2015Oct},
a researcher who has conducted many regression sessions with individuals claiming alien abduction, identified patterns in their accounts that formed a classic abduction scenario. He concluded that these experiences were not merely occupations but rather invasions in which the aliens were creating human--alien hybrids. Jacobs believes that these beings have been coexisting with humans for a considerable time and that hybridization is not a recent occurrence.

Jacobs has also pointed out that the memories of abductees, if not false, may be the only way of ``looking into alien craft from the inside.''

\section{Fort's ``we're property'' hypothesis}
\label{2023-UFO-part-Perception-abductions-wao}

There are some darker theories that suggest aliens see us as their property and Earth as their farm. According to these theories, we are nothing more than mere possessions to them, despite our belief in our own independence and free will. Essentially, these theories posit that we are under their control.

Charles Fort\index{Fort, Charles}, in Chapter 12 of ``The Book of the Damned"~\cite{FortBotD,Fort1975-vd}, was one of the first authors to suggest this hypothesis in a secular context. While religions often point this out, Fort proposed this theory back in 1919:
\begin{svgraybox}
``$\ldots$~that near approach by another world to this world would be catastrophic:
that navigable worlds would avoid proximity;
that others that have survived have organized into protective remotenesses,
or orbits which approximate to regularity, although by no means to the degree of popular supposition.''

$\ldots$

``Would we, if we could, educate and sophisticate pigs, geese, cattle?''

``Would it be wise to establish diplomatic relations with the hen that now functions,
satisfied with mere sense of achievement by way of compensation?''


``I think we're property.

I should say we belong to something:

That once upon a time, this earth was No-man's Land, that other worlds explored and colonized here,
and fought among themselves for possession, but that now it's owned by something:

That something owns this earth---all others warned off.''

$\ldots$

``Nothing in our own times---perhaps---because I am thinking of certain notes I have---has ever appeared upon this earth, from somewhere else,
so openly as Columbus landed upon San Salvador, or as Hudson sailed up his river.
 But as to surreptitious visits to this earth, in recent times, or as to emissaries, perhaps, from other worlds,
or voyagers who have shown every indication of intent to evade and avoid,
we shall have data as convincing as our data of oil or coal-burning aerial super-constructions.''


$\ldots$

``Pigs, geese, and cattle.

First find out that they are owned.

Then find out the whyness of it.

I suspect that, after all, we're useful---that among contesting claimants, adjustment has occurred,
 or that something now has a legal right to us, by force,
or by having paid out analogues of beads for us to former, more primitive,
owners of us---all others warned off---that all this has been known, perhaps for ages, to certain ones upon this earth,
a cult or order, members of which function like bellwethers
 to the rest of us, or as superior slaves or overseers,
directing us in accordance with instructions received---from Somewhere else---in our mysterious usefulness.

But I accept that, in the past, before proprietorship was established,
inhabitants of a host of other worlds have---dropped here, hopped here,
wafted, sailed, flown, motored---walked here, for all I know---been pulled here, been pushed;
have come singly, have come in enormous numbers; have visited occasionally,
 have visited periodically for hunting, trading, replenishing harems, mining:
have been unable to stay here, have established colonies here, have been lost here;
far-advanced peoples, or things, and primitive peoples or whatever they were:
white ones, black ones, yellow ones~$\ldots$''
\end{svgraybox}

Fort's hypothesis can undoubtedly be considered as wild speculation with seemingly no substantial empirical evidence.
I suppose many critical readers might blame me for even bringing it up.
However, it may still be worthwhile to include this material for the sake of historical completeness.
Additionally, discussing this hypothesis can shed light on the innate human tendency to fantasize and invent stories and devise explanations,
regardless of how far-fetched they may appear.

On the other hand, some individuals may argue that Fort's ideas have not been conclusively disproven, and that their personal experiences lend support to at least some of his speculations.
However, we need to recall Carl Sagan's critical stance on any appeal to ignorance---the claim that
``whatever has not been proved false must be true''~\cite{Sagan1997Feb},
and that ``extraordinary claims require extraordinary evidence''~\cite{Laplace1814,Flournoy1900,Russell1952}.

In Fort's line of speculation, and similar to Brinsley Le Poer Trench's\index{Le Poer Trench, Brinsley}
``The Sky People''~\cite{lePoerTrench1961Jan},
William Bramley's~\cite{Bramley1993Mar}  ``Gods of Eden''
is a critical look back at human history that starts in ancient
times and moves forward, examining ancient texts and relating them (in an almost literary way at times) to the UFO phenomenon in prehistory.
He quotes the Bible, interpreting the descriptions of ``God'' as literal accounts of thundering,
smoke, and aerial phenomena that resemble a modern rocketship.
Bramley argues that a custodial race---the Others---of advanced beings formed the basis of ancient religions such as the
Hebrews, Sumerians, Mayans, and Aztecs.

He claims that these same rulers manipulated and controlled secret societies,
religions, and wars to maintain their power and foster violence.
Bramley also connects the custodians to modern events such as the Crusades,
the Inquisition, the financial system, and revolutions, showing their role in creating a system of economics, politics,
and social control based on war and a ``strategia della tensione''
as Italians call it, a strategy of tension---thereby maintaining control over humanity.
According to Bramley, there are no significant events in human history that occur by mere chance,
but rather they are the result of the hidden influence and manipulation of the custodials, the Others.


\section{Demonology}
\label{2023-UFO-part-Perception-abductions-dem}

The reception of the UFO phenomenon could not possibly become any more bizarre:
Nick Redfern reports~\cite{Redfern2010Aug,Heiser2010Nov} about the mysterious and frightening realm of the alleged ``Collins Elite'',\index{Collins Elite}
a covert group within the US government. This group holds the belief that extraterrestrial life as we know it is actually comprised of deceitful demons and fallen angels, serving as minions of Satan and harvesting humanity's souls,
ultimately leading to the end of the world and the Day of Judgement.

The Collins Elite believes that the extraterrestrial life that we are aware of is actually comprised of demonic entities.
The group  came to believe that the wave of UFO sightings was the result of these entities coming into our world through portals
opened by occult practices.

Their research focuses on the relationship between these entities, UFO sightings,
alien abductions, and other paranormal issues such as out-of-body experiences and life after death.

The Collins Elite believes that the ultimate goal of the demons masquerading as aliens
is to control humanity through a new world order and enslave humanity by harvesting our souls.
The group is allegedly divided on what to do about this situation.
Some members have attempted to establish a Christian theocracy to combat the demonic forces.
The group fears that publicizing their findings would result in mass panic and hysteria,
leading to biblical end times.


Complaints have been made that the Collins Elite in the Pentagon has effectively and successfully blocked research into
the UFO phenomenon, for the reasons just explained.

\section{Alien implants and DNA analysis?}
\label{2023-UFO-part-Perception-abductions-alim}

Roger Leir allegedly extracted alien implants from people who claimed to have been abducted and
found ``objects'' in their body~\cite{Leir2000}.
This raises the possibility that the implants are used for monitoring people~\cite{DolanDisclosure2018Sep} but
thus far, all of these ``objects'' seem to have mundane origins.

One example of an investigation~\cite{ScottLittle-implant} into an alleged ``implant'' found that the individual who had it removed was likely struck by a piece of a tool while they were preoccupied with other events. This resulted in a cut that was thought to have healed, but over time, the body developed a reddish lump of scar tissue to isolate the object.

In my opinion, the claims of ``alien implants'' are either false or a deliberate deception.

The ``Hair of the Alien'' is a book on the DNA analysis of alleged alien origin~\cite{Chalker2005Jul}.

% Dr. Alex Mosier

\section{Connections to psychoanalysis and similarities to Freud's seduction hypothesis}
\label{2023-UFO-part-Perception-abductions-sfsh}

\subsection{Freud's abandonment of hypnosis}

Hypnotherapists have published books on alleged abductions~\cite{Marden2019Aug,Lalich2020,Smith2008Feb}
that are directed toward those who feel uneasy and have alleged memories or flashbacks of abduction episodes.
It is interesting to compare these and Mack's earlier attempts~\cite{Mack1994Apr,Mack1999PC} using hypnosis
with Freud's own experiences and methodological transformations.
Freud started using hypnosis in his early work as a therapist,
but eventually abandoned it in favor of his new method of psychoanalysis,
which involved the use of free association and interpretation of unconscious thoughts and feelings.
He found that hypnosis was not as effective as he had hoped and that it often led to false memories and confabulations.
Additionally, he believed that the talking cure of psychoanalysis
was a more direct and efficient way of accessing the unconscious mind.

For these very reasons, I suggest that one should be very cautious when interpreting statements obtained by hypnosis.


\subsection{Memories of childhood abductions}

In 1896, Sigmund Freud presented his revolutionary ``seduction theory''
in his article ``The Etiology of Hysteria,''
arguing that acts of sexual abuse and violence inflicted on children are the direct cause of adult mental illness.
These ideas were not well received, since they might have implied that many of his clients and patients
in ``reality'' had been sexually abused by their fathers or male relatives---in particular, many colleagues and prominent members of
the Viennese society.

Nine years later, Freud completely reversed his position:
he blamed his patients for having deceived him on that issue,
insisting that these sexual memories were actually fantasies that never happened.
He subsequently launched a ``fantasy'' theory to explain the development of hysterical
symptoms without the need for any actual abuses~\cite{Masson1984,Eisen2001Jan,Bogousslavsky_2014}

In his book ``The Assault on Truth: Freud's Suppression of the Seduction Theory,''
Jeffrey Masson criticizes Sigmund Freud's rejection of seduction theory,
which posited that many of his patients' psychological problems were rooted in childhood sexual abuse.
According to Masson, Freud initially believed that his patients' accounts of sexual abuse were true,
but later discarded this theory under pressure from colleagues and his own unconscious conflicts.
Masson argues that Freud's rejection of seduction theory marked a turning point in his work and had far-reaching implications for the development of psychoanalysis.
The book remains a controversial and widely debated aspect of Freudian scholarship.

The intriguing similarity between Masson's claims and the disregard of supposed alien abductions is that both raise the question of what actually occurred. Did the patients actually experience abductions,
or did they succumb to false memories and hysterical delusions?


\subsection{Some personal observations}

I am convinced that we need to be aware that not all experiencers are hoaxers or profiteers, and that what some experiencers share is at least authentic, mind-boggling, and subjectively true. However, this kind of openness and suspended attention should not, under any circumstances, lead researchers to unquestioningly accept the factuality of these narratives and memories.

Likewise, experiencers need to be aware that their memories and experiences may not be entirely factual. These mental contents can be interpreted in various ways:
\begin{itemize}
\item they might be (partly) imagined and fantasized,
\item some profound trauma may have been translated or transferred by the mind to an alien context,
\item or they might not be imagined at all, but have actually occurred to that person in some form.
\end{itemize}
In any case, it is crucial to strive for a rigorous and empirical approach when investigating anomalous experiences. One effective way to do this is to augment, falsify, or corroborate such experiences with physical sensors, such as infrared, radar, or visual sensors. By doing so, we can attempt to establish a correlation between these objective measurements and the subjective experiences reported by individuals.

However, simply not believing, laughing at, or distrusting genuine subjective experiences by experiencers is,
in my opinion, an inadequate response. Many experiencers seem to be very much afraid---almost ashamed---to talk about their experiences.
This hesitancy might also be due in part to the attempt to self-repress such experiences, as a way to cope with the ontological shock that comes with them.
Indeed, if they told these stories out loud to the world, this would document their own (maybe involuntary) difference, or even strangeness,
which might be a challenge both to them and also to their peers and the religious or secular societies they live in.

Therefore, I would suggest creating friendly spaces and environments so that nobody is afraid to recount their extraordinary, sometimes very strange, and highly personal stories. Additionally, maintaining a critical scientific attitude that offers clinical psychological help and interdisciplinary research into these matters seems essential.
