%%%%%%%%%%%%%%%%%%%%% chapter.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% sample chapter
%
% Use this file as a template for your own input.
%
%%%%%%%%%%%%%%%%%%%%%%%% Springer-Verlag %%%%%%%%%%%%%%%%%%%%%%%%%%
%\motto{Use the template \emph{chapter.tex} to style the various elements of your chapter content.}
\chapter{Crash retreivals and UFO materials}
\label{2023-UFO-part-Perception-crash-retreivals} % Always give a unique label
% use \chaptermark{}
% to alter or adjust the chapter heading in the running head


\abstract*{In this speculative and possibly fictional chapter, the authenticity of two crash retrieval accounts is uncertain: Corso's alleged experiences with the Roswell incident and beyond and the alleged notes taken by Eric Davis from a meeting with Admiral Wilson. These accounts, akin to Rosetta stones, are used to understand the relationship between the ``canonical normalcy'' of the world and secret black ops programs related to alien technology. If genuine, they imply that the US industrial-military complex has tried to comprehend and replicate technology from recovered alien craft to gain supremacy over competitors on earth, as well as to deal with the possibility of alien domination. This interpretation would, if correct, create a modified version of history after the Trinity test. One implication is that extraterrestrial crash materials and attempts to recreate alien vehicles are likely to be found in carve-out unacknowledged special access programs with precarious Congressional oversight.}

\abstract{In this speculative and possibly fictional chapter, the authenticity of two crash retrieval accounts is uncertain: Corso's alleged experiences with the Roswell incident and beyond and the alleged notes taken by Eric Davis from a meeting with Admiral Wilson. These accounts, akin to Rosetta stones, are used to understand the relationship between the ``canonical normalcy'' of the world and secret black ops programs related to alien technology. If genuine, they imply that the US industrial-military complex has tried to comprehend and replicate technology from recovered alien craft to gain supremacy over competitors on earth, as well as to deal with the possibility of alien domination. This interpretation would, if correct, create a modified version of history after the Trinity test. One implication is that extraterrestrial crash materials and attempts to recreate alien vehicles are likely to be found in carve-out unacknowledged special access programs with precarious Congressional oversight.}



\section{Caveat}

This chapter presents two potentially authentic but highly speculative and unconfirmed accounts of events related to crash recoveries: the Roswell incident and a meeting with Admiral Wilson. These accounts offer insight into secret programs aimed at recovering, understanding, and replicating alien technology, as well as the US military and industrial complex's attempts to cope with the potential superiority of aliens and the technological advancements that may have been gained from recovered crafts. If true, these accounts suggest a revisionist history in which the US has sought to adapt to the presence of extraterrestrial beings.

It is important to note, however, that these accounts, particularly Corso's story and the Wilson Memo, are speculative and may be purely fictional. They are highly controversial and have not been independently verified. Therefore, it is crucial to approach such claims with skepticism and evaluate the credibility of the sources and evidence presented. Additionally, it is not uncommon for individuals to make false or exaggerated claims, particularly in the context of controversial or sensational topics such as UFO sightings and alleged recoveries of alien technology.

\section{Missing artifacts from ancient aliens}
\label{2023-UFO-part-Perception-crash-retreivals-aa}

Although it is not possible to say exactly how certain objects from ancient times---such as the pyramids in Egypt---were built, there are plausible scenarios that make it seem at least possible that these objects could have been constructed with the technology and tools available at the time of construction~\cite{Arnold1997Feb,Arnold1992}.

The same can be said about specific objects or groups of objects, such as granite masonry with almost perfect joints and obelisks, such as the Serapeum of Saqqara. Nothing points to extraterrestrial help; their origins appear prosaic~\cite{Arnold1988,Arnold1992,Arnold1997Feb,Arnold2000,Arnold2008}.

% https://www.quora.com/How-did-an-ancient-civilization-build-the-granite-boxes-at-the-Serapeum-of-Saqqara

There may be artifacts, such as grave goods, that surpass the materials of the Bronze Age. For instance, an iron dagger was found in the tomb or mummy of Tutankhamun, but it was later determined that it was probably made from meteorite iron that was processed~\cite{Wertime1980Jan,Wertime1973Nov,Matsui2022Apr}. Additionally, there was recently a discovery at the Red Sea that included records dating back to the time of the construction of the ``Cheops Pyramid'' under Khufu~\cite{Tallet2016,Tallet2017,Tallet2021}, which revealed diligent bookkeeping of stones and ships involved.

Therefore, despite esoteric myths and grandiose claims to the contrary, there are no reliable indications of any ``nonprosaic, exotic, out-of-time'' agent in human history. Of course, one can never exclude that the ``Nazca Lines'' are signs directed toward ``real'' compared to ``projected'' aliens; they may well be. However, at the moment, from published records, there is not a single artifact or fact that would indicate the presence of ``ancient aliens'' from any material objects and that could not have been created with a lot of people and ``a lot of sweat.''



\section{Why do UFOs crash? How common are crashes?}
\label{2023-UFO-part-Perception-crash-retreivals-wduc}

The obvious question of ``why do UFOs crash?'' has not been definitively answered. This is even more pressing given that if the extraterrestrial hypothesis is correct, these crafts might have to master physical environments of a much harsher type than the situation on our planet.

Some, such as William Bramley and Richard Dolan, point out that in comparison to the overall activity, only a tiny fraction of UFOs crash. This might be analogous to plane crashes: most of the time, planes do not crash, but occasionally they do.

Bramley~\cite[pp.~18-19]{Bramley1993Mar} estimates the chances of finding crashed ancient alien craft based on statistics from the Federal Aviation Administration. He assumes that if aliens have the same safety record as commercial jet aircraft, then there would be approximately one serious accident per one million flights. With the assumption that 2000 alien flights occur every year, which is approximately 5-6 per day, Bramley calculates that there would be only one crash or significant debris every $1000000/2000 = 500$ years, that is, every five centuries. Of course, assuming more flights per day, such as 500 per day, would result in approximately one crash every $1000000/(500 \times 365) \approx 5-6$ years. Bramley concludes that even if extraterrestrial crafts have been flying in Earth's skies for a long time, finding wreckage or debris is unlikely, and the best evidence we might have of their existence is eyewitness testimony.

Another option is that UFOs cannot deal with very specific situations, such as some types of radar pulses or with electromagnetic (or hypothetical other) pulses released during nuclear tests. A third option is ``gifting:'' just as US troops gifted items or food, or left behind things too difficult to remove, to indigenous populations in Melanesia during WWII, UFOs might come down on purpose, maybe to foster our technology.

I find the first theory plausible, but my confidence is weak. Of course, if there is no crash retrieval going on, then the entire issue disappears into thin air.

\section{Can UFOs be downed by electromagnetic pulses?}
\label{2023-UFO-part-Perception-crash-retreivals-htdau}

``Operation Dominic'' was a series of 31 nuclear test explosions conducted in 1962 in the Pacific.
Within this larger series of tests, ``Operation Fishbowl'' consisted of five high-altitude tests, and ``Starfish Prime''
was one of them.

On July 9, 1962, the United States detonated the Starfish Prime nuclear bomb in outer space as part of Operation Fishbowl.
The explosion, with a yield of approximately 1.4 megatons, occurred at an altitude of 250 miles (400 km) over a point 19 miles (31 km)
southwest of Johnston Atoll, approximately 700 nautical miles (1400 km) west of Hawaii.
The test aimed to study the effects of nuclear explosions in space on the Earth's magnetic field and determine
whether such explosions could be used as weapons against the Soviet Union.
The explosion generated strong electromagnetic pulses
that caused auroras to be seen across the sky and resulted in electronics failing in the Hawaiian Islands.
Additionally, the explosion created an artificial radiation belt that persisted for several years.

According to some unconfirmed reports and eyewitness accounts, the incident also resulted in a UFO fallout~\cite[Chapter~BDM - EMP, pp.~64f]{Omega_Point2022Nov}.
In Elizondo's words, the UFO came down like a brick ~\cite[time=38 min 44 sec]{Iandoli-Elizondo-Starfish-prime}.

\subsection{Personal account of a sailor}

Dolan quotes an unconfirmed
eye witness called David, who gave another recollection of maybe a similar event~\cite[time:~22:08 min:sec]{DolanDisclosure2020Jan}:
\begin{svgraybox}
``Lo and behold, what appeared in the sky---what I saw---was
a huge cigar shaped object coming toward us at ten degrees and approximately 15--20 degrees off the horizon.
It looked like it was two football fields in the length.
It appeared to have some kind of blue glow at the stern [[its back side, Dolan:
at its rear]]
as it was moving past us at a moderate rate.
My judgment is that it had to be traveling at least 400 miles per hour.
When it reached directly in front of us,
it picked up speed. It just went from there to gone.''
\end{svgraybox}

David Noble Whitecrow also recalls that during such tests, UFO sightings, both underwater and above sea level,
were quite common, but they usually vanished shortly before the tests.
David appeared in another series of interviews~\cite{Hein2018Dec,Hein2018Dec2} with his full name and gave a similar, more detailed account~\cite[time:~1:09:55 hour:min:sec]{Hein2018Dec}:
\begin{svgraybox}
``the order came from the bitch  box---that's that overhead ships command speaker---the order came for all hands on the starboard side to
look 90 degrees shift eyes forward 90 degrees off the starboard side and observe in silence, tracking visually what you observe in every detail
[[$\ldots$]]
dead silence, the surrounding side of the ship and the sea vanished, the huge object approaching the Finch off the starboard bow was not of this
world [[$\ldots$]] I went into brain freeze---I can only recall [[$\ldots$]] the moment in time
%[[$\ldots$]]
as I watch this immense unidentified object move downrange in the opposite direction of the Finch
[[USS Finch, DER 328, a radar picket escort vessel David was on]]
and it was like the fuselage---two football fields in length without props jet engines or wings---slowly moving down.
It was so immense I thought at the time why doesn't it fall, why doesn't it go
into the ocean?
[[$\ldots$]]
I first caught the sight of the object. It did not resemble any logical airborne aircraft
that I could recognize, I've been taught to recognize, as a lookout.
I had been trained to recognize as a watchstander surface ships
[[$\ldots$]] military or commercial aircrafts in sky watch experience.
Silence was broken by a whisper.
I could hear someone comment `what the  f-u-[[$\ldots$]] is that?'
The opposite charge of the group---he broke ranks and ordered us to report to the ship's galley,
for further instruction---before the strange objects slip downrange toward the horizon
I caught a short slight glimpse of the UFO as it slipped out of sight
downrange vanishing at the point of the horizon.''


``We were ordered to
the ship's galley by twos and we sat at the ship's galley tables by twos in order to sit and remain until ordered to the
ship's personnel office, which was down below the [[?]] deck.''

``I was still trying to make some sense of what I witnessed on deck moments before. No one had a good answer before ordered
to remain silent until our names were called to report to the personnel office.''

``The debriefing [[$\ldots$~?]] in it was my turn to describe what I have seen and witnessed the strange object,
 and I don't think a stenographer could keep up because the agitation and the gargle
and the confusion and our thoughts and words just didn't make sense---we had nothing to relate it to---and
so I just came out said:
`the strange object approach at a very slow speed [[$\ldots$]] I mean for its size and shape.
For a short time I watched [[it]]---and maybe
a minute to a minute and a half.
I wondered how could this huge cigar shape stay up so high [[$\ldots$]] above the sea.
That was my first thought. I could tell it was whitish in color without markings.
It was a cylinder with a smooth round front nose,
like a gigantic cigar tube---you know the aluminum cigar tubes [[$\ldots$]]
Only long---about one and a half to two football fields in length---and I'm just using that as a gauge because I used to watch football.
It was long, okay, and it was like 10 to 15 degrees above the horizon, above [[$\ldots$]]  the sea surface.'~''
\end{svgraybox}


David recounted a story about a military diver who had been tasked with searching for and exploring a sunken UFO. The diver was carried down to the ocean floor and quickly realized that there was a gigantic translucent object behind him. He made a slow 180-degree turn and faced the object, following instructions to get close to it and touch or kick it. He tentatively put his hand out to touch it and discovered that the object was not solid but was instead organic, resembling a gigantic jellyfish or sponge with a honeycomb-like surface. As soon as his hand penetrated the surface, he got the subjective impression that the object was going to ``suck him in.'' Although the object did not actually pull him in, he grew increasingly nervous and began to panic. He could see ``alien'' figures moving inside and shining through the illuminated object. When he finally surfaced, he vomited into his diving suit.

Because of his condition, the diver was flown to a military hospital in Hawaii.
He later heard rumors from sailors who had remained on his ship that the US, along with other nations such as Russia,
the UK, and France, had attempted to salvage the huge object he had encountered underwater. This cigar-shaped object was most likely
the same one that David had seen drifting by the Finch just days before. At one point during the night,
the object began to glow brightly inside the water, much like a neon tube. It then moved carefully so as not to damage
the surrounding ships above and quickly disappeared beneath the water's surface~\cite{Hein2018Dec}.


Years later, David and the diver accidentally met at a bar in Guantanamo Bay, commonly referred to as ``Gitmo.''
The diver told David that he had been working on a book about his encounter with the cigar-shaped craft.
Unfortunately, he never realized his plan because he was electrocuted in the bathroom just a week before his retirement.
Prior to this bizarre incident, the diver's personal yeoman, a sort of secretary, was killed in another skateboarding accident.
This nerd-like person had been typing up a transcript based on the loose leaf papers,
binders, and sketches that the diver had made~\cite{Hein2018Dec}.


\subsection{UFO coming down like a brick}


Here are excerpts from a transcribed~\cite[pp.~69,70]{Penetrator2022NovLooseThreads}
conversation between James Iandoli~\index{Iandoli, James} and  Lue Elizondo\index{Elizondo, Lue}
~\cite[time:~38:19 min:sec]{Iandoli-Elizondo-Starfish-prime}

\begin{svgraybox}
\noindent Iandoli: ``Just one crash retrieval case I don't hear talked about much: Starfish Prime. $\ldots $
They were doing a nuclear test to explode a nuclear warhead in space. And allegedly something
crashed in the ocean. And it was retrieved.''

\noindent Elizondo: ``$\ldots $ There may be some significance to [[electromagnetic pulses]]
EMPs. $\ldots $
This is at this point pure speculation
based on some potential observations made in the past. There may be some truth that an
electromagnetic pulse of energy can interfere with whatever this technology is and its propulsion. And
if it interferes with it, then you now~$\ldots $  have a very interesting scenario where whatever
is keeping these things up in the sky~$\ldots $  it no longer does that.
$\ldots $  And so now all of the sudden
if you will, the bubble pops. And so all of a sudden comes crashing down this object that has no wings,
no tail, no ailerons, no obvious signs of propulsion. And now it really does become a brick. And that
brick falls. Now all of a sudden, gravity has a say and Mother Nature takes over. And that's probably
all I'll say about that right now.''
\end{svgraybox}

The search for the downed UFO was also reflected in ``Sekret Machines Book 2: {A} Fire Within'' by Tom DeLonge and Andrew James Hartley~\cite{DeLongeHartley-SM2}. In this fiction, a second UFO appeared that vaporized the wreckage that had fallen with precise stabs of energy.




\section{Smith memo}
\label{2023-UFO-part-Perception-crash-retreivals-SM}
\index{Smith memo}

%The following paragraphs are excerpts of a 2019 dissertation on Canada's UFO history.


Wilbert Brockhouse Smith, a Canadian government official and electrical engineer, developed an interest in unidentified flying objects (UFOs) in the late 1940s after reading newspaper reports about them. These reports followed the 1947 Roswell UFO incident and the attention being given to the US military's projects Sign and Grudge, which were related to UFO sightings~\cite{MatthewHayes-PhD}.

In 1950, Smith traveled to the US and met with Dr. Robert Irving Sarbacher,\index{Sarbacher, Robert Irving}, an American physicist, a student of Einstein, and a defense consultant, at a radio broadcasting conference in Washington DC. According to Smith's notes, Sarbacher told him that UFOs are real, that the subject is the most highly classified in the US government, even more so than the hydrogen bomb, and that the technology used in UFOs is so advanced that the US government has not been able to replicate it. Sarbacher also claimed that UFOs do not originate from Earth. The reason for Smith's meeting with Sarbacher and the specifics of their conversation are not clear. However, it is believed that this meeting had a significant impact on Smith's views on UFOs.

The following document~\cite{SmithWilbertNotes}, written by Smith on 15 September 1950, is about his interview with Sarbacher. It can be found in the archives at the University of Ottawa as part of the Arthur Bray fonds. This document is believed to be a recollection of the interview, written by Smith after their meeting, rather than an official document. The document is not part of the national archives collection. Arthur Bray, an independent researcher of UFOs who became interested in Smith's story after the engineer's death, self-published two books about UFOs and mentioned Smith and his work in them. He obtained Smith's personal papers, which included copies of official documents from the Department of Transportation or their drafts, from Smith's widow and later donated his collection of materials on Smith to the University of Ottawa.

\begin{svgraybox}
\noindent  Smith: ``I am doing some work on the collapse of the earth's magnetic field as a source of energy, and I think our work may have a bearing on the flying saucers.

\noindent  Sarbacher: ``What do you want to know[[?]]''

\noindent  Smith: ``I have read [[Frank]] Scully's book [[``Behind the Flying Saucers'']]
on the saucers and would like to know how much of it is true.''

\noindent  Sarbacher: ``The facts reported in the book are substantially correct.''

\noindent  Smith: ``Then the saucers do exist?''

\noindent  Sarbacher: ``Yes, they exist.''

\noindent  Smith: ``Do they operate as Scully suggests, on magnetic principles?''

\noindent  Sarbacher: ``We have not been able to duplicate their performance.''

\noindent  Smith: ``Do they come from some other planet?''

\noindent  Sarbacher: ``All we know is, we didn't make them, and it's pretty certain they didn't originate on the earth.''

\noindent  Smith: ``I understand the whole subject of saucers is classified.''

\noindent  Sarbacher: ``Yes, it is classified two points higher even than the H-bomb.
In fact it is the most highly classified subject in the U.S. government at the present time.''

\noindent  Smith: ``May I ask the reason for the classification?''

\noindent  Sarbacher: ``You may ask, but I can't tell you.''

\noindent  Smith: ``Is there any way in which I can get more information, particularly as it might fit in with our own work?''

\noindent  Sarbacher: ``I suppose you could be cleared through your own Defense Department and
I am pretty sure arrangements could be made to exchange information. If you have anything to contribute we should be glad to talk it over,
but I can't give you any more at the present time.''
\end{svgraybox}

However, the context surrounding Wilbur Smith's meeting with Sarbacher is unclear, as this handwritten note~\cite{SmithWilbertNotes} is prefaced with the following line:
``Notes on interview through Lt/C. Bremner with Dr. Robert Sarbacher.''
This suggests that Smith may not have personally spoken with Sarbacher,
but rather passed on questions to the scientist through Bremner, who was a defense attach\'e at the Canadian embassy in Washington D.C.
It remains unclear whether Smith actually met
with Sarbacher in person or communicated with him through intermediaries such as Bremner.

On November 21, 1950, Smith sent a memorandum~\cite{SmithWilbertMemo} on the letterhead of the Canadian Department of Transportation to the Controller of Telecommunications. The memo discussed Smith's thoughts on flying saucers and mentioned that he had a meeting with a well-known American scientist. Smith had arranged this meeting through ``discreet enquiries'' made by the Canadian embassy staff in Washington.
During the meeting, Smith reported that he had received the following information:
\begin{svgraybox}
\begin{enumerate}
\item[a.] ``The matter is the most highly classified subject in the United States
Government, rating higher even than the H-bomb.
\item[b.] Flying saucers exist.
\item[c.] Their modus operandi is unknown but concentrated effort is being
made
by a small group headed by Doctor Vannevar Bush.
\item[d.] The entire matter is considered by the United States authorities to be
of tremendous significance.''
\end{enumerate}

[[In this memorandum Smith also states:]] ``I was further informed that the United States authorities are
investigating along quite a nubmer of lines which might possibly be
related to the saucers such as mental phenomena and I gather that they are
not doing too well since they indicated that if Canada is doing anything
at all in geo-magnetics they would welcome a discussion with suitably
accredited Canadians.''
\end{svgraybox}
This is remarkable, as such information was not mentioned in the aforementioned note quoting Sarbacher. It remains unclear on what basis Smith made those remarks.

Several individuals attempted to locate Robert Sarbacher,\index{Sarbacher, Robert} who was living in Florida at the time~\cite{DolanDisclosure2018Oct}. UFO investigator William Steinman was successful in reaching Sarbacher and had several phone conversations with him. As a result, Sarbacher wrote a two-page, single-spaced typed letter detailing his meeting with Smith. In this letter, dated November 29, 1983, Sarbacher confirmed the following~\cite{Sarbacher-83}:
\begin{svgraybox}
\begin{enumerate}
\item[1.] ``Relating to my own experience regarding recovered flying saucers, I had no
association with any of the people involved in the recovery and have no
knowledge regarding the dates of the recoveries. If I had I would send it to you.
\item[2.] Regarding verification that persons you list were involved, I can say only this:
John von Neuman\index{von Neumann, John} was definitely involved. Dr. Vannever Bush\index{Bush, Vannever}
was definitely involved, and I think Dr. Robert Oppenheimer\index{Oppenheimer, Robert} also.

My association with the Research and Development Board under Doctor
Compton during the Eisenhower administration was rather limited so that
although I had been invited to participate in several discussions associated with
the reported recoveries, I could not personally attend the meetings. I am sure
that they would have asked Dr. von Braun\index{von Braun, Wernher} and the others that you listed were
probably asked and may or may not have attended. This is all I know for sure.

\item[3.] I did receive some official reports when I was in my office at the Pentagon but all
of these were left there as the time we were never supposed to take them out of
the office.''
\end{enumerate}

[[On the second page of his letter Sarbacher writes:]]
``I recall the interview with Dr. Brenner of the Canadian Embassy. I think the answers I
gave him were the ones you listed. Naturally, I was more familiar with the subject
matter under discussion, at that time. Actually I would have been able to give more
specific answers had I attended the meetings concerning the subject. You must
understand that I took this assignment as a private contribution. We were called
``dollar-a-year men''. My first responsibility was the maintenance of my own business
activity so that my participation was limited.

About the only thing I remember at this time is that certain materials reported to have
come from flying saucer crashes were extremely light and very tough. I am sure our
laboratories analyzed them very carefully.

There were reports that instruments or people operating these machines were also of
very light weight, sufficient to withstand the tremendous deceleration and acceleration
associated with their machinery. I remember in talking with some of the people at the
office that I got the impression these ``aliens'' were constructed like certain insects we
have observed on earth, wherein because of the low mass the inertial forces involved in
operation of these instruments would be quite low.

I still do not know why the high order of classification has been given and why the
denial of the existence of these devices.''
\end{svgraybox}

Sarbacher led researchers to another individual,  Eric Walker~\cite{DolanDisclosure2019Mar}.
Walker\index{Walker, Eric} was a significant figure who served as the president of Penn State University and headed one of the most influential military
think tanks in the country, the Institute for Defense Analyses. He was also a friend of President Dwight Eisenhower and was still alive at the time of the investigation. According to Steinman, Walker confirmed the existence of crash retrievals and recoveries of bodies and attended meetings on the subject. He also verified the existence of an organization similar to MJ-12. However, he warned Steinman to stop his investigation and leave the matter alone, stating that it was an area that could not be pursued.

Smith claimed to be in regular contact with an alien called ``Affa'', receiving alien artifacts from the US government and having mastered anti-gravity.

\section{Corso's memories}
\label{2023-UFO-part-Perception-crash-retreivals-CM}

What follows are speculations and fantasies. This brief review concerns unconfirmed memories of Philip James Corso, \index{Corso, Philip James}, a former US Army officer who served in various positions within the military, including as a member of President Dwight D. Eisenhower's National Security Council. Corso is known for his claims about his involvement in the alleged recovery of extraterrestrial technology. In 1997, Corso copublished a book called ``The Day After Roswell,''~\cite{Corso1998Jun} in which he claimed that he had played a key role in the distribution of some of that technology into various military and civilian applications. This would be similar to the exploitation of recovered foreign technology, as Corso became Chief of the Pentagon's Foreign Technology desk in 1961 and was a Special Assistant to Lt General Arthur Trudeau, who headed Army Research and Development.

It is unclear to me what Corso's original statements in the book ``The Day After Roswell'' were as he coauthored the book with Birnes. Birnes' status is mentioned as ``with'' on the book cover. There exists another, apocryphal account~\cite{Corso}. However, the authenticity of this second account remains unclear.

\section{Stalin's craft}
At this point, I would also like to mention another highly speculative account of the Roswell incident mentioned in Annie Jacobsen's book ``Area 51: An Uncensored History of America's Top Secret Military Base''~\cite{Jacobsen2011}. Jacobsen's story is about five EG{\&}G engineers who, as told by one of them, claimed that the federal government does not officially acknowledge the existence of Area 51 due to a program related to the Roswell crash remains, which predated the establishment of the CIA facility, constructed in 1955.

According to this narrative, the name ``Area 51'' is not arbitrary but rather refers to the transfer of the Roswell crash remains from Wright-Patterson Air Force Base to a secret location in Nevada's desert in 1951. The crashed craft had been sent by Stalin with Russian writing imprinted on its interior. It had been the main focus of the project at Wright-Patterson Air Force Base. The secret behind the craft's ability to hover and fly had remained a mystery. As a result, Vannevar Bush instructed the team to reverse engineer the craft, taking it apart and reassembling it in an effort to discover its flight capabilities. Allegedly, that problem was solved.

\section{Wilson notes}
\label{2023-UFO-part-Perception-crash-retreivals-wn}

The following assumptions remain officially unconfirmed and speculative.
Indeed, the authenticity of this narrative has allegedly been denied by Admiral Thomas Ray Wilson:
Wilson himself called the notes ``fiction'' and stated, ``I wouldn't know Eric Davis if he walked in right now''~\cite{cox20,Cox2021Aug}.
If, however, the subject is authentic, this has profound consequences, the weight of which cannot be overestimated.

\subsection{Context}

The Wilson UFO Leak refers to a set of 13 pages, known as the Wilson notes~\index{Wilson notes}\cite{WilsonNotes-Davis}, which were written by Dr. Eric Davis in 2002 about an alleged meeting with Admiral Thomas Ray Wilson. However, Wilson has categorically denied that such a meeting ever took place\cite[Chapters9, 21, Ref.9.23]{Coulthart2021Aug} and has also claimed to have no recollection of the other individuals mentioned in the notes\cite{cox20, Cox2021Aug}.

Wilson is a retired vice admiral in the US Navy who served as the director of the Defense Intelligence Agency (DIA) from 1999 until 2002. The DIA is a federal agency responsible for collecting, analyzing, and disseminating military intelligence to support military planning and operations, as well as providing intelligence support to the President, the National Security Council, and other government agencies. Prior to serving as the DIA director, Wilson was the director of intelligence for the joint chiefs of staff and deputy director of intelligence for the joint chiefs. This position, known as J-2, is responsible for managing the collection, analysis, and dissemination of intelligence information, providing guidance and direction to the Joint Staff (JS) on intelligence matters, and serving as the principal intelligence advisor to the Chairman of the Joint Chiefs of Staff (CJCS) and the JS. Wilson retired from the Navy in 2002.

One would assume that J-2 would be privy to ``all state secrets'' and have access to even the Critical Nuclear Weapons Design Information (CNWDI) to be ``on top of the information chain.'' However, the Wilson notes reveal Wilson's frustrating attempts to access certain carve-out/waived special access programs executed by private contractors and dealing with extraterrestrial matters.

The documents were discovered among the personal effects of the late Dr. Edgar Mitchell, a prominent astronaut and MIT-trained aeronautics engineer who was the ``sixth man to walk on the moon.'' The significance of the documents lies in their confirmation of the existence of US projects related to the study of extraterrestrial technology and beings.

The notes describe an alleged meeting that Wilson previously had with UFO researchers Dr. Stephen Greer, Edgar Mitchell, and others in April 1997 (according to Dr. Edgar Mitchell, it took place on April 9, 1997~\cite[p.~1]{WilsonNotes-Davis}). In this first meeting---presumably arranged by Dr. Edgar Mitchell---they had discussed UFOs, UFO crash retrievals, and the existence of black budget programs to study the acquired technology. The notes further describe Wilson's subsequent attempts to locate such program(s) and his subsequent meeting with the ``gatekeepers'' of at least one program run by a private contractor---probably Lockheed Martin Corporation---that was attempting to reverse engineer an alien spacecraft. Davis and Wilson also discuss UFO and alien-related topics, including the existence of organizations tasked with studying extraterrestrial technology and bodies retrieved from crashed UFOs. The notes also mention the Roswell UFO incident and Area 51.

The Admiral Wilson memo was officially recorded during Congressional hearings in May 2022~\cite{WilsonNotes-Davis}, giving credibility to previously suspected fake information about UFO/UAP. This information had previously been circulated in various online groups and websites but is now part of the public record.

There has been some debate over the authenticity of the notes. Many researchers believe that they are genuine. For instance, Christopher Mellon\index{Mellon, Christopher Karl} corroborated authenticity by stating that ``Dr. Davis, [[is the]] author of the famous Wilson-Davis memo''~\cite{Mellon2022Dec}. He is a former Deputy Assistant Secretary of Defense for Intelligence in the United States government. Mellon served in this position from 1997 to 2002, during which time he oversaw the Department of Defense's intelligence programs and activities related to national security. After leaving the government, Mellon worked in the private sector as a security and intelligence consultant and has been involved in efforts to investigate and disclose information about unidentified flying objects.

It might not be totally unreasonable to speculate---but far from certain---that, carefully drafted, leaked, and exposed by a certain fraction of persons with some knowledge of the subject matter, the Wilson memo was purposefully intended as a sort of ``screwdriver'' to open up the can of (reconstructed) crafts and crash retrievals.

\subsection{Excerpts from the Wilson notes}

By now, the reader should be familiar with the categories of the US Department of Defense, specifically carve-out, waived, and unacknowledged special access programs (USAPs). The following excerpts from the Wilson memo are a slightly annotated (annotations marked by double brackets ``[[$\ldots$]]'') walkthrough of these notes. Its essential and smoothed thread exposes the most critical issues raised.

\begin{svgraybox}
On page 6 Eric W. Davis (EWD) asks Thomas R. Wilson (TW): ``Okay then, what happened in April - June '97?''

TW (Wilson) reponds: ``$\ldots$~I made calls, knocked on a few doors, talked to people - went on for 45 days (thereabouts) on and off''

On page 7 TW (Wilson) continues:
\begin{enumerate}
\item[-]
``Suggestion came from Ward (Gen. M. Ward) to go through the records groups
files (like an index system [[a list of all SAPs]]) in OUSDAT\index{OUSDAT} (Office of the Under Secretary of Defense for Acquisition and Technology)
\item[-]
Ran into Bill Perry [[William James Perry  was the US Secretary of Defense from February 3, 1994, to January 23, 1997]] in May '97---talked
about this quietly---he suggested the same thing
\item[-]
They told me of a special projects record group not belonging to usual SAP---a
special subset of the unacknowledged/carve-outs/waived programs---not
belonging to usual SAP divisions as organized in '94 by Perry himself---set
apart from rest but buried/covered by conventional SAPs''
\end{enumerate}
$\ldots$

\noindent TW (Wilson): ``So found the unusual record group---read the index abstracts [[of those carve-out/waived USAPs]].''
$\ldots$



On page 8 EWD (Davis) asks: ``Change subject: so what SAP compartment did you find in?''

\noindent TW (Wilson): ``Core secret - won't say''

\noindent EWD (Davis): ``Code name?''

\noindent TW (Wilson): ``Again won't say - core secret''

\noindent EWD (Davis): ``Who was the project contractor or USG agency that runs program?''

\noindent TW (Wilson): ``An aerospace technology contractor - one of the top ones in US''

\noindent EWD (Davis): ``Who?''

\noindent TW (Wilson): ``Core secret - can't tell''

\noindent EWD (Davis): ``Defense contractor?''

\noindent TW (Wilson): ``Yes, the best one of them.''

\noindent EWD (Davis): ``Intelligence too?''

\noindent TW (Wilson): ``In their corporate portfolio''

\noindent EWD (Davis): ``Give a hint?''

\noindent TW (Wilson): ``Sorry, no.''

Let us suppose for a moment that this ``best defense contractor'' is Lokheed Martin. This is a mere speculation~\cite{Dolan-MrX-Disclosure2020Jul}

\noindent EWD (Davis): ``What happened when you found contractor?''

\noindent TW (Wilson): ``I made several calls (end of May '97), first to Paul
[[maybe Paul Garrett Kaminski~\cite{cox20,Dolan-MrX-Disclosure2020Jul}?]], Mike [[maybe Michael Costanik?
Both Kaminski and Costanik may have members of the previous leadership team of the senior review group~\cite{Dolan-MrX-Disclosure2020Jul}]]
{\&} Perry [[maybe William James Perry~?]] to confirm I had right contractor and program manager to talk to.''

\noindent EWD (Davis): ``They confirm?''

\noindent TW (Wilson): ``Yes.''

Page 9 continues with EWD (Davis) asking Wilson: ``Then?''

\noindent TW (Wilson): ``(End of May '97) Made three calls to the program manager---one of them conferece call with security director and corporate attorney.
[[There was]] -- Confusion on their part as to why I was looking for them and what I wanted from them or wanted to know about.
-- Very testy tone from all of them.''

\noindent EWD (Davis): ``What do you mean? ''

\noindent TW (Wilson): ``They were agitated about my calling---surprised by call''

\noindent EWD (Davis): ``What [[did]] you asked them?''

\noindent TW (Wilson): ``Yes.''

\noindent EWD (Davis): ``What was that---what words?''

\noindent TW (Wilson): ``I told I read their program record in the OUSDAT\index{OUSDAT} [[Office of the Under Secretary of Defense for Acquisition and Technology]]
special program records group  and wanted to know about their crashed UFO program, what their role in that was, what
they had, etc. Also asked if they heard of MJ-12 or some such organization code relating to crashed/recovered UFO craft.''

\noindent EWD (Davis): ``Reaction on phone to that? ''

\noindent TW (Wilson): ``Yes! Asked who I talked to before I called them so I told them and they weren't happy with that answer.''

\noindent EWD (Davis): ``You mean about Perry, Paul, etc.?''

\noindent TW (Wilson): ``Oh, no! I didn't tell them [[that]] I talked to those guys!''

$\ldots$

On page~10 EWD (Davis) ask Wilson: ``What then?''

\noindent TW (Wilson): ``I told threesome I wanted formal briefing, tour, etc.---was exploiting my regulatory
authority as Deputy Director DIA/Assistant Joint Chief of Staff J-2---Told them my not
being briefed was oversight they needed to correct---I demanded!''

\noindent TW (Wilson): ``They needed to discuss this (his demand) so hung up. Got called 2 days later and
they said they don't want to talk on phone and arranged for face-to-face meeting at their  facility.''

\noindent EWD (Davis): ``Did you go?''

\noindent TW (Wilson): ``Yes, ten days later (mid-June or so). Flew out there -- Met in their conference room in their secure vault -- Three of them show up''

\noindent EWD (Davis): ``3 guys with whom you had telecon?''

\noindent TW (Wilson): ``Yes, same 3
\begin{enumerate}
\item[-]
Security director (NSA-retired, a CI expert)
\item[-]
Program director
\item[-]
Corporate attorney
\item[-]
Called themselves ''the watch committee,'' or gate keepers
\end{enumerate}

On page 11 TW (Wilson) mentions: ``[[$\ldots$]] they weren't going to let me in the door''

\noindent EWD (Davis): ``Why?''

\noindent TW (Wilson): ``They said my tickets were all confirmed and valid, but I was not on the bigot list
\begin{enumerate}
\item[-]
My tickets alone were not enough
\item[-]
I didn't meet the special criteria so need to know authorization was not being granted
\item[-] Went back 'n forth with them over these points (primarily with security director {\&} attorney)
\end{enumerate}

Page 12 continues with \noindent TW (Wilson): ``Argued more---they wouldn't accept my arguments that they fell under my statutory
oversight and regulatory authority as Deputy Director DIA---under purview for my right
to have need to know (oversight, audit, justification issues, etc., etc.)
\begin{enumerate}
\item[-]
Regulatory and statutory authority as Deputy Director DIA not relevant or pertinent to nature of their program!
\item[-]
Then they pulled out their bigot list to convince me otherwise - several pages long-dated 1990, updated 1993.''
\end{enumerate}
$\ldots$


\noindent TW (Wilson):
\begin{enumerate}
\item[-]
``Program Manager said they were
\begin{enumerate}
\item[$\circ$]  Not any weapons program
\item[$\circ$] Not any intelligence program
\item[$\circ$] Not any special ops or logistics program
\item[$\circ$] Doesn't fit these categories
\end{enumerate}
\item[-] I asked what they were then
\item[-] Loud groan from Program Manager
\item[-] Security Director and attorney say it's okay to say it.''
\end{enumerate}

\noindent EWD (Davis): ``Say what?''

\noindent TW (Wilson): ``They were a reverse engineering program---
\begin{enumerate}
\item[-]
Something recovered years ago in the past
\item[-]
Technological hardware was recovered
\item[-]
So I thought they meant recovered Soviet/Chinese, etc. hardware and reverse
engineer it---like a missile or intell platform or aircraft---actually came to
meeting expecting to find a sensitive foreign collection and reverse
engineering operation---thought ``UFOs'' used as a cover for that
\item[-]
So I said that and they said they weren't that either
\item[]page 13
\item[-]
They had (program manager talking) a craft---an intact craft they believed
could fly (space? air? water? dimensions?)
\item[-]  Was it from overseas or not?
\item[-]  Said NO! Could not be-not possible!!
\item[-]  Why, I asked-where did it come from?
\item[-]  Program manager said they didn't know where it was from [they had some ideas on this] -- it was technology that {\it  was not of this Earth -- not made by man -- not by human hands}
\item[-] Said were trying to understand and exploit technology; their program was going on for years and years with very slow progress
\item[-] Agonizingly slow with little or no success -- painful lack of collaboration to get help from outside community of experts and facilities to assist
effort -- must remain isolated and use own facilities and cleared personnel -- tough
environment to work -- about 400-800 (bigot list count) workers varying in number with funding or personnel changes
\item[-] Miller questions asked:\begin{enumerate}
\item[$\circ$] Roswell -- craft/bodies/autopsies
\item[$\circ$] Holloman AFB Landing
\item[$\circ$] MJ-12 and leaked docs
\item[$\circ$] Zamora {\&} Bentwaters, etc.
\end{enumerate}
\item[-] They were mum - declined to discuss these
\item[-] TW [[Wilson]] threatened to go to SAPOC\index{SAPOC} [[Special Access Program Oversight Committee]] to complain, gain access to their program
\item[-] They said go ahead and do what you must
\item[-] I was angry because they defied my authority to be read-in with good logical reason---wouldn't budge
\item[-] Their tone was very testy/terse throughout conversation''
\end{enumerate}

\noindent EWD (Davis): ``What was outcome?''

\noindent TW (Wilson): ``Meeting broke up and I returned to Washington''

$\ldots$

\noindent EWD (Davis): ``Did you complain to SAPOC [[Special Access Program Oversight Committee]]?''\index{SAPOC}
\noindent TW (Wilson): ``Yes---called the subgroup members (Senior Review Group members) to a meeting at Pentagon
\begin{enumerate}
\item[-]
Told them what happened at meeting
\item[]page 14
\item[-]
They responded that they would sustain the contractor on their access denial
\item[-]
So I ended up arguing with them a while
\item[-]
Broke up in 20 minutes and they would meet me in 2-3 days
\item[-]
Got the call 2 days later (near end of June) and met again with Senior Review Group members''
\end{enumerate}

\noindent EWD (Davis): ``When?''

\noindent TW (Wilson): ``Before last week of June (' 97)''
\begin{enumerate}
\item[-]  They told me (TW [[Wilson]]) that they were sustaining the contractor, that I was to
immediately drop the matter and let it go -- forget about it as I did not have
purview over their project, it didn't fall within my oversight, etc.
\item[-]  I became very angry---started yelling when should have kept my mouth shut''
\end{enumerate}

\noindent EWD (Davis): ``Miller and Greer said you nearly got busted.''

\noindent TW (Wilson): ``Close to that!

$\ldots$

Why such a big deal over this considering the position of trust I have in the
Pentagon---I do have relevant regulatory/statutory authority over their program---that's my position!!!''

\noindent EWD (Davis): ``Is it because funding comes from you or through you? Or Director DIA?''

\noindent TW (Wilson): ``Core secret---can't answer''

\noindent EWD (Davis): ``Back to bigot list-can you describe type of people?''

\noindent TW (Wilson): ``Corporate types---scientists and technicians- engineers, scientists, managers, etc.''

Page 15 closes by stating that Wilson ``[[w]]ill retire to Midwest and close office in 2003.''
\end{svgraybox}

I am not sure of how to read and interpret these lines, but assuming this conversation had actually taken place, it appears more fascinating than a central scene in a science fiction movie: a high-ranking officer being told by his elected masters and an industry contractor to ``forget about it'' as he ``did not have purview over their project.'' How must it feel to be told to suppress that the US possesses, analyzes, and reproduces technology that is ``not of this Earth--not made by man--not by human hands?'' How should we cope with such a possibility?


\section{Where would alien stuff most likely reside?}
\label{2023-UFO-part-Perception-crash-retreivals-where}


What follows are speculations and fantasies.

Suppose for a moment that the USA was in possession of ``alien stuff'' and artifacts. Then, ask yourself: How would they be categorized and dealt with? Where would this material be located?

Suppose for a moment that the USA attempted to ``process'' alien artifacts in a similar manner to Nazi and other Earth-bound (e.g., Soviet) technology. It would not be unreasonable to speculate that early on, the US Air Force handled most of these efforts, coordinated by what was then the Air Materiel Command at Wright-Patterson Field, Ohio. From there, the material might have been distributed to various sites in industry and government.

Currently, such material and information, if administered by the DOD, would be buried away in carve-out USAPs. That means that they might be passed on to private corporations such as Lockheed Martin or Battelle. If they were in the possession of the DOE, they might also end up in private hands. However, they might have also been locked away in national laboratories such as Los Alamos National Lab (LANL) or Lawrence Livermore National Laboratory (LLNL), or Sandia National Laboratories, a long time ago.

If one believes in Philip Corso's original manuscript~\cite{Corso}, he reportedly gave alien artifacts and technology similar to what companies were already working on to these companies for them to analyze for a time. He did not disclose where the technology came from or any other information about it. He asked the companies to pretend that they had developed any ideas or innovations based on the technology themselves and to offer the US military first access to any resulting technology.

%\section{Hoe common would be a crash}
%\label{2023-UFO-part-Perception-crash-retreivals-hvwc}




\label{2023-UFO-part-Perception-crash-retreivals-wkoam}

Corso claimed that he was told that in some materials that he allegedly recovered and distributed to US corporations and research institutions, the atoms and molecules were ordered in one direction -- like in spider silk -- which was unlike anything they had ever seen before~\cite{Corso1998Jun}.

While some technologies that were on the verge of discovery have made some advances, central alien technologies such as anti-gravity or
propulsion could not be reverse-engineered, as discussed in Appendix~\ref{2023-UFO-Appendix-Explanation-Trap}.
This can be attributed to an explanation trap caused by conceptual and theoretical overreach.
Without an ``alien Prometheus'' to explain the semantics and associated scientific principles necessary for comprehending more advanced aspects of alien technology,
such technology might have been far too advanced for human science to understand.
Human scientists, in their expert pose, may act like shamans explaining an airplane to an indigenous tribe.

Nat Kobitz, the former Navy Director of Science and Technology, told Ross Coulthart~\cite{Coulthart2021Aug} that at Wright-Patterson Air Force Base, he saw a piece made out of a titanium alloy not known to the Air Force. Kobitz did not know what the piece was or where it came from. It was approximately three by four feet in size and showed no attachment except for something that appeared to be welded to it. However, this something was actually integral with the material, like a bulkhead to the skin, as if it were cast. According to Kobitz's expert opinion, there was no known industrial process that could replicate the bond he saw on that curious fragment.






A book by Tom Carey and Don Schmitt~\cite{CareySchmitt} suggests that materials from the Roswell\index{Roswell incident} crash were first transferred to Wright-Patterson Air Force Base, which was then called Air Force Technical Base. Later, the Battelle Memorial Institute was contracted for the reproduction of a shape-recovery alloy.

Following the alleged 1947 Roswell UFO crash, secret studies on previously unknown material were conducted to examine the debris reported by witnesses. This material is known as ``memory metal'' or Nitinol, which is a shape-recovery alloy that can instantly return to its original state after being deformed. The studies were conducted by the Battelle Memorial Institute in Ohio under the direction of a metallurgist who also worked on the US Air Force's official UFO study, Project Bluebook. The article suggests that the discoveries from these studies were then utilized for further technology development by other government agencies and military contracts to universities and industry. It is also suggested that the Wright-Patterson Air Force Base was the location where the Roswell UFO debris was flown after the crash, and the Battelle Memorial Institute was contracted by Wright-Patterson to perform the studies.



\section{Could we hack ourselves in their database?}
\label{2023-UFO-part-Perception-crash-retreivals-hack}


An additional option to the aforementioned ``alien Prometheus'' explaining their technology to us personally may be to locate their ``computer,''
that is, the information storage and handling facility, in a recovered craft.
Most likely, this will be guarded and provided by an advanced form of artificial intelligence.

If we were able to access and penetrate it, we might be able to recover
its information content---a sort of encyclopedia galactica---which would also include their means of propulsion and energy source.
But this maybe hopeless: their technology might be so advanced that our own computers would seem like mere abacuses in comparison.
Attempting to hack into their systems would be like trying to operate our own computers with an abacus---an
endeavor that would highlight the vast technological gulf between our civilizations.

A first attack on such a database would be to identify its hardware. Suppose for a moment we had retrieved a (partially) intact flying saucer from a crash.
How should we go about in identifying potential storage facilities in such a craft?
I suggest that, as any such storage can be assumed to be ``very advanced'' its material layer can be expected to be ``very complex''---that is, the information density stored per ``stuff'' is high.
That in turn means that such a memory device, if analyzed on an atomic level, should appear to be highly irregular, randomly aligned, akin to white noise.
There is almost paradoxical feature of highly organized, dense coding: the ciphertext appears random in the
very precise sense of algorithmic information theory (with randomness roughly meaning algorithmic incompressibility.
So, one criterium to identify storage on such a recovered craft would be that it appears disorganized and ``amorphous.''


To identify potential storage facilities in a recovered flying saucer, we must first determine its hardware.
Given that any such storage is likely to be highly advanced, we can assume that its material layer is also highly complex.
This means that the information density stored per unit of material is high, and any memory device would appear irregular and randomly aligned at an atomic level,
similar to white noise.

However, this paradoxical feature of highly organized, dense coding also means that the ciphertext appears random in the very precise sense
of algorithmic information theory~\cite{calude:02}). Informally speaking, it would be algorithmically incompressible, making it difficult to identify
using conventional methods.

Therefore, a key criterion to identify storage on a recovered craft would be its disorganized and ``amorphous'' appearance.
By looking for areas of the craft that appear structurally irregular and chaotic, we may be able to locate potential storage devices.
Additionally, we could use advanced scanning techniques to analyze the material at an atomic level and identify any areas of high information density.
Overall, a multi-faceted approach
that combines structure analysis of the material with algortithmic pattern analysis
would be necessary to successfully identify and cryptanalize storage facilities in a recovered flying saucer.

One argument against conventional search for extraterrestrial intelligence (SETI) is that any signals emitted by a civilization that are identifiable through conventional analysis might only be detectable for a relatively short period, perhaps just one or two centuries. This is due to the fact that as technology progresses and broadcasts become digitized, the emitted signals tend to approach white noise, making them much more difficult to analyze and decipher. Therefore, traditional SETI approaches that rely solely on detecting narrowband radio signals may
not be effective in detecting advanced civilizations that have moved beyond using this type of technology.

\section{Human ``implants'' as alien material?}


Roger Leir allegedly performed numerous extractions of implants from individuals who claimed to have had abduction experiences and believed these objects to be of extraterrestrial origin~\cite{Leir2000}. Typically, these implants were smaller than a grain of rice~\cite{DolanDisclosure2018Sep}. It would be intriguing to examine these ``implants'' as potential ``alien'' materials and subject them to detailed analysis. However, all objects investigated thus far have been found to have mundane origins~\cite{ScottLittle-implant}.



\section{Speculations of a possible involvement of Einstein}
\label{2023-UFO-part-Perception-crash-retreivals-einstein}

\subsection{Memories of ``Einstein's pet''}

I hesitated for some time to include this very speculative addendum to an already very speculative chapter.
There have been claims that Einstein was invited to a secret trip to a base---presumably Wright-Patterson Field, Ohio---where
the remains of the Roswell crash was investigated.
Allegedly, he took a young female student with him (no affirmative account of Einstein is available):
Shirley Wright~\cite{ShirleyWrightObituaryLegacy2015Jul}, who later earned two PhDs in Physical Chemistry and Physical Science,
and still much later recalled her trip with Einstein.
In her own description, she
called herself ``Einstein's pet,''
chosen among a group of students meeting and working with Einstein in the summer of 1947~\cite{BragaliaEinstein}.

Wright gave an unconfirmed oral statement that partially survived due to the flow of information from three UFO research enthusiasts and authors
Sheila Franklin, Leonard Stringfield and Anthony Bragalia.
According to Shirley Wright's alleged testimony, she saw a badly damaged vehicle---one is reminded of the alleged
``tear-drop spaped escape pod'' mentioned by Carey and Schmitt~\cite{CareySchmitt}.
She also recalls having seen five, and in another location nine aliens, at least one of them still alive~\cite{BragaliaEinstein}.
Wright was not permitted to enter the disk-shaped vehicle, as in her personal recollection she was only a
``gopher $\ldots$ a nobody''
supposed to assist Einstein.
The communication with the alien was telepathic. ``They wanted to know how long we lived. $\ldots$ They were very quick to let us know we did not know anything
$\ldots$ Why would they come to Earth? $\ldots$ One ship he claim crashed in Siberia $\ldots$ definitely not from our galaxy $\ldots$
I wondered whether somebody playing a terrible untruth to us $\ldots$''

\subsection{Einstein's letter of ignorance}

In the midst of the 1952 UFO craze (or flap, depending on your preferences), Einstein refused to discuss flying saucers:
in a letter dated July 23~\cite{EinsteinGardner1952}, 1952, Einstein wrote that ``those people have seen \underline{something}.
What it is I do not know, and I am not curious to know.'' This is an amazing statement, given the potential impact on his own opus magnus,
general relativity.

One might speculate that, at the time, Einstein had already phased out thoughts on general relativity.
In 1949, Einstein might also have gotten weary over
Kurt G\"odel's 1949 solution to the field equations of general relativity~\cite{godel-rmp} which described a
spacetime with some unusual properties. This G\"odel universe permitted closed timelike curves; hence, a
kind of time travel and thereby might be seen as not admitting a forward lapse of time by successive moments of time.
In the same year,
G\"odel published ``A Remark about the Relationship between Relativity Theory and Idealistic Philosophy''~\cite{godel-sch}
in which he used certain properties of this solution to argue for a kind of temporal idealism,
whereby ``it seems that one obtains
an unequivocal proof for the view of those philosophers who,
like Parmenides, Kant, and the modern idealists, deny the objectivity of change and consider change as an illusion or an
appearance due to our special mode of perception.''

Therefore, late in his life, Einstein might have been fully aware of the explanation trap encountered
with regard to flying saucers.
Indeed, in addition to Manhatten Project celebrities such as von Neumann and Oppenheimer,
one might think that Einstein would have been the first to be adressed and called in case of crashed UFOs
(maybe in spite of FBI's concerns that he might have been a double agent, anarchist and pacifist).
